### 2020-02-13 src/eba_sims_w_data.Rmd
First scripts

---
title: "eba_sims_data"
author: "Esa Turkulainen"
date: "1/13/2020"
output: pdf_document
---

```{r setup, echo = FALSE}
setwd("~/eba_sims/src")
library(ggplot2)
library(gridExtra)
library(knitr)
library(plyr)
library(lubridate)
library(numbers)
library(data.table)
library(R.utils)
library(forecast)
source("pffunctions.R")
```

```{r load_data, echo = FALSE}
load("~/eba_sims/data/20191211CollectionAmounts.rdata")
```

## Inspect data
```{r inspect}
summary(donation10)
```

```{r filtering_function, echo = FALSE}
# Let's build a tool to filter quickly as we please
# TODO: Let the user select multiple levels for each factor
filter_donations <- function(data, 
                             donor=NULL, donation13=NULL, site=NULL,
                             date=NULL, phleb_start=NULL, status=NULL,
                             donat_phleb=NULL, Hb=NULL, gender=NULL,
                             dob=NULL, aborh=NULL, age=NULL,
                             agegroup=NULL, Hb_deferral=NULL){
  temp <- data
  if(!is.null(donor)){
    temp <- temp[temp$donor==donor,]
  }
  if(!is.null(donation13)){
    temp <- temp[temp$donation13==donation13,]
  }
  if(!is.null(site)){
    temp <- temp[temp$site==site,]
  }
  if(!is.null(date)){
    temp <- temp[temp$date > date[1] & temp$date < date[1],]
  }
  if(!is.null(phleb_start)){
    temp <- temp[temp$phleb_start==phleb_start,]
  }
  if(!is.null(status)){
    temp <- temp[temp$status==status,]
  }
  if(!is.null(donat_phleb)){
    temp <- temp[temp$donat_phleb==donat_phleb,]
  }
  if(!is.null(Hb)){
    temp <- temp[temp$Hb > Hb[1] & temp$Hb < Hb[2],]
  }
  if(!is.null(gender)){
    temp <- temp[temp$gender==gender,]
  }
  if(!is.null(site)){
    temp <- temp[temp$site==site,]
  }
  if(!is.null(dob)){
    temp <- temp[temp$dob > dob[1] & temp$dob < dob[2],]
  }
  if(!is.null(aborh)){
    temp <- temp[temp$aborh==aborh,]
  }
  if(!is.null(age)){
    temp <- temp[temp$age > age[1] & temp$age < age[2],]
  }
  if(!is.null(agegroup)){
    temp <- temp[temp$agegroup==agegroup,]
  }
  if(!is.null(Hb_deferral)){
    temp <- temp[temp$Hb_deferral==Hb_deferral,]
  }

  return(temp)
}
```

```{r filter_test}
# Only donat_phleb="K" and status="-"
dons <- filter_donations(data = donation10, donat_phleb = "K", status = "-")

```

```{r distribution_of_event_counts}
eventcounts <- table(dons$donor)
hist(eventcounts)
```
Sanity check: passed. (60 is the maximum number of donations in 10 years time)

```{r}
head(dons)
```

```{r}
dons <- dons[with(dons, order(date)), ]
firsts <- dons[!duplicated(dons$donor),]
```

```{r}
firsts[1:20,]
```

```{r}
counts <- data.frame()
weeks <- c()
for (i in 1:dim(firsts)[1]) {
  stamp <- paste0(year(firsts[i,]$date), '.', week(firsts[i,]$date))
  if(!(stamp %in% weeks)){
    weeks <- c(weeks, stamp)
    counts[stamp, "Count"] <- 1
  }
  else{
    counts[stamp, "Count"] <- counts[stamp, "Count"] + 1
  }
}
```

```{r}
ggplot(data = counts, aes(x=row.names(counts), y=Count)) + 
  geom_point(color = "dark red") +
  theme(axis.text.x = element_text(angle = 45)) +
  scale_x_discrete(breaks=c("2010.1", "2011.1", "2012.1", "2013.1", "2014.1", "2015.1", "2016.1", "2017.1", "2018.1", "2019.1")) +
  theme_bw() +
  labs(title = "Number of weekly first time donations",
       subtitle = "Assuming 2010 as anomalous, the weekly donations fall to the 500 range with small seasonality.",
       x = "Weeks",
       y = "Count")
```

### 2020-02-17 src/eba_sims_w_data.Rmd
update file

@@
 library(knitr)
+library(dplyr)
 library(plyr)
@@
 ```{r}
+# Function to transfrom dates to year.week 
+yearweek <- function(x = Sys.Date()) {
+  xday <- ISOdate(year(x), month(x), day(x), tz=tz(x))
+  dn <- 1 + (wday(x) + 5)%%7
+  nth <- xday + ddays(4 - dn)
+  jan1 <- ISOdate(year(nth), 1, 1, tz=tz(x))
+  return(sprintf("%s.%d", format(nth, "%Y"), 1 + (nth - jan1)%/%ddays(7)))
+}
+```
+
+```{r}
+# Order donations by date
 dons <- dons[with(dons, order(date)), ]
+
+# Transform dates to yearweeks
+dons$date <- yearweek(dons$date)
+
+# Extract first time donations
 firsts <- dons[!duplicated(dons$donor),]
@@
 ```{r}
+# Check
 firsts[1:20,]
+```
+```{r}
+# Attempting vectorization of the processes below
+
+# Firsts
+first_count <- as.data.frame(table(factor(firsts$date, levels=unique(firsts$date))))
+# (We needed to call factor() to retain the chronological order of the rows)
+colnames(first_count) <- c("week", "firsts")
+
+# All
+all_count <- as.data.frame(table(factor(dons$date, levels=unique(dons$date))))
+colnames(all_count) <- c("week", "all")
 ```
@@
 ```{r}
-counts <- data.frame()
-weeks <- c()
-for (i in 1:dim(firsts)[1]) {
-  stamp <- paste0(year(firsts[i,]$date), '.', week(firsts[i,]$date))
-  if(!(stamp %in% weeks)){
-    weeks <- c(weeks, stamp)
-    counts[stamp, "Count"] <- 1
-  }
-  else{
-    counts[stamp, "Count"] <- counts[stamp, "Count"] + 1
-  }
-}
+# Combine into one neat df
+counts_df <- data.frame(week=first_count$week, 
+                        first=first_count$firsts, 
+                        all=all_count$all)
 ```
@@
 ```{r}
-ggplot(data = counts, aes(x=row.names(counts), y=Count)) + 
+# Plot
+ggplot(data = counts_df, aes(x=week, y=first)) + 
   geom_point(color = "dark red") +
+  geom_point(aes(x=week, y=all), color = "pink") +
   theme(axis.text.x = element_text(angle = 45)) +
+  geom_abline(intercept = mean(counts_df[158:520,2]), slope=0) +
+  geom_abline(intercept = mean(counts_df[158:520,3]), slope=0) +
   scale_x_discrete(breaks=c("2010.1", "2011.1", "2012.1", "2013.1", "2014.1", "2015.1", "2016.1", "2017.1", "2018.1", "2019.1")) +
   theme_bw() +
-  labs(title = "Number of weekly first time donations",
-       subtitle = "Assuming 2010 as anomalous, the weekly donations fall to the 500 range with small seasonality.",
+  labs(title = "Number of weekly first time donations vs. all donations",
+       subtitle = paste0("Each week around ", round(mean(counts_df[158:520,2])/mean(counts_df[158:520,3])*100, 2), "%", " of all donations comprise first time donors (2013 onwards)."),
        x = "Weeks",
-       y = "Count")
+       y = "Count") +
+  annotate(geom="text", 
+           x="2019.1", 
+           y=mean(counts_df[158:520,2])+400, 
+           label=paste0("Firsts: ", round(mean(counts_df[158:520,2]), 1))) +
+  annotate(geom="text",
+           x="2019.1",
+           y=mean(counts_df[158:520,3])+400, 
+           label=paste0("All: ", round(mean(counts_df[158:520,3]), 1)))
 ```

### 2020-05-06 src/eba_sims_w_data.Rmd
cost surface functions

@@
 ```{r setup, echo = FALSE}
-setwd("~/eba_sims/src")
+setwd("~/interval-sims/src")
 library(ggplot2)
+library(plotly)
 library(gridExtra)
@@
 source("pffunctions.R")
+source("helper_functions.R")
 ```
@@
 ```{r}
-# Function to transfrom dates to year.week 
-yearweek <- function(x = Sys.Date()) {
-  xday <- ISOdate(year(x), month(x), day(x), tz=tz(x))
-  dn <- 1 + (wday(x) + 5)%%7
-  nth <- xday + ddays(4 - dn)
-  jan1 <- ISOdate(year(nth), 1, 1, tz=tz(x))
-  return(sprintf("%s.%d", format(nth, "%Y"), 1 + (nth - jan1)%/%ddays(7)))
-}
-```
-
-```{r}
 # Order donations by date
@@
 ```
+
 ```{r}
@@
 ```
+
+## Cost analysis
+
+```{r}
+actual.deferral.rate <- 76565/2464190
+```
+
+```{r}
+p <- 0.031
+price_ratio <- 200
+surface <- cost_surface(p, price_ratio)
+plot_surface(surface, price_ratio, p)
+plot_interactive_surface(surface)
+```

### 2020-05-06 src/helper_functions.R

@@ -1,0 +1,61 @@
+generate_population <- function(n, prob){
+  rand <- runif(n)
+  population <- as.integer(rand <= prob)
+  
+  return(population)
+}
+
+# Function to transfrom dates to year.week 
+yearweek <- function(x = Sys.Date()) {
+  xday <- ISOdate(year(x), month(x), day(x), tz=tz(x))
+  dn <- 1 + (wday(x) + 5)%%7
+  nth <- xday + ddays(4 - dn)
+  jan1 <- ISOdate(year(nth), 1, 1, tz=tz(x))
+  return(sprintf("%s.%d", format(nth, "%Y"), 1 + (nth - jan1)%/%ddays(7)))
+}
+
+cost_surface <- function(p, price_ratio){
+  
+  # Cut FPR at 100x%
+  x = 0.5
+  
+  # Set axis 
+  tp.axis <- seq(0, (p-p/100), p/100) # TPs go from 0 to p
+  fp.axis <- seq(0, (1-p)*x - ((1-p)*x)/100, ((1-p)*x)/100) # FPs need to be limited: they explode the cost (due to reinvites) approaching 1-p
+  
+  # Set grid
+  cost <- matrix(, nrow = length(tp.axis), ncol = length(fp.axis))
+  
+  tpi <- 0
+  for(Rtp in tp.axis){
+    tpi <- tpi + 1
+    fpi <- 0
+    for(Rfp in fp.axis){
+      fpi <- fpi + 1
+      E <- ((1/(1 - p - Rfp)) - (1/(1 - p))) + price_ratio * (((p - Rtp)/(1 - p - Rfp)) - (p/(1 - p)))
+      cost[tpi, fpi] <- E
+    }
+  }
+  return(cost)
+}
+
+plot_surface <- function(surface, price_ratio, p){
+  x <- seq(0, (50-0.1), 0.5)
+  y <- seq(0, 99, 1)
+  data <- expand.grid(FPR=x, TPR=y)
+  data$Cost.effect <- as.vector(t(surface))
+  
+  ggplot(data, aes(FPR, TPR, fill = Cost.effect)) + 
+    geom_tile() + 
+    
+    labs(title = "Cost Effect of Classifier in ROC Space",
+         subtitle = paste0("Using a price ratio of ", price_ratio, " and positive prevalence of ", p, ".")) +
+    geom_abline(intercept = 0, slope = 1, color="white") + 
+    theme_minimal() 
+}
+
[…]

### 2020-05-24 src/eba_sims_w_data.Rmd
costfunction into savingsfunction + first functions for storage sustainability analysis

@@
 
-```{r}
+```{r savings effect, echo = FALSE}
+p <- actual.deferral.rate
+response.rate <- 1 - 1/10  # 10 invites -> 1 donor
+invite.price <- 0.5  # Guess
+deferral.price <- 30  # Guesstimate
+savings <- savings_effect(p, response.rate, invite.price, deferral.price)
+
+# Alternative p for comparison
+p2 <- 0.1
+alt.p <- savings_effect(p2, response.rate, invite.price, deferral.price)
+```
+
+```{r plot savings}
+savings.df <- data.frame(savings = savings, alt = alt.p)
+p <- ggplot(data = savings.df, aes(x = seq(0, 100, 1), y = -1*savings, color = "savings")) + 
+  geom_line() +
+  geom_line(aes(x = seq(0, 100, 1), y = -1*alt.p, color = "alt.p")) +
+  theme_minimal() +
+  theme(legend.position = "none") +
+  labs(title = "Savings as a function of True Positive Rate",
+       subtitle = "p = 0.031 and p = 0.1, response.rate = 0.9, invite.price = 0.5, deferral.price = 30",
+       y = "Savings (€)",
+       x = "TPR") +
+  geom_text(aes(x = 100, y = 0.9, label = "p = 0.031", color = "savings")) +
+  geom_text(aes(x = 100, y = 3.5, label = "p = 0.1", color = "alt.p"))
+
+p
+```
+
+```{r deprecated_roc_space, include = FALSE}
+# DEPRECATED!!! WORKS BUT IS INCORRECT. DO NOT USE.
 p <- 0.031
-price_ratio <- 200
+price_ratio <- 13.683
 surface <- cost_surface(p, price_ratio)
@@
 ```
+
+# Storage level simulation
+```{r storage level}
+# Create "pool series" based on the following "selections"
+# We designate 5 different levels for donation intervals to approximate personalized donation intervals
+# This is a "male pool" in that we set the intervals at 30, 40, 60, 70 and 90 days (below and above the current)
+# We guess that those interval categories appear at rates 8, 12, 50, 20 and 10 percent
+# We run the series for 1 year (365 days)
+# We have a pool of 1000 invitees
+# To simplify, it is furthermore assumed that invites are sent with 100% response rate at earliest eligibility
+
+pool.series <- generate_pool_series(5, c(0.08, 0.12, 0.5, 0.20, 0.10), 1000, 365, c(30, 40, 60, 70, 90))
+# TODO: Non-discretize the above (build on distributions)
+
+# Let's simulate a 1000 realities with these parameters and find the worst case scenario
+realities.matrix <- matrix(, nrow = 1000, ncol = 365)
+for(i in 1:1000){
+  pool.series <- generate_pool_series(5, c(0.08, 0.12, 0.5, 0.20, 0.10), 1000, 365, c(30, 40, 60, 70, 90))
+  series <- colSums(pool.series)
+  realities.matrix[i, ] <- series
+}
+
+# Find worst cases
+worsts <- c()
+for(i in 1:1000){
+  worst <- min(realities.matrix[i, ][90:365])
+  worsts <- c(worsts, worst)
+}
+```

### 2020-05-24 src/helper_functions.R

@@ -1,8 +1,38 @@
-generate_population <- function(n, prob){
-  rand <- runif(n)
-  population <- as.integer(rand <= prob)
+generate_pool_series <- function(num.levels, props, pool_size, days, exclusion.levels){
+  if(length(props) != num.levels){
+    stop(paste0("Length of proportions list does not match given number of levels: ", length(props)," != ", num.levels))
+  }
   
-  return(population)
+  levels <- 1:num.levels
+  elig.probs <- 1/exclusion.levels
+  
+  # Initialize population based on levels
+  population <- sample(levels, pool_size, replace = TRUE, prob = props)
+  
+  # Initialize series
+  init.days <- c()
+  for(donor in population){
+    days.until <- sample(0:(exclusion.levels[donor]-1), 1)
+    init.days <- c(init.days, days.until)
+  }
+  
+  # Initialize time series grid
+  pool.series.matrix <- matrix(, nrow = pool_size, ncol = days)
+  
+  # Construct series
+  i <- 0
+  for(donor.time in init.days){
+    i <- i + 1
+    donor.init <- rep(0, donor.time)
+    level.cycle <- c(rep(0, exclusion.levels[population[i]]), c(1))
+    if(days <= length(level.cycle)){
+      donor.series <- c(donor.init, level.cycle)[1:days]
+    } else{
+      donor.series <- c(donor.init, rep(level.cycle, ceiling(days/length(level.cycle))))[1:days]
+    }
+    pool.series.matrix[i, ] <- donor.series
+  }
+  return(pool.series.matrix)
 }
 
 # Function to transfrom dates to year.week 
@@ -14,7 +44,27 @@ yearweek <- function(x = Sys.Date()) {
   return(sprintf("%s.%d", format(nth, "%Y"), 1 + (nth - jan1)%/%ddays(7)))
 }
 
+savings_function <- function(p, q, response.rate, invite.price, deferral.price){
+  # This is for the current implementation: we are only aware of savings
+  cost <- (1/(1 - q - response.rate + q*response.rate) - 
+             1/(1 - p - response.rate + p*response.rate))*invite.price + 
+          (q/(1 - q) - p/(1 - p))*deferral.price
+  return(cost)
+}
+
+savings_effect <- function(p, response.rate, invite.price, deferral.price){
+  
[…]

### 2020-08-16 subgroup_analysis.Rmd
Update repo

---
title: "Subgroup cost analysis"
output: html_notebook
---

```{r setup, echo = FALSE}
setwd("~/interval-sims/src")
library(knitr)
library(dplyr)
library(plyr)
library(lubridate)
library(numbers)
library(data.table)
library(R.utils)
source("pffunctions.R")
source("helper_functions.R")
```

```{r load_data, echo = FALSE}
load("~/eba_sims/data/20191211CollectionAmounts.rdata")
```

# Create subgroups from donation data
```{r subgroups}
# All successful whole blood donations
dons <- filter_donations(data = donation10, donat_phleb = "K", status = "-")

# Women under 30
fU30Hbdeferred <- filter_donations(data = donation10, donat_phleb = "*", gender = "Women", age = c(17, 30), Hb_deferral = 1)
fU30don <- filter_donations(data = donation10, donat_phleb = "K", gender = "Women", status = "-", age = c(17, 30))

# Women under 40
fU40Hbdeferred <- filter_donations(data = donation10, donat_phleb = "*", gender = "Women", age = c(30, 40), Hb_deferral = 1)
fU40don <- filter_donations(data = donation10, donat_phleb = "K", gender = "Women", status = "-", age = c(30, 40))

# Men under 40
mU40Hbdeferred <- filter_donations(data = donation10, donat_phleb = "*", gender = "Men", age = c(17, 40), Hb_deferral = 1)
mU40don <- filter_donations(data = donation10, donat_phleb = "K", gender = "Men", status = "-", age = c(17, 40))
```

We don't need to analyse the entire available history to get a robust average. It is also important that the statistics reflect the current state of the donation population. This is why we cut the sets into windows.

## Windowing
```{r windowing}
# Windowing between 2018-2020
wfU30don <- subset(fU30don, date >= "2018-01-01" & date <= "2019-12-30")
wfU30def <- subset(fU30Hbdeferred, date >= "2018-01-01" & date <= "2019-12-30")

wfU40don <- subset(fU40don, date >= "2018-01-01" & date <= "2019-12-30")
wfU40def <- subset(fU40Hbdeferred, date >= "2018-01-01" & date <= "2019-12-30")

wmU40don <- subset(mU40don, date >= "2018-01-01" & date <= "2019-12-30")
wmU40def <- subset(mU40Hbdeferred, date >= "2018-01-01" & date <= "2019-12-30")
```

# Strategy 1

In the first strategy we limit the maximum donations for women under 30 to 1 per year. We thus iterate through the 2 year window, identify donors and see how many times they've donated and how many times they've been deferred to calculate costs and savings with this strategy. We limit the analysis to first 10,000 events to reduce the analysis time.

```{r tallying1}
vec <- c()
defvec <- c()
for(i in 1:10000){
  # Get donor and date
  donor <- wfU30don$donor[i]
  dondate <- as.Date(head(wfU30don$date[i], 10))

  # Get donor window
  yearw <- subset(wfU30don, date >= dondate & date <= (dondate + years(1)))
  defyearw <- subset(wfU30def, date >= dondate & date <= (dondate + years(1)))

  # Tally up yo
  num <- as.integer(table(yearw$donor)[donor])
  vec <- c(vec, num)
  defnum <- as.integer(table(defyearw$donor)[donor])
  defvec <- c(defvec, defnum)
}
```

We also need to check what percentage of all donations come from this subgroup
```{r total_proportion}
wdons <- subset(dons, date >= "2018-01-01" & date <= "2019-12-30")
p <- nrow(wfU30don)/nrow(wdons)
p
```

## Cost effect

Cost effect is calculated by multiplying the lost flux by the unit marketing cost (representing increasing marketing costs) and then subtracting savings generated by avoided deferrals from it. Deferrals are scaled w.r.t total donation flux (max avoided is around 3 %).

```{r dons_lost1}
# All donations lost if interval extended (%)
vec[is.na(vec)] <- 0
subflux_effect <- sum(vec[vec > 1] - 1)/sum(vec)
dons_lost <- subflux_effect * p
dons_lost
```

```{r deferreals_avoided1}
# Deferrals avoided if interval extended (%)
defvec[is.na(defvec)] <- 0
sub_avoided <- sum(defvec)/sum(vec)
def_avoided <- sub_avoided * p
def_avoided
```

```{r cost_effect1}
E <- 2*dons_lost - 60*def_avoided
E
```

# Strategy 2

Second strategy limits donations from women under 30 years to 1 per year and to 2 per year for women between 30-40. We'll use the result of the previous strategy and add to it results from limiting women between 30-40 to 2 donations a year.

```{r tallying2}
vec <- c()
defvec <- c()
for(i in 1:10000){
  # Get donor and date
  donor <- wfU40don$donor[i]
  dondate <- as.Date(head(wfU40don$date[i], 10))

  # Get donor window
  yearw <- subset(wfU40don, date >= dondate & date <= (dondate + months(6)))
  defyearw <- subset(wfU40def, date >= dondate & date <= (dondate + months(6)))

  # Tally up yo
  num <- as.integer(table(yearw$donor)[donor])
  vec <- c(vec, num)
  defnum <- as.integer(table(defyearw$donor)[donor])
  defvec <- c(defvec, defnum)
}
```

We also need to check what percentage of all donations come from this subgroup
```{r total_proportion2}
p <- nrow(wfU40don)/nrow(wdons)
p
```

## Cost effect

```{r dons_lost2}
# All donations lost if interval extended (%)
vec[is.na(vec)] <- 0
subflux_effect <- sum(vec[vec > 1] - 1)/sum(vec)
dons_lost <- subflux_effect * p
dons_lost
```

```{r deferreals_avoided2}
# Deferrals avoided if interval extended (%)
defvec[is.na(defvec)] <- 0
sub_avoided <- sum(defvec)/sum(vec)
def_avoided <- sub_avoided * p
def_avoided
```

```{r cost_effect2}
E2 <- 2*dons_lost - 60*def_avoided
E12 <- E + E2
E12
```

# Strategy 3

Under the 3rd strategy, women under 30 may also donate twice per year. We recalulate and sum with E2.

```{r tallying3}
vec <- c()
defvec <- c()
for(i in 1:10000){
  # Get donor and date
  donor <- wfU30don$donor[i]
  dondate <- as.Date(head(wfU30don$date[i], 10))

  # Get donor window
  yearw <- subset(wfU30don, date >= dondate & date <= (dondate + months(6)))
  defyearw <- subset(wfU30def, date >= dondate & date <= (dondate + months(6)))

  # Tally up yo
  num <- as.integer(table(yearw$donor)[donor])
  vec <- c(vec, num)
  defnum <- as.integer(table(defyearw$donor)[donor])
  defvec <- c(defvec, defnum)
}

# Reinstance p
p <- nrow(wfU30don)/nrow(wdons)
p
```

## Cost effect

```{r dons_lost3}
# All donations lost if interval extended (%)
vec[is.na(vec)] <- 0
subflux_effect <- sum(vec[vec > 1] - 1)/sum(vec)
dons_lost <- subflux_effect * p
dons_lost
```

```{r deferreals_avoided3}
# Deferrals avoided if interval extended (%)
defvec[is.na(defvec)] <- 0
sub_avoided <- sum(defvec)/sum(vec)
def_avoided <- sub_avoided * p
def_avoided
```

```{r cost_effect3}
E3 <- 2*dons_lost - 60*def_avoided
E23 <- E2 + E3
E23
```

# Strategy 4

4 donations for men under 40.

```{r tallying4}
vec <- c()
defvec <- c()
for(i in 1:10000){
  # Get donor and date
  donor <- wmU40don$donor[i]
  dondate <- as.Date(head(wmU40don$date[i], 10))

  # Get donor window
  yearw <- subset(wmU40don, date >= dondate & date <= (dondate + months(3)))
  defyearw <- subset(wmU40def, date >= dondate & date <= (dondate + months(3)))

  # Tally up yo
  num <- as.integer(table(yearw$donor)[donor])
  vec <- c(vec, num)
  defnum <- as.integer(table(defyearw$donor)[donor])
  defvec <- c(defvec, defnum)
}

# Reinstance p
p <- nrow(wmU40don)/nrow(wdons)
p
```

## Cost effect

```{r dons_lost4}
# All donations lost if interval extended (%)
vec[is.na(vec)] <- 0
subflux_effect <- sum(vec[vec > 1] - 1)/sum(vec)
dons_lost <- subflux_effect * p
dons_lost
```

```{r deferreals_avoided4}
# Deferrals avoided if interval extended (%)
defvec[is.na(defvec)] <- 0
sub_avoided <- sum(defvec)/sum(vec)
def_avoided <- sub_avoided * p
def_avoided
```

```{r cost_effect4}
E4 <- 2*dons_lost - 60*def_avoided
E4
```

### 2020-08-16 helper_functions.R

@@ -1,0 +1,121 @@
+generate_pool_series <- function(num.levels, props, pool_size, days, exclusion.levels){
+  if(length(props) != num.levels){
+    stop(paste0("Length of proportions list does not match given number of levels: ", length(props)," != ", num.levels))
+  }
+  
+  levels <- 1:num.levels
+  elig.probs <- 1/exclusion.levels
+  
+  # Initialize population based on levels
+  population <- sample(levels, pool_size, replace = TRUE, prob = props)
+  
+  # Initialize series
+  init.days <- c()
+  for(donor in population){
+    days.until <- sample(0:(exclusion.levels[donor]-1), 1)
+    init.days <- c(init.days, days.until)
+  }
+  
+  # Initialize time series grid
+  pool.series.matrix <- matrix(, nrow = pool_size, ncol = days)
+  
+  # Construct series
+  i <- 0
+  for(donor.time in init.days){
+    i <- i + 1
+    donor.init <- rep(0, donor.time)
+    level.cycle <- c(rep(0, exclusion.levels[population[i]]), c(1))
+    if(days <= length(level.cycle)){
+      donor.series <- c(donor.init, level.cycle)[1:days]
+    } else{
+      donor.series <- c(donor.init, rep(level.cycle, ceiling(days/length(level.cycle))))[1:days]
+    }
+    pool.series.matrix[i, ] <- donor.series
+  }
+  return(pool.series.matrix)
+}
+
+# Function to transfrom dates to year.week 
+yearweek <- function(x = Sys.Date()) {
+  xday <- ISOdate(year(x), month(x), day(x), tz=tz(x))
+  dn <- 1 + (wday(x) + 5)%%7
+  nth <- xday + ddays(4 - dn)
+  jan1 <- ISOdate(year(nth), 1, 1, tz=tz(x))
+  return(sprintf("%s.%d", format(nth, "%Y"), 1 + (nth - jan1)%/%ddays(7)))
+}
+
+savings_function <- function(p, q, response.rate, invite.price, deferral.price){
+  # This is for the current implementation: we are only aware of savings
+  cost <- (1/(1 - q - response.rate + q*response.rate) - 
+             1/(1 - p - response.rate + p*response.rate))*invite.price + 
+          (q/(1 - q) - p/(1 - p))*deferral.price
+  return(cost)
+}
+
[…]

### 2020-08-16 pffunctions.R

@@ -1,0 +1,375 @@
+# Helper functions
+
+extract_type <- function(red.distr, type){
+  # Create a full sequence of dates for imputation purposes
+  all.dates <- (seq.Date(min(red.distr$date),
+                         max(red.distr$date),
+                         "day"))
+  # Find type
+  typed <- red.distr[red.distr$ABO == type, ]
+  typed.distr <- aggregate(typed$quantity, by = list(typed$date), sum); colnames(typed.distr) <- c("date", "pcs")
+  # Merge into a whole set with NAs
+  typed.distr <- merge(x = data.frame(date = all.dates),
+                       y = typed.distr,
+                       all.x = TRUE)
+  # Replace with zeroes
+  typed.distr[is.na(typed.distr)] <- 0
+  # Cut to time after 2014
+  typed.distr <- typed.distr[typed.distr$date >= as.Date("2014-01-06"), ]
+  
+  return(typed.distr)
+}
+
+aggregate_weekly <- function(series){
+  pcslist <- list(); weeklist <- list(); datelist <- list()  # Create storage lists
+  j <- 0; k <- 0  # j is for resetting week counter, k is for gathering dates
+  for(i in seq(to = length(series$date), by = 7)){
+    if(j == 52){j <- 0}
+    j <- j + 1
+    k <- k + 1
+    pcslist[[k]] <- sum(series$pcs[i : (i + 6)])
+    weeklist[[k]] <- j
+    datelist[[k]] <- series$date[i]
+  }
+  weekly <- data.frame(week = unlist(weeklist), startdate = as.Date.numeric(unlist(datelist), origin = "1970-01-01"), pcs = unlist(pcslist))
+  return(weekly)
+}
+
+find_errors <- function(segment, beginning, series.ts, method = "none", smooth = "none", freq = "monthly"){
+  apes <- c()
+  if(freq == "monthly"){
+    for(i in seq(length(series.ts) - 37)){
+      # Define training and testing set as ROLLING WINDOW
+      if(smooth == ".25"){
+        train <- ts(itsmr::smooth.fft(series.ts[(1 + i):(36 + i)], .25), 
+                    start = decimal_date(beginning + i), 
+                    frequency = 12)
+        test <- ts(series.ts[(37 + i)], start = decimal_date(beginning + 37 + i), frequency = 12)
+      }
+      if(smooth == ".10"){
[…]

### 2020-08-16 helper_functions.R
Delete helper_functions.R

@@ -1,121 +1,0 @@
-generate_pool_series <- function(num.levels, props, pool_size, days, exclusion.levels){
-  if(length(props) != num.levels){
-    stop(paste0("Length of proportions list does not match given number of levels: ", length(props)," != ", num.levels))
-  }
-  
-  levels <- 1:num.levels
-  elig.probs <- 1/exclusion.levels
-  
-  # Initialize population based on levels
-  population <- sample(levels, pool_size, replace = TRUE, prob = props)
-  
-  # Initialize series
-  init.days <- c()
-  for(donor in population){
-    days.until <- sample(0:(exclusion.levels[donor]-1), 1)
-    init.days <- c(init.days, days.until)
-  }
-  
-  # Initialize time series grid
-  pool.series.matrix <- matrix(, nrow = pool_size, ncol = days)
-  
-  # Construct series
-  i <- 0
-  for(donor.time in init.days){
-    i <- i + 1
-    donor.init <- rep(0, donor.time)
-    level.cycle <- c(rep(0, exclusion.levels[population[i]]), c(1))
-    if(days <= length(level.cycle)){
-      donor.series <- c(donor.init, level.cycle)[1:days]
-    } else{
-      donor.series <- c(donor.init, rep(level.cycle, ceiling(days/length(level.cycle))))[1:days]
-    }
-    pool.series.matrix[i, ] <- donor.series
-  }
-  return(pool.series.matrix)
-}
-
-# Function to transfrom dates to year.week 
-yearweek <- function(x = Sys.Date()) {
-  xday <- ISOdate(year(x), month(x), day(x), tz=tz(x))
-  dn <- 1 + (wday(x) + 5)%%7
-  nth <- xday + ddays(4 - dn)
-  jan1 <- ISOdate(year(nth), 1, 1, tz=tz(x))
-  return(sprintf("%s.%d", format(nth, "%Y"), 1 + (nth - jan1)%/%ddays(7)))
-}
-
-savings_function <- function(p, q, response.rate, invite.price, deferral.price){
-  # This is for the current implementation: we are only aware of savings
-  cost <- (1/(1 - q - response.rate + q*response.rate) - 
-             1/(1 - p - response.rate + p*response.rate))*invite.price + 
-          (q/(1 - q) - p/(1 - p))*deferral.price
-  return(cost)
-}
-
[…]

### 2020-08-16 pffunctions.R
Delete pffunctions.R

@@ -1,375 +1,0 @@
-# Helper functions
-
-extract_type <- function(red.distr, type){
-  # Create a full sequence of dates for imputation purposes
-  all.dates <- (seq.Date(min(red.distr$date),
-                         max(red.distr$date),
-                         "day"))
-  # Find type
-  typed <- red.distr[red.distr$ABO == type, ]
-  typed.distr <- aggregate(typed$quantity, by = list(typed$date), sum); colnames(typed.distr) <- c("date", "pcs")
-  # Merge into a whole set with NAs
-  typed.distr <- merge(x = data.frame(date = all.dates),
-                       y = typed.distr,
-                       all.x = TRUE)
-  # Replace with zeroes
-  typed.distr[is.na(typed.distr)] <- 0
-  # Cut to time after 2014
-  typed.distr <- typed.distr[typed.distr$date >= as.Date("2014-01-06"), ]
-  
-  return(typed.distr)
-}
-
-aggregate_weekly <- function(series){
-  pcslist <- list(); weeklist <- list(); datelist <- list()  # Create storage lists
-  j <- 0; k <- 0  # j is for resetting week counter, k is for gathering dates
-  for(i in seq(to = length(series$date), by = 7)){
-    if(j == 52){j <- 0}
-    j <- j + 1
-    k <- k + 1
-    pcslist[[k]] <- sum(series$pcs[i : (i + 6)])
-    weeklist[[k]] <- j
-    datelist[[k]] <- series$date[i]
-  }
-  weekly <- data.frame(week = unlist(weeklist), startdate = as.Date.numeric(unlist(datelist), origin = "1970-01-01"), pcs = unlist(pcslist))
-  return(weekly)
-}
-
-find_errors <- function(segment, beginning, series.ts, method = "none", smooth = "none", freq = "monthly"){
-  apes <- c()
-  if(freq == "monthly"){
-    for(i in seq(length(series.ts) - 37)){
-      # Define training and testing set as ROLLING WINDOW
-      if(smooth == ".25"){
-        train <- ts(itsmr::smooth.fft(series.ts[(1 + i):(36 + i)], .25), 
-                    start = decimal_date(beginning + i), 
-                    frequency = 12)
-        test <- ts(series.ts[(37 + i)], start = decimal_date(beginning + 37 + i), frequency = 12)
-      }
-      if(smooth == ".10"){
[…]

### 2020-08-16 src/helper_functions.R
Update repo

@@ -63,50 +63,59 @@ savings_effect <- function(p, response.rate, invite.price, deferral.price){
   return(E)
 }
 
-cost_surface <- function(p, price_ratio){
-  # DEPRECATED!!! WORKS BUT IS NOT CORRECT. DO NOT USE.
-  
-  # Cut FPR at 100x%
-  x = 0.5
-  
-  # Set axis 
-  tp.axis <- seq(0, (p-p/100), p/100) # TPs go from 0 to p
-  fp.axis <- seq(0, (1-p)*x - ((1-p)*x)/100, ((1-p)*x)/100) # FPs need to be limited: they explode the cost (due to reinvites) approaching 1-p
-  
-  # Set grid
-  cost <- matrix(, nrow = length(tp.axis), ncol = length(fp.axis))
-  
-  tpi <- 0
-  for(Rtp in tp.axis){
-    tpi <- tpi + 1
-    fpi <- 0
-    for(Rfp in fp.axis){
-      fpi <- fpi + 1
-      E <- ((1/(1 - p - Rfp)) - (1/(1 - p))) + price_ratio * (((p - Rtp)/(1 - p - Rfp)) - (p/(1 - p)))
-      cost[tpi, fpi] <- E
-    }
+
+filter_donations <- function(data, 
+                             donor=NULL, donation13=NULL, site=NULL,
+                             date=NULL, phleb_start=NULL, status=NULL,
+                             donat_phleb=NULL, Hb=NULL, gender=NULL,
+                             dob=NULL, aborh=NULL, age=NULL,
+                             agegroup=NULL, Hb_deferral=NULL){
+  temp <- data
+  if(!is.null(donor)){
+    temp <- temp[temp$donor==donor,]
   }
-  return(cost)
-}
-
-plot_surface <- function(surface, price_ratio, p){
-  # DEPRECATED!!! SEE cost_surface()!!
-  x <- seq(0, (50-0.1), 0.5)
-  y <- seq(0, 99, 1)
-  data <- expand.grid(FPR=x, TPR=y)
-  data$Cost.effect <- as.vector(t(surface))
+  if(!is.null(donation13)){
+    temp <- temp[temp$donation13==donation13,]
+  }
+  if(!is.null(site)){
+    temp <- temp[temp$site==site,]
+  }
+  if(!is.null(date)){
+    temp <- temp[temp$date > date[1] & temp$date < date[1],]
+  }
[…]

### 2020-08-23 src/subgroup_analysis.Rmd
Add files via upload

--- 2020-08-16 subgroup_analysis.Rmd
@@
 title: "Subgroup cost analysis"
-output: html_notebook
+output:
+  html_document:
+    df_print: paged
 ---
 
-```{r setup, echo = FALSE}
+```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
 setwd("~/interval-sims/src")

### 2020-08-23 src/eba_sims_w_data.Rmd
Add files via upload

--- 2020-05-06 src/eba_sims_w_data.Rmd
@@
 
-```{r filtering_function, echo = FALSE}
-# Let's build a tool to filter quickly as we please
-# TODO: Let the user select multiple levels for each factor
-filter_donations <- function(data, 
-                             donor=NULL, donation13=NULL, site=NULL,
-                             date=NULL, phleb_start=NULL, status=NULL,
-                             donat_phleb=NULL, Hb=NULL, gender=NULL,
-                             dob=NULL, aborh=NULL, age=NULL,
-                             agegroup=NULL, Hb_deferral=NULL){
-  temp <- data
-  if(!is.null(donor)){
-    temp <- temp[temp$donor==donor,]
-  }
-  if(!is.null(donation13)){
-    temp <- temp[temp$donation13==donation13,]
-  }
-  if(!is.null(site)){
-    temp <- temp[temp$site==site,]
-  }
-  if(!is.null(date)){
-    temp <- temp[temp$date > date[1] & temp$date < date[1],]
-  }
-  if(!is.null(phleb_start)){
-    temp <- temp[temp$phleb_start==phleb_start,]
-  }
-  if(!is.null(status)){
-    temp <- temp[temp$status==status,]
-  }
-  if(!is.null(donat_phleb)){
-    temp <- temp[temp$donat_phleb==donat_phleb,]
-  }
-  if(!is.null(Hb)){
-    temp <- temp[temp$Hb > Hb[1] & temp$Hb < Hb[2],]
-  }
-  if(!is.null(gender)){
-    temp <- temp[temp$gender==gender,]
-  }
-  if(!is.null(site)){
-    temp <- temp[temp$site==site,]
-  }
-  if(!is.null(dob)){
-    temp <- temp[temp$dob > dob[1] & temp$dob < dob[2],]
-  }
-  if(!is.null(aborh)){
-    temp <- temp[temp$aborh==aborh,]
-  }
-  if(!is.null(age)){
-    temp <- temp[temp$age > age[1] & temp$age < age[2],]
-  }
-  if(!is.null(agegroup)){
-    temp <- temp[temp$agegroup==agegroup,]
-  }
-  if(!is.null(Hb_deferral)){
-    temp <- temp[temp$Hb_deferral==Hb_deferral,]
-  }
-
-  return(temp)
-}
-```
-
 ```{r filter_test}
@@
 dons <- filter_donations(data = donation10, donat_phleb = "K", status = "-")
-
 ```
@@
 Sanity check: passed. (60 is the maximum number of donations in 10 years time)
-
-```{r}
-head(dons)
-```
 
@@
 ```{r}
-# Check
-firsts[1:20,]
-```
-
-```{r}
-# Attempting vectorization of the processes below
-
 # Firsts
@@
 
-## Cost analysis
+```{r}
+library(truncnorm)
 
-```{r}
-actual.deferral.rate <- 76565/2464190
+q <- seq(0, 0.031, 0.031/1000)
+x <- seq(1, 1.5, 1/10000)
+adjustments <- sort(round(rtruncnorm(n=length(q), a=1, b=1.5, mean=1.1, sd=0.1), digits = 4))
+E.matrix <- matrix(, nrow = 1001, ncol = 1001)
+Pm <- 2
+Pd <- 60 # from Satu
+rl <- 0
+Fn = 0.259 # Vuosikertomus 2019
+qalt <- seq(0, 0.06, 0.06/1000) # Larger q
+q <- seq(0, 0.031, 0.031/1000) # from data
+for(i in 1:1001){
+  Fratio <- adjustments[i] / (1 + Fn * (adjustments[i] - 1))
+  for(j in 1:1001){
+    Em <- Pm * (Fratio - 1 - q[j] * rl)
+    Ed <- Pd * q[j]
+    E.matrix[i, j] <- Em - Ed
+    }
+}
+
+E.long <- melt(E.matrix)
+
 ```
@@
 ```{r}
-p <- 0.031
-price_ratio <- 200
-surface <- cost_surface(p, price_ratio)
-plot_surface(surface, price_ratio, p)
-plot_interactive_surface(surface)
+colnames(E.long) <- c("adjustment", "q", "cost")
+ggplot(E.long, aes(x = q, y = adjustment, z = cost)) +
+geom_tile(aes(fill=cost)) +
+  stat_contour(color = "white", size = 0.7, bins = 2) +
+  scale_fill_gradientn(colours=c("blue", "red")) +
+labs(x="q%", y="Adjustment", title="Cost surface", subtitle = paste0("Surface minimum: ", round(min(E.long$cost), 2), ". Cost effect thresholds highlighted at 0 and -1 euros.")) +
+theme_bw() +
+scale_x_continuous(breaks = c(1, 1001), labels = c("0", "3.1")) +
+scale_y_continuous(breaks = c(1, 1001), labels = c(as.character(min(adjustments)), as.character(max(adjustments))))
 ```

### 2020-09-21 src/current_model_effect.Rmd
Update to most recent

---
title: "Current model cost effect"
author: "Esa Turkulainen"
date: "8/25/2020"
output: html_document
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
setwd("~/interval-sims/src")
library(knitr)
library(dplyr)
library(plyr)
library(lubridate)
library(numbers)
library(data.table)
library(R.utils)
source("pffunctions.R")
source("helper_functions.R")
```

```{r load_data, echo = FALSE}
load("~/eba_sims/data/20191211CollectionAmounts.rdata")
```

# Create subgroups from donation data
```{r subgroups}
# All successful whole blood donations
dons <- filter_donations(data = donation10, donat_phleb = "K", status = "-")
# Order donations by date
dons <- dons[with(dons, order(date)), ]
# Find only returning donors
rdons <- dons[duplicated(dons$donor),]

# Successful male
mdon <- filter_donations(data = rdons, donat_phleb = "K", status = "-", gender = "Men")
# Successful female
fdon <- filter_donations(data = rdons, donat_phleb = "K", status = "-", gender = "Women")
# Deferrals
def <- filter_donations(data = donation10, donat_phleb = "*", Hb_deferral = 1)
mdef <- filter_donations(data = def, gender = "Men")
fdef <- filter_donations(data = def, gender = "Women")

# Windows
walldons <- subset(dons, date >= "2018-01-01" & date <= "2019-12-30")
wdons <- subset(rdons, date >= "2018-01-01" & date <= "2019-12-30")
wmdon <- subset(mdon, date >= "2018-01-01" & date <= "2019-12-30")
wfdon <- subset(fdon, date >= "2018-01-01" & date <= "2019-12-30")
wdef <- subset(def, date >= "2018-01-01" & date <= "2019-12-30")
wmdef <- subset(mdef, date >= "2018-01-01" & date <= "2019-12-30")
wfdef <- subset(fdef, date >= "2018-01-01" & date <= "2019-12-30")
```

# Find ratios
```{r}
# Male ratio
mr <- nrow(wmdon)/nrow(wdons)
# Female ratio
fr <- nrow(wfdon)/nrow(wdons)
# Male deferral ratio
mdr <- nrow(wmdef) / nrow(wdef)
# Female deferral ratio
fdr <- nrow(wfdef) / nrow(wdef)
# Total deferral prevalence
d <- nrow(wdef)/(nrow(wdef) + nrow(walldons))
```

The current model finds 1653/2116 = 78.1% of deferrals. In addition, it misclassifies 22.57% of donations as deferrals.
```{r}
TPR <- 0.781
FPR <- 0.2257
# Find q
q <- d * TPR
# Find a for 6 month category
# correctly predicted non-deferrals: 77.43% of the returning donations (all - d): (0.7743 * (1 - d))
# incorrectly predicted non-deferrals: 22.57% of the returning donations (all - d): (0.2257 * (1 - d))
a6 <- 1 * ((1 - FPR) * (1 - d)) + 
  3 * mr * (FPR * (1 - d)) + 2 * fr * (FPR * (1 - d)) + 
  3 * mdr * d + 2 * fdr * d

a12 <- 1 * ((1 - FPR) * (1 - d)) + 
  6 * mr * (FPR * (1 - d)) + 4 * fr * (FPR * (1 - d)) + 
  6 * mdr * d + 4 * fdr * d
```

# Final cost
```{r}
Pm <- 2
Pd <- 60 # from Satu
rl <- 0
Fn = 0.1066 # Data

# a6
Fratio <- a6 / (1 + Fn * (a6 - 1))
Em <- Pm * (Fratio - 1 - q * rl)
Ed <- Pd * q
E6 <- Em - Ed

# a12
Fratio <- a12 / (1 + Fn * (a12 - 1))
Em <- Pm * (Fratio - 1 - q * rl)
Ed <- Pd * q
E12 <- Em - Ed
```

```{r}
E6
E12
```

### 2020-09-21 src/eba_sims_w_data.Rmd
Update to most recent

@@
 
-q <- seq(0, 0.031, 0.031/1000)
-x <- seq(1, 1.5, 1/10000)
-adjustments <- sort(round(rtruncnorm(n=length(q), a=1, b=1.5, mean=1.1, sd=0.1), digits = 4))
+q <- seq(0, 0.032, 0.032/1000)
+adjustments <- seq(1, 2, 1/1000)
 E.matrix <- matrix(, nrow = 1001, ncol = 1001)
@@
 rl <- 0
-Fn = 0.259 # Vuosikertomus 2019
-qalt <- seq(0, 0.06, 0.06/1000) # Larger q
-q <- seq(0, 0.031, 0.031/1000) # from data
+Fn = 0.1066 # Data
 for(i in 1:1001){
@@
   scale_fill_gradientn(colours=c("blue", "red")) +
-labs(x="q%", y="Adjustment", title="Cost surface", subtitle = paste0("Surface minimum: ", round(min(E.long$cost), 2), ". Cost effect thresholds highlighted at 0 and -1 euros.")) +
+labs(x="q%", y="Adjustment", title="Cost surface (2018-2020)", subtitle = paste0("Surface minimum: ", round(min(E.long$cost), 2), ".
+Cost effect thresholds highlighted at 0 and -1 euros.", "\n", "Dots mark current model performance with a 6 or 12 month extension strategy.")) +
+  annotate("point", x = 800, y = which(adjustments > 1.3619 & adjustments < (1.3619 + 0.001)), colour = "white", size = 2, alpha = 1) +
+  annotate("point", x = 800, y = which(adjustments > 1.9749 & adjustments < (1.9749 + 0.001)), colour = "white", size = 2, alpha = 1) +
+  annotate("text", x = 760, y = which(adjustments > 1.3619 & adjustments < (1.3619 + 0.001)), label = "6", colour = "white") +
+  annotate("text", x = 760, y = which(adjustments > 1.9749 & adjustments < (1.9749 + 0.001)), label = "12", colour = "white") +
 theme_bw() +
-scale_x_continuous(breaks = c(1, 1001), labels = c("0", "3.1")) +
+scale_x_continuous(breaks = c(1, 1001), labels = c("0", "3.2")) +
 scale_y_continuous(breaks = c(1, 1001), labels = c(as.character(min(adjustments)), as.character(max(adjustments))))

### 2020-09-21 src/subgroup_analysis.Rmd
Update to most recent

@@
 
+# Women 18-25
+f25Hbdeferred <- filter_donations(data = donation10, donat_phleb = "*", gender = "Women", age = c(17, 26), Hb_deferral = 1)
+f25don <- filter_donations(data = donation10, donat_phleb = "K", gender = "Women", status = "-", age = c(17, 26))
+
 # Women under 30
@@
 # Windowing between 2018-2020
+wf25don <- subset(f25don, date >= "2018-01-01" & date <= "2019-12-30")
+wf25def <- subset(f25Hbdeferred, date >= "2018-01-01" & date <= "2019-12-30")
+
 wfU30don <- subset(fU30don, date >= "2018-01-01" & date <= "2019-12-30")
@@
 ```
+
+# Extra strategy
+```{r tallyingExtra}
+vec <- c()
+defvec <- c()
+for(i in 1:10000){
+  # Get donor and date
+  donor <- wf25don$donor[i]
+  dondate <- as.Date(head(wf25don$date[i], 10))
+
+  # Get donor window
+  yearw <- subset(wf25don, date >= dondate & date <= (dondate + years(1)))
+  defyearw <- subset(wf25def, date >= dondate & date <= (dondate + years(1)))
+
+  # Tally up yo
+  num <- as.integer(table(yearw$donor)[donor])
+  vec <- c(vec, num)
+  defnum <- as.integer(table(defyearw$donor)[donor])
+  defvec <- c(defvec, defnum)
+}
+
+# Reinstance p
+p <- nrow(wf25don)/nrow(wdons)
+p
+```
+
+## Cost effect
+
+Cost effect is calculated by multiplying the lost flux by the unit marketing cost (representing increasing marketing costs) and then subtracting savings generated by avoided deferrals from it.
+Deferrals are scaled w.r.t total donation flux (max avoided is around 3 %).
+
+```{r dons_lostExtra}
+# All donations lost if interval extended (%)
+vec[is.na(vec)] <- 0
+subflux_effect <- sum(vec[vec > 1] - 1)/sum(vec)
+dons_lost <- subflux_effect * p
+dons_lost
+```
+
+```{r deferreals_avoidedExtra}
+# Deferrals avoided if interval extended (%)
+defvec[is.na(defvec)] <- 0
+sub_avoided <- sum(defvec)/sum(vec)
+def_avoided <- sub_avoided * p
+def_avoided
+```
+
+```{r cost_effectExtra}
+E <- 2*dons_lost - 60*def_avoided
+E
+```

### 2020-09-21 src/pffunctions.R

@@ -1,33 +1,43 @@
 # Helper functions
 
-extract_type <- function(red.distr, type){
+extract_type <- function(red.distr, type) {
   # Create a full sequence of dates for imputation purposes
-  all.dates <- (seq.Date(min(red.distr$date),
-                         max(red.distr$date),
-                         "day"))
+  all.dates <- (seq.Date(
+    min(red.distr$date),
+    max(red.distr$date),
+    "day"
+  ))
   # Find type
   typed <- red.distr[red.distr$ABO == type, ]
-  typed.distr <- aggregate(typed$quantity, by = list(typed$date), sum); colnames(typed.distr) <- c("date", "pcs")
+  typed.distr <- aggregate(typed$quantity, by = list(typed$date), sum)
+  colnames(typed.distr) <- c("date", "pcs")
   # Merge into a whole set with NAs
-  typed.distr <- merge(x = data.frame(date = all.dates),
-                       y = typed.distr,
-                       all.x = TRUE)
+  typed.distr <- merge(
+    x = data.frame(date = all.dates),
+    y = typed.distr,
+    all.x = TRUE
+  )
   # Replace with zeroes
   typed.distr[is.na(typed.distr)] <- 0
   # Cut to time after 2014
   typed.distr <- typed.distr[typed.distr$date >= as.Date("2014-01-06"), ]
-  
+
   return(typed.distr)
 }
 
-aggregate_weekly <- function(series){
-  pcslist <- list(); weeklist <- list(); datelist <- list()  # Create storage lists
-  j <- 0; k <- 0  # j is for resetting week counter, k is for gathering dates
-  for(i in seq(to = length(series$date), by = 7)){
-    if(j == 52){j <- 0}
+aggregate_weekly <- function(series) {
+  pcslist <- list()
+  weeklist <- list()
+  datelist <- list() # Create storage lists
+  j <- 0
+  k <- 0 # j is for resetting week counter, k is for gathering dates
+  for (i in seq(to = length(series$date), by = 7)) {
+    if (j == 52) {
+      j <- 0
+    }
     j <- j + 1
     k <- k + 1
-    pcslist[[k]] <- sum(series$pcs[i : (i + 6)])
+    pcslist[[k]] <- sum(series$pcs[i:(i + 6)])
     weeklist[[k]] <- j
     datelist[[k]] <- series$date[i]
   }
@@ -35,176 +45,188 @@ aggregate_weekly <- function(series){
   return(weekly)
 }
 
[…]

### 2020-09-21 src/current_model_effect.Rmd
cleaning and testing variable probability cut-off

@@
 date: "8/25/2020"
-output: html_document
+output:
+  pdf_document: default
+  html_document: default
 ---
@@
 ```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
-setwd("~/interval-sims/src")
-library(knitr)
-library(dplyr)
-library(plyr)
-library(lubridate)
-library(numbers)
-library(data.table)
-library(R.utils)
-source("pffunctions.R")
+#setwd("~/interval-sims/src")
+library(tidyverse)
+library(truncnorm)
+library(caret)
 source("helper_functions.R")
@@
 ```{r load_data, echo = FALSE}
-load("~/eba_sims/data/20191211CollectionAmounts.rdata")
+load("..//data/20191211CollectionAmounts.rdata")
+```
+
+The data object donation10 contains donation history data from beginning of 2010 to end of 2019 it has following columns.
+
+```{r }
+lapply(donation10,class)
 ```
@@
 
-The current model finds 1653/2116 = 78.1% of deferrals. In addition, it misclassifies 22.57% of donations as deferrals.
+The current model finds 1662/2116 = 78.5% of deferrals. In addition, it misclassifies 21.45% of donations as deferrals.
 ```{r}
-TPR <- 0.781
-FPR <- 0.2257
+TPR <- 1662/2116
+FPR <- 1- (1662 /2116)
 # Find q
@@
   6 * mdr * d + 4 * fdr * d
+TPR
+FPR
+a6
+a12
 ```
@@
 ```{r}
-Pm <- 2
-Pd <- 60 # from Satu
-rl <- 0
-Fn = 0.1066 # Data
+Pm <- 2 # estimate collected from FRCBS experts
+Pd <- 60 # estimate collected from FRCBS experts
+rl <- 0 # estimated previously from data
+Fn = 0.1066 # etimated from data in https://github.com/FRCBS/interval-sims/blob/master/src/eba_sims_w_data.Rmd
 
@@
 E12
+q
 ```
+
+# Plot final cost on data surface
+
+```{r}
+
+q.seq <- seq(0, 0.032, 0.032/1000)
+adjustments <- seq(1, 2, 1/1000)
+E.matrix <- matrix(nrow = 1001, ncol = 1001)
+Fn = 0.1066 # 
+for(i in 1:1001){
+  Fratio <- adjustments[i] / (1 + Fn * (adjustments[i] - 1))
+  for(j in 1:1001){
+    Em <- Pm * (Fratio - 1 - q.seq[j] * rl)
+    Ed <- Pd * q.seq[j]
+    E.matrix[i, j] <- Em - Ed
+    }
+}
+
+E.long <- reshape2::melt(E.matrix)
+
+```
+
+```{r}
+colnames(E.long) <- c("adjustment", "q", "cost")
+ggplot(E.long, aes(x = q, y = adjustment, z = cost)) +
+geom_tile(aes(fill=cost)) +
+  stat_contour(color = "white", size = 0.7, bins = 2) +
+  scale_fill_gradientn(colours=c("blue", "red")) +
+labs(x="q%", y="Adjustment", title="Cost surface (2018-2020)", subtitle = paste0("Surface minimum: ", round(min(E.long$cost), 2), ".
+Cost effect thresholds highlighted at 0 and -1 euros.", "\n", "Dots mark current model performance with a 6 or 12 month extension strategy.")) +
+  annotate("point", x = head(which(q.seq > q ),1), y = which(adjustments > a6 & adjustments < (a6 + 0.001)), colour = "white", size = 2, alpha = 1) +
+  annotate("point", x = head(which(q.seq > q ),1), y = which(adjustments > a12 & adjustments < (a12 + 0.001)), colour = "white", size = 2, alpha = 1) +
+  annotate("text", x = head(which(q.seq > q ),1) - 40 , y = which(adjustments > a6 & adjustments < (a6 + 0.001)), label = "6", colour = "white") +
+  annotate("text", x = head(which(q.seq > q ),1) -40, y = which(adjustments > a12 & adjustments < (a12 + 0.001)), label = "12", colour = "white") +
+theme_bw() +
+scale_x_continuous(breaks = c(1, 1001), labels = c("0", "3.2")) +
+scale_y_continuous(breaks = c(1, 1001), labels = c(as.character(min(adjustments)), as.character(max(adjustments))))
+```
+
+#What if we use some other than the default cut-off for model class assingment
+
+```{r}
+#Load a data frame with observed classes and probability of deferral from  a classifier
+load("../data//rrfFit_roc_validate_probs.rdata")
+summary(df)
+```
+
+```{r}
+#cut prediction.probs with seq(0, 1, .1)
+x <- seq(0.1, .9, .1)
+pred.classes <- data.frame(matrix(nrow=nrow(df),ncol=length(x)))
+colnames(pred.classes) <- paste0("p_",x)
+for (i in x ) {
+  pred.classes[,paste0("p_",i)] <-  factor( df$Deferred <= i ,labels = c("Deferred","Accepted"))
+}
+
+head(pred.classes)
+```
+
+```{r}
+#count TPR and FPR for each probability
+s <- vector()
+#df$obs has factor levels in wrong
+df$obs <- relevel(df$obs,"Deferred")
+
+for(i in names(pred.classes)) {
+  conf <- caret::confusionMatrix(reference=df$obs, data=pred.classes[,i], positive = "Deferred",mode="everything")
+  #Sensitivity  = TPR
+  s <- c(s,conf$byClass['Sensitivity']) 
+}
+tprs <- tibble(probability = x,TPR=s,FPR=1-s)
+tprs
+```
+
+```{r}
+
+get_cost <- function(TPR, FPR,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) {
+  q <- d * TPR
+  # Find a for 6 month category
+  # correctly predicted non-deferrals: 77.43% of the returning donations (all - d): (0.7743 * (1 - d))
+  # incorrectly predicted non-deferrals: 22.57% of the returning donations (all - d): (0.2257 * (1 - d))
+  a6 <- 1 * ((1 - FPR) * (1 - d.)) + 
+    3 * mr. * (FPR * (1 - d.)) + 2 * fr. * (FPR * (1 - d.)) + 
+    3 * mdr. * d. + 2 * fdr. * d.
+
+  a12 <- 1 * ((1 - FPR) * (1 - d.)) + 
+    6 * mr. * (FPR * (1 - d.)) + 4 * fr. * (FPR * (1 - d.)) + 
+    6 * mdr. * d. + 4 * fdr. * d.
+
+  # a6
+  Fratio <- a6 / (1 + Fn. * (a6 - 1))
+  Em <- Pm. * (Fratio - 1 - q * rl.)
+  Ed <- Pd. * q
+  E6 <- Em - Ed
+
+  # a12
+  Fratio <- a12 / (1 + Fn. * (a12 - 1))
+  Em <- Pm. * (Fratio - 1 - q * rl.)
+  Ed <- Pd. * q
+  E12 <- Em - Ed
+  tibble(q=q,a6=a6,a12=a12,E6=E6,E12=E12)
+}
+
+#Do we get the same as before
+# > get_cost(TPR=1662/2116, FPR=1- (1662 /2116),d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) 
+# # A tibble: 1 x 5
+#        q    a6   a12     E6     E12
+#    <dbl> <dbl> <dbl>  <dbl>   <dbl>
+# 1 0.0257  1.35  1.93 -0.945 -0.0255
+#Yes
+
+tprs <- tprs %>% group_by(probability,TPR,FPR) %>%  do(get_cost(.$TPR,.$FPR) )
+
+```
+
+```{r}
+p <- ggplot(tprs)
+p <- p + geom_line(aes(x=probability,y=E6))
+p <- p + geom_line(aes(x=probability,y=E12)) + ylab("cost")
+p
+```

### 2020-09-21 src/eba_sims_w_data.Rmd
cleaning and testing variable probability cut-off

@@
 ```{r setup, echo = FALSE}
-setwd("~/interval-sims/src")
+#setwd("~/interval-sims/src")
 library(ggplot2)
-library(plotly)
-library(gridExtra)
-library(knitr)
-library(dplyr)
-library(plyr)
-library(lubridate)
-library(numbers)
-library(data.table)
-library(R.utils)
-library(forecast)
-source("pffunctions.R")
-source("helper_functions.R")
+library(tidyverse)
+library(truncnorm)
 ```
@@
 ```{r load_data, echo = FALSE}
-load("~/eba_sims/data/20191211CollectionAmounts.rdata")
+load("../data/20191211CollectionAmounts.rdata")
 ```
@@
 ```{r inspect}
-summary(donation10)
+lapply(donation10,class)
 ```
@@
 ```{r}
-library(truncnorm)
 
@@
 adjustments <- seq(1, 2, 1/1000)
-E.matrix <- matrix(, nrow = 1001, ncol = 1001)
-Pm <- 2
-Pd <- 60 # from Satu
-rl <- 0
-Fn = 0.1066 # Data
+E.matrix <- matrix(nrow = 1001, ncol = 1001)
+Pm <- 2 # estimate collected from FRCBS experts
+Pd <- 60 # estimate collected from FRCBS experts
+rl <- 0 # estimated previously from data
+Fn = 0.1066 # estimated from data above round(mean(counts_df[158:520,2])/mean(counts_df[158:520,3]), 4)
 for(i in 1:1001){
@@
 
-E.long <- melt(E.matrix)
+E.long <- reshape2::melt(E.matrix)
 

### 2020-09-22 src/current_model_effect.Rmd
prob cut wrong way twice :-)

@@
 for (i in x ) {
-  pred.classes[,paste0("p_",i)] <-  factor( df$Deferred <= i ,labels = c("Deferred","Accepted"))
+  pred.classes[,paste0("p_",i)] <-  factor( df$Deferred >= i ,labels = c("Accepted","Deferred"))
 }
@@
 head(pred.classes)
+```
+
+```{r}
+summary(pred.classes)
 ```
@@
 
+tprs
 ```

### 2020-09-22 src/current_model_effect.Rmd
started bootstrapping

@@
 library(caret)
+library(boot)
 source("helper_functions.R")
@@
 
-The current model finds 1662/2116 = 78.5% of deferrals. In addition, it misclassifies 21.45% of donations as deferrals.
+The current model finds 1662/2116 = 78.5% of deferrals. In addition, it misclassifies 22.4% of donations as deferrals.
 ```{r}
-TPR <- 1662/2116
-FPR <- 1- (1662 /2116)
+# > conf
+# Confusion Matrix and Statistics
+# 
+#           Reference
+# Prediction Accepted Deferred
+#   Accepted    31719      454
+#   Deferred     9137     1662
+TPR <- 1662/(1662 + 454) #78.5%
+FPR <- 9137 / (31719 + 9137) # 22.4%
 # Find q
@@
 scale_y_continuous(breaks = c(1, 1001), labels = c(as.character(min(adjustments)), as.character(max(adjustments))))
+ggsave(filename="../results/cme_cost_surface.pdf",  width = 180,  height = 120,units="mm", dpi=600, scale=1.0)
 ```
@@
 #count TPR and FPR for each probability
-s <- vector()
+sen <- vector()
+spe <- vector()
 #df$obs has factor levels in wrong
-df$obs <- relevel(df$obs,"Deferred")
+#df$obs <- relevel(df$obs,"Deferred")
 
@@
   #Sensitivity  = TPR
-  s <- c(s,conf$byClass['Sensitivity']) 
+  sen <- c(sen,conf$byClass['Sensitivity']) 
+#   > conf$table
+#           Reference
+# Prediction Accepted Deferred
+#   Accepted    31719      454
+#   Deferred     9137     1662
+# > conf$table[2,1] / (conf$table[1,1] + conf$table[2,1])
+# [1] 0.2236391
+# > 1- conf$byClass['Specificity'] 
+# Specificity 
+#   0.2236391 
+  spe <- c(spe,1- conf$byClass['Specificity'])
+
 }
-tprs <- tibble(probability = x,TPR=s,FPR=1-s)
+tprs <- tibble(probability = x,TPR=sen,FPR=spe)
 tprs
@@
 #Do we get the same as before
-# > get_cost(TPR=1662/2116, FPR=1- (1662 /2116),d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) 
+# > get_cost(TPR=1662/2116, FPR=9137 / (31719 + 9137),d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) 
 # # A tibble: 1 x 5
@@
 #    <dbl> <dbl> <dbl>  <dbl>  <dbl>
-# 1 0.0257  1.35  1.93 -0.945 -0.0255
+# 1 0.0257  1.36  1.97 -0.923 0.0256
 #Yes
+
+# > get_cost(TPR=0/2116, FPR=1- (0 /2116),d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) 
+# # A tibble: 1 x 5
+#       q    a6   a12    E6   E12
+#   <dbl> <dbl> <dbl> <dbl> <dbl>
+# 1     0  2.47  4.93  2.26  4.95
+# > get_cost(TPR=2116/2116, FPR=1- (2116 /2116),d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) 
+# # A tibble: 1 x 5
+#        q    a6   a12    E6   E12
+#    <dbl> <dbl> <dbl> <dbl> <dbl>
+# 1 0.0327  1.04  1.11 -1.89 -1.76
+#Yes, perfect model would give best savings
 
@@
 ```{r}
-p <- ggplot(tprs)
-p <- p + geom_line(aes(x=probability,y=E6))
-p <- p + geom_line(aes(x=probability,y=E12)) + ylab("cost")
+
+d <- tprs %>% ungroup() %>% select(probability,E6,E12) %>% pivot_longer(cols=c(E6,E12)) %>% mutate(name=gsub("^E",'',name))
+
+size <- 2
+p <- ggplot(d)
+p <- p + geom_line(aes(x=probability,y=value,col=name)) + ylab("Cost effect (€ / donation)") + xlab("Probability of deferral cut-off") + labs(color="Deferral length (mo)")
+p <- p + geom_point(aes( y =tprs$E6[which(tprs$E6 == min(tprs$E6))] , x = tprs$probability[which(tprs$E6 == min(tprs$E6))] ))
+p <- p + geom_label(aes( y =tprs$E6[which(tprs$E6 == min(tprs$E6))] , x = tprs$probability[which(tprs$E6 == min(tprs$E6))] , label = paste0("p=",round(tprs$probability[which(tprs$E6 == min(tprs$E6))],2) , " ; ",  round(tprs$E6[which(tprs$E6 == min(tprs$E6))],2)," €")), nudge_y=0.2,size=size)
+p <- p + geom_point(aes( y =tprs$E12[which(tprs$E12 == min(tprs$E12))] , x = tprs$probability[which(tprs$E12 == min(tprs$E12))] ))
+p <- p + geom_label(aes( y =tprs$E12[which(tprs$E12 == min(tprs$E12))] , x = tprs$probability[which(tprs$E12 == min(tprs$E12))] , label = paste0("p=",round(tprs$probability[which(tprs$E12 == min(tprs$E12))],2) , " ; ",  round(tprs$E12[which(tprs$E12 == min(tprs$E12))],2)," €")), nudge_y=0.2,size=size)
+p <- p + theme(legend.position="bottom")
 p
+ggsave(filename="../results/cme_prob_cut_curve.pdf", p, width = 90,  height = 80,units="mm", dpi=600, scale=1.0,encoding = "ISOLatin9.enc")
 ```
+
+Using a bit heigher than the default cut-off  appears slightly better.
+Let's bootstrap predictions to create error estimates for the predictions and subsequently on the cost effects.
+
+```{r}
+
+# https://www.painblogr.org/2017-10-18-purrring-through-bootstraps.html
+get_cost <- function(data, boot_ind){
+  #classify predictions
+  #get TPR and FPR
+    conf <- caret::confusionMatrix(reference=df$obs, data=pred.classes[,i], positive = "Deferred",mode="everything")
+    sen <- conf$byClass['Sensitivity'] 
+    spe <- 1- conf$byClass['Specificity'] 
+#gte cost
+    get_cost(TPR=sen, FPR=spe,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl)
+
+#    return(coef(fit))
+}
+
+fit_boot <- boot(data, statistic= get_coefficients, R = nb_boot)
+fit_boot_Bca <- compute_Bca_CI(fit_boot)
+
+```

### 2020-09-23 src/current_model_effect.Rmd
adding errors to surface plot

@@
 source("helper_functions.R")
+set.seed(123)
 ```
@@
 d <- nrow(wdef)/(nrow(wdef) + nrow(walldons))
+mr
+fr
+mdr
+fdr
+d
 ```
@@
 
-d <- tprs %>% ungroup() %>% select(probability,E6,E12) %>% pivot_longer(cols=c(E6,E12)) %>% mutate(name=gsub("^E",'',name))
+tprs.plot <- tprs %>% ungroup() %>% select(probability,E6,E12) %>% pivot_longer(cols=c(E6,E12)) %>% mutate(name=gsub("^E",'',name))
 
 size <- 2
-p <- ggplot(d)
+p <- ggplot(tprs.plot)
 p <- p + geom_line(aes(x=probability,y=value,col=name)) + ylab("Cost effect (€ / donation)") + xlab("Probability of deferral cut-off") + labs(color="Deferral length (mo)")
@@
 Using a bit heigher than the default cut-off  appears slightly better.
+
+# Bootstrap cost confidence intervals
+
 Let's bootstrap predictions to create error estimates for the predictions and subsequently on the cost effects.
@@
 ```{r}
-
+#Define functions
 # https://www.painblogr.org/2017-10-18-purrring-through-bootstraps.html
-get_cost <- function(data, boot_ind){
+boot_cost <- function(boot_data, boot_ind, prob.cut = 0.6){
+  #sample
+    boot_data <- boot_data[boot_ind,]
   #classify predictions
+    pred.class <- factor( boot_data$Deferred >= prob.cut)
+    levels(pred.class ) <- c("Accepted","Deferred")
   #get TPR and FPR
-    conf <- caret::confusionMatrix(reference=df$obs, data=pred.classes[,i], positive = "Deferred",mode="everything")
+    conf <- caret::confusionMatrix(reference=boot_data$obs, data=pred.class, positive = "Deferred",mode="sens_spec")
     sen <- conf$byClass['Sensitivity'] 
     spe <- 1- conf$byClass['Specificity'] 
-#gte cost
-    get_cost(TPR=sen, FPR=spe,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl)
-
-#    return(coef(fit))
+  #get cost
+#    cat(sen,spe,"\n")
+     c <- get_cost(TPR=sen, FPR=spe,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl)
+     c <- as.vector(t(c))
+     names(c) <- c('q','a6','a12','E6','E12')
+  return(c)
 }
 
-fit_boot <- boot(data, statistic= get_coefficients, R = nb_boot)
-fit_boot_Bca <- compute_Bca_CI(fit_boot)
+compute_CI <-function(fit_boot){
+  CI_inf = rep(0, length(fit_boot$t0))
+  CI_sup = rep(0, length(fit_boot$t0))
+  for (i_regressor in 1:length(fit_boot$t0)){
+    #https://stackoverflow.com/questions/6791290/r-boot-package-not-enough-memory-to-get-confidence-intervals
+    #CI <- boot.ci(fit_boot, type = "bca", index=i_regressor)
+    #Bca_inf[i_regressor] <- CI$bca[4] # N.B type = 'normal' has 3 columns, while the other types 5 columns
+    #Bca_sup[i_regressor] <- CI$bca[5] #
+    #there is just too much data for bca to be used
+    CI <- boot.ci(fit_boot, type = "norm", index=i_regressor)
+    CI_inf[i_regressor] <- CI$normal[2]
+    CI_sup[i_regressor] <- CI$normal[3]
+  }
+  return(tibble(variable = names(fit_boot$t0),CI_inf,CI_sup))
+}
+```
 
+```{r}
+
+nb_boot <- 1000
+fit_boot <- boot(df,prob.cut=0.6, statistic= boot_cost, R = nb_boot)
+p6.ci <- compute_CI(fit_boot)
+
+fit_boot <- boot(df,prob.cut=0.8, statistic= boot_cost, R = nb_boot)
+p8.ci <- compute_CI(fit_boot)
+
+p6.ci
 ```
+
+```{r}
+p8.ci
+```
+
+```{r}
+#a and q values do not come from tprs! they are not optimised
+
+colnames(E.long) <- c("adjustment", "q", "cost")
+p <- ggplot(E.long, aes(x = q, y = adjustment, z = cost)) +
+  geom_tile(aes(fill=cost)) +
+  stat_contour(color = "black", size = 0.7, bins = 2) +
+  scale_fill_gradientn(colours=c("blue", "red")) +
+  labs(x="q% i.e. deferral rate", y="Adjustment \n i.e. donation interval co-efficient ")  + 
+  annotate("point", x = head(which(q.seq > q ),1), y = which(adjustments > a6 & adjustments < (a6 + 0.001)), colour = "white", size = 2, alpha = 1) +
+  annotate("point", x = head(which(q.seq > q ),1), y = which(adjustments > a12 & adjustments < (a12 + 0.001)), colour = "white", size = 2, alpha = 1) +
+  theme_bw() +
+  scale_x_continuous(breaks = c(1, 1001), labels = c("0", "3.2")) +
+  scale_y_continuous(breaks = c(1, 1001), labels = c(as.character(min(adjustments)), as.character(max(adjustments))))
+#Add numeric values of costs with optimised classifier probability cut-offs
+p <- p + annotate("text" ,
+                  x = head(which(q.seq > q ),1) , 
+                  y = head(which(adjustments > a6),1) - 50,
+                  label = paste0("6 mo: ", round(tprs$E6[which(tprs$E6 == min(tprs$E6))],2),'(',round(deframe(p6.ci[4,2]),2)," - ",round(deframe(p6.ci[4,3]),2),') €'  ),
+                  colour = "white")
+
+p <- p + annotate("text" ,
+                  x = head(which(q.seq > q ),1) , 
+                  y = head(which(adjustments > a12),1) - 50,
+                  label = paste0("12 mo: ", round(tprs$E12[which(tprs$E12 == min(tprs$E12))],2),'(',round(deframe(p8.ci[5,2]),2)," - ",round(deframe(p8.ci[5,3]),2),') €'  ),
+                  colour = "white")
+
+#Add error bars of q and a
+# p <- p + annotate("text" ,
+#                   x = head(which(q.seq > q ),1) , 
+#                   y = head(which(adjustments > a12),1) - 50,
+#                   label = paste0("12 mo: ", round(E12,2),'(',round(deframe(p8.ci[5,2]),2)," - ",round(deframe(p8.ci[5,3]),2),') €'  ),
+#                   colour = "white")
+p <- p + labs(color="Cost per donation (€)")
+p
+ggsave(filename="../results/cme_cost_surface_werr.pdf",  width = 180,  height = 120,units="mm", dpi=600, scale=1.0,encoding = "ISOLatin9.enc")
+```
+
+```{r}
+labs(x="q%", y="Adjustment", title="Cost surface (2018-2020)", subtitle = paste0("Surface minimum: ", round(min(E.long$cost), 2), ".
+Cost effect thresholds highlighted at 0 and -1 euros.", "\n", "Dots mark current model performance with a 6 or 12 month extension strategy."))
+```

### 2020-09-23 src/current_model_effect.Rmd
error bootstrapping works

--- 2020-09-22 src/current_model_effect.Rmd
@@
 title: "Current model cost effect"
-author: "Esa Turkulainen"
-date: "8/25/2020"
+author: "Esa Turkulainen & Mikko Arvas"
+date: "`r Sys.time()`"
 output:
@@
 source("helper_functions.R")
+set.seed(123)
+#echo "rmarkdown::render('current_model_effect.Rmd', clean=TRUE,output_format='pdf_document',output_file='../results/current_model_effect.pdf')" | R --slave
 ```
+
+# Calculate parameters from donation history data
 
@@
 
-The data object donation10 contains donation history data from beginning of 2010 to end of 2019 it has following columns.
+The data object donation10 contains donation history data from beginning of 2010 to end of 2019. It has the following columns.
 
@@
 
-# Create subgroups from donation data
+## Create subgroups from donation data
 ```{r subgroups}
@@
 
-# Find ratios
+## Find ratios from data
 ```{r}
@@
 d <- nrow(wdef)/(nrow(wdef) + nrow(walldons))
+mr
+fr
+mdr
+fdr
+d
 ```
 
-The current model finds 1662/2116 = 78.5% of deferrals. In addition, it misclassifies 22.4% of donations as deferrals.
+# Find interval adjustments from ratios and model metrics
+
+The current model finds in validation data 1662/2116 = 78.5% of deferrals. In addition, it misclassifies 22.4% of donations as deferrals.
 ```{r}
@@
 # Find a for 6 month category
-# correctly predicted non-deferrals: 77.43% of the returning donations (all - d): (0.7743 * (1 - d))
-# incorrectly predicted non-deferrals: 22.57% of the returning donations (all - d): (0.2257 * (1 - d))
+# correctly predicted non-deferrals: 78.5% of the returning donations (all - d): (0.785 * (1 - d))
+# incorrectly predicted non-deferrals: 22.4% of the returning donations (all - d): (0.224 * (1 - d))
 a6 <- 1 * ((1 - FPR) * (1 - d)) + 
@@
 rl <- 0 # estimated previously from data
-Fn = 0.1066 # etimated from data in https://github.com/FRCBS/interval-sims/blob/master/src/eba_sims_w_data.Rmd
+Fn = 0.1066 # estimated from donation history data in here https://github.com/FRCBS/interval-sims/blob/master/src/eba_sims_w_data.Rmd
 
@@
 
-#What if we use some other than the default cut-off for model class assingment
+# What if we use some other than the default cut-off for model class assingment
 
@@
 ```
+
+## Create different cuts
 
@@
 
+# Count TPR and FPR for each probability cut-off
+
 ```{r}
-#count TPR and FPR for each probability
+
 sen <- vector()
 spe <- vector()
-#df$obs has factor levels in wrong
-#df$obs <- relevel(df$obs,"Deferred")
 
@@
   q <- d * TPR
-  # Find a for 6 month category
-  # correctly predicted non-deferrals: 77.43% of the returning donations (all - d): (0.7743 * (1 - d))
-  # incorrectly predicted non-deferrals: 22.57% of the returning donations (all - d): (0.2257 * (1 - d))
   a6 <- 1 * ((1 - FPR) * (1 - d.)) + 
@@
 
-# > get_cost(TPR=0/2116, FPR=1- (0 /2116),d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) 
-# # A tibble: 1 x 5
-#       q    a6   a12    E6   E12
-#   <dbl> <dbl> <dbl> <dbl> <dbl>
-# 1     0  2.47  4.93  2.26  4.95
-# > get_cost(TPR=2116/2116, FPR=1- (2116 /2116),d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl) 
-# # A tibble: 1 x 5
-#        q    a6   a12    E6   E12
-#    <dbl> <dbl> <dbl> <dbl> <dbl>
-# 1 0.0327  1.04  1.11 -1.89 -1.76
-#Yes, perfect model would give best savings
-
 tprs <- tprs %>% group_by(probability,TPR,FPR) %>%  do(get_cost(.$TPR,.$FPR) )
@@
 
-```{r}
+## Plot effect of probability cut-offs
 
-d <- tprs %>% ungroup() %>% select(probability,E6,E12) %>% pivot_longer(cols=c(E6,E12)) %>% mutate(name=gsub("^E",'',name))
+```{r , warning=FALSE}
+
+tprs.plot <- tprs %>% ungroup() %>% select(probability,E6,E12) %>% pivot_longer(cols=c(E6,E12)) %>% mutate(name=gsub("^E",'',name))
 
 size <- 2
-p <- ggplot(d)
+p <- ggplot(tprs.plot)
 p <- p + geom_line(aes(x=probability,y=value,col=name)) + ylab("Cost effect (€ / donation)") + xlab("Probability of deferral cut-off") + labs(color="Deferral length (mo)")
 p <- p + geom_point(aes( y =tprs$E6[which(tprs$E6 == min(tprs$E6))] , x = tprs$probability[which(tprs$E6 == min(tprs$E6))] ))
-p <- p + geom_label(aes( y =tprs$E6[which(tprs$E6 == min(tprs$E6))] , x = tprs$probability[which(tprs$E6 == min(tprs$E6))] , label = paste0("p=",round(tprs$probability[which(tprs$E6 == min(tprs$E6))],2) , " ; ",  round(tprs$E6[which(tprs$E6 == min(tprs$E6))],2)," €")), nudge_y=0.2,size=size)
+p <- p + geom_label(aes( y =tprs$E6[which(tprs$E6 == min(tprs$E6))] , x = tprs$probability[which(tprs$E6 == min(tprs$E6))] , label = paste0("p>=",round(tprs$probability[which(tprs$E6 == min(tprs$E6))],2) , " ; ",  round(tprs$E6[which(tprs$E6 == min(tprs$E6))],2)," €")), nudge_y=0.2,size=size)
 p <- p + geom_point(aes( y =tprs$E12[which(tprs$E12 == min(tprs$E12))] , x = tprs$probability[which(tprs$E12 == min(tprs$E12))] ))
-p <- p + geom_label(aes( y =tprs$E12[which(tprs$E12 == min(tprs$E12))] , x = tprs$probability[which(tprs$E12 == min(tprs$E12))] , label = paste0("p=",round(tprs$probability[which(tprs$E12 == min(tprs$E12))],2) , " ; ",  round(tprs$E12[which(tprs$E12 == min(tprs$E12))],2)," €")), nudge_y=0.2,size=size)
+p <- p + geom_label(aes( y =tprs$E12[which(tprs$E12 == min(tprs$E12))] , x = tprs$probability[which(tprs$E12 == min(tprs$E12))] , label = paste0("p>=",round(tprs$probability[which(tprs$E12 == min(tprs$E12))],2) , " ; ",  round(tprs$E12[which(tprs$E12 == min(tprs$E12))],2)," €")), nudge_y=0.2,size=size)
 p <- p + theme(legend.position="bottom")
@@
 
-Using a bit heigher than the default cut-off  appears slightly better.
+Using a bit higher than the default cut-off  appears slightly better.
+
+# Bootstrap cost confidence intervals
+
 Let's bootstrap predictions to create error estimates for the predictions and subsequently on the cost effects.
@@
 ```{r}
-
-# https://www.painblogr.org/2017-10-18-purrring-through-bootstraps.html
-get_cost <- function(data, boot_ind){
+#Define functions
+#boot code from https://github.com/FRCBS/changes_in_donor_health/blob/master/src/hypothesis_regressions.Rmd
+boot_cost <- function(boot_data, boot_ind, prob.cut = 0.6){
+  #sample
+    boot_data <- boot_data[boot_ind,]
   #classify predictions
+    pred.class <- factor( boot_data$Deferred >= prob.cut)
+    levels(pred.class ) <- c("Accepted","Deferred")
   #get TPR and FPR
-    conf <- caret::confusionMatrix(reference=df$obs, data=pred.classes[,i], positive = "Deferred",mode="everything")
+    conf <- caret::confusionMatrix(reference=boot_data$obs, data=pred.class, positive = "Deferred",mode="sens_spec")
     sen <- conf$byClass['Sensitivity'] 
     spe <- 1- conf$byClass['Specificity'] 
-#gte cost
-    get_cost(TPR=sen, FPR=spe,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl)
-
-#    return(coef(fit))
+  #get cost
+#    cat(sen,spe,"\n")
+     c <- get_cost(TPR=sen, FPR=spe,d.=d, mr.=mr,fr.=fr,mdr.=mdr,fdr.=fdr,Pm.=Pm,Pd.=Pd,Fn.=Fn,rl.=rl)
+     c <- as.vector(t(c))
+     names(c) <- c('q','a6','a12','E6','E12')
+  return(c)
 }
 
-fit_boot <- boot(data, statistic= get_coefficients, R = nb_boot)
-fit_boot_Bca <- compute_Bca_CI(fit_boot)
+compute_CI <-function(fit_boot){
+  CI_inf = rep(0, length(fit_boot$t0))
+  CI_sup = rep(0, length(fit_boot$t0))
+  for (i_regressor in 1:length(fit_boot$t0)){
+    #https://stackoverflow.com/questions/6791290/r-boot-package-not-enough-memory-to-get-confidence-intervals
+    #CI <- boot.ci(fit_boot, type = "bca", index=i_regressor)
+    #Bca_inf[i_regressor] <- CI$bca[4] # N.B type = 'normal' has 3 columns, while the other types 5 columns
+    #Bca_sup[i_regressor] <- CI$bca[5] #
+    #there is just too much data for bca to be used
+    CI <- boot.ci(fit_boot, type = "norm", index=i_regressor)
+    CI_inf[i_regressor] <- CI$normal[2]
+    CI_sup[i_regressor] <- CI$normal[3]
+  }
+  return(tibble(variable = names(fit_boot$t0),CI_inf,CI_sup))
+}
+```
 
+```{r}
+
+#nb_boot <- 1000
+nb_boot <- nrow(df)
+fit_boot <- boot(df,prob.cut=0.6, statistic= boot_cost, R = nb_boot)
+p6.ci <- compute_CI(fit_boot)
+
+fit_boot <- boot(df,prob.cut=0.8, statistic= boot_cost, R = nb_boot)
+p8.ci <- compute_CI(fit_boot)
+
+p6.ci
 ```
+
+```{r}
+p8.ci
+```
+
+## Plot surface with confidence intervals
+
+```{r ,warning=FALSE}
+#Get the a (y) and q (x) mappings for the optimised 6 and 12 month scenarios
+x6 <- head(which(q.seq > tprs$q[which(tprs$E6 == min(tprs$E6))] ),1)
+y6 <- head(which(adjustments > tprs$a6[which(tprs$E6 == min(tprs$E6))]),1)
+x12 <- head(which(q.seq > tprs$q[which(tprs$E12 == min(tprs$E12))] ),1)
+y12 <- head(which(adjustments > tprs$a12[which(tprs$E12 == min(tprs$E12))]),1)
+#Plot them
+colnames(E.long) <- c("adjustment", "q", "cost")
+p <- ggplot(E.long, aes(x = q, y = adjustment, z = cost)) +
+  geom_tile(aes(fill=cost)) +
+  stat_contour(color = "white", size = 0.7, binwidth =  1) +
+  scale_fill_gradientn(colours=c("blue", "red")) +
+  labs(x="q% i.e. deferral rate", y=expression(a[tot]~i.e.~donation~interval~coefficient ))  + 
+  theme_bw() +
+  scale_x_continuous(breaks = c(1, 1001), labels = c("0", "3.2")) +
+  scale_y_continuous(breaks = c(1, 1001), labels = c(as.character(min(adjustments)), as.character(max(adjustments))))
+#Add numeric values of costs with optimised classifier probability cut-offs
+p <- p + annotate("text" ,
+                  x = x6 , 
+                  y = y6 - 50,
+                  label = paste0("6 mo: ", round(tprs$E6[which(tprs$E6 == min(tprs$E6))],2),' (',round(deframe(p6.ci[4,2]),2)," - ",round(deframe(p6.ci[4,3]),2),') €'  ),
+                  colour = "white")
+
+p <- p + annotate("text" ,
+                  x = x12 , 
+                  y = y12 - 50,
+                  label = paste0("12 mo: ", round(tprs$E12[which(tprs$E12 == min(tprs$E12))],2),' (',round(deframe(p8.ci[5,2]),2)," - ",round(deframe(p8.ci[5,3]),2),') €'  ),
+                  colour = "white")
+#Add points estimates
+p <- p + annotate("point", x = x6, y = y6, colour = "white", size = 3, alpha = 1) 
+p <- p +  annotate("point", x = x12, y = y12, colour = "white", size = 3, alpha = 1)
+
+#Get x and y values for error bars
+xmin.p6  <- head(which(q.seq >  deframe(p6.ci[1,'CI_inf'])) ,1) 
+xmax.p6  <- head(which(q.seq >  deframe(p6.ci[1,'CI_sup'])) ,1) 
+ymin.p6  <- head(which(adjustments >  deframe(p6.ci[2,'CI_inf'])) ,1) 
+ymax.p6  <- head(which(adjustments >  deframe(p6.ci[2,'CI_sup'])) ,1)
+
+xmin.p12  <- head(which(q.seq >  deframe(p8.ci[1,'CI_inf'])) ,1) 
+xmax.p12  <- head(which(q.seq >  deframe(p8.ci[1,'CI_sup'])) ,1) 
+ymin.p12  <- head(which(adjustments >  deframe(p8.ci[3,'CI_inf'])) ,1) 
+ymax.p12  <- head(which(adjustments >  deframe(p8.ci[3,'CI_sup'])) ,1) 
+#Add error bars of q and a
+
+p <- p + annotate("errorbar",x=x6,ymin = ymin.p6, ymax= ymax.p6,size=1   )
+p <- p + annotate("errorbarh",y=y6,xmin = xmin.p6, xmax= xmax.p6,size=1   )
+
+p <- p + annotate("errorbar",x=x12,ymin = ymin.p12, ymax= ymax.p12,size=1  )
+p <- p + annotate("errorbarh",y=y12,xmin = xmin.p12, xmax= xmax.p12,size=1   )
+
+p <- p + labs(fill="Cost effect (€ / donation)")  + theme(legend.position="bottom")
+p
+ggsave(filename="../results/cme_cost_surface_werr.pdf",  width = 180,  height = 120,units="mm", dpi=600, scale=1.0,encoding = "ISOLatin9.enc")
+```

### 2020-09-24 src/current_model_effect.Rmd
critical x-axes label fix

@@
   scale_fill_gradientn(colours=c("blue", "red")) +
-  labs(x="q% i.e. deferral rate", y=expression(a[tot]~i.e.~donation~interval~coefficient ))  + 
+  labs(x="q% i.e. rate of avoided deferrals", y=expression(a[tot]~i.e.~donation~interval~coefficient ))  + 
   theme_bw() +

### 2020-09-30 src/current_model_effect.Rmd
plotting device improved

--- 2020-09-23 src/current_model_effect.Rmd
@@
   html_document: default
+dev: cairo_pdf 
 ---
@@
 p
-ggsave(filename="../results/cme_prob_cut_curve.pdf", p, width = 90,  height = 80,units="mm", dpi=600, scale=1.0,encoding = "ISOLatin9.enc")
+#ggsave(filename="../results/cme_prob_cut_curve.pdf", p, width = 90,  height = 80,units="mm", dpi=600, scale=1.0,encoding = "ISOLatin9.enc")
+ggsave(filename="../results/cme_prob_cut_curve.pdf", p, width = 90,  height = 80,units="mm", dpi=600, scale=1.0,device=cairo_pdf)
 ```
@@
                   y = y6 - 50,
-                  label = paste0("6 mo: ", round(tprs$E6[which(tprs$E6 == min(tprs$E6))],2),' (',round(deframe(p6.ci[4,2]),2)," - ",round(deframe(p6.ci[4,3]),2),') €'  ),
+                  label = paste0("6 mo: ", round(tprs$E6[which(tprs$E6 == min(tprs$E6))],2),' (',round(deframe(p6.ci[4,2]),2)," – ",round(deframe(p6.ci[4,3]),2),') €'  ),
                   colour = "white")
@@
                   y = y12 - 50,
-                  label = paste0("12 mo: ", round(tprs$E12[which(tprs$E12 == min(tprs$E12))],2),' (',round(deframe(p8.ci[5,2]),2)," - ",round(deframe(p8.ci[5,3]),2),') €'  ),
+                  label = paste0("12 mo: ", round(tprs$E12[which(tprs$E12 == min(tprs$E12))],2),' (',round(deframe(p8.ci[5,2]),2)," – ",round(deframe(p8.ci[5,3]),2),') €'  ),
                   colour = "white")
@@
 p
-ggsave(filename="../results/cme_cost_surface_werr.pdf",  width = 180,  height = 120,units="mm", dpi=600, scale=1.0,encoding = "ISOLatin9.enc")
+#ggsave(filename="../results/cme_cost_surface_werr.pdf",  width = 180,  height = 120,units="mm", dpi=600, scale=1.0,encoding = "ISOLatin9.enc")
+ggsave(filename="../results/cme_cost_surface_werr.pdf",  width = 180,  height = 120,units="mm", dpi=600, scale=1.0, device=cairo_pdf)
+#check this from the performance plot
 ```

### 2020-10-30 src/article_figures.Rmd
Added Rmd that generates all article figures

---
title: "Article figures"
author: "Jarkko Toivonen"
date: "10/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
save_figs = FALSE
fig_path="~/FRCBS/results-for-eba-paper/pdf"
```

## Effect sizes and importances of variables

Effect sizes of variables from eProgesa and Biobank using dynamic linear mixed model. In addition, variables
importance from random forest algorithm.

```{r}
# Creates a forest plot of means and confidence intervals from posterior distributions.
# Puts both male and female results in the same plot.
create_double_forest_plot <- function(male_posterior, female_posterior, variables, combined_variables, base_size = 11) {
  for (gender in c("male", "female")) {
    #posterior <- ifelse(gender == "male", male_posterior, female_posterior)
    if (gender == "male") {
      posterior <- male_posterior
    } else {
      posterior <- female_posterior
    }

    for (i in seq_along(posterior)) {
      #cat(sprintf("Gender %s, Column %i\n", gender, i))
      v <- posterior[[i]]
      ci_hdi <- bayestestR::ci(v, method = "HDI", ci=0.95)
      #str(ci_hdi)
      L <- list(names = variables[[i]], gender=gender, mean = mean(v), low = ci_hdi$CI_low, high = ci_hdi$CI_high)
      if (i == 1 && gender=="male") {
        result <- data.frame(L, stringsAsFactors = FALSE)
      } else {
        result <- rbind(result, L)
      }
    }
  }
  result <- as_tibble(result)
  result <- result %>%
    mutate(gender = factor(gender, levels=c("female", "both", "male")))
  cis <- result
  #result <- DescTools::Rev(result, margin=1)   # Reverse the order of rows so they appear in correct order in the forest plot
  print(head(result))
  combined_variables <- combined_variables %>% rename(names=Pretty)
  result <- left_join(combined_variables, result, by="names")
  plot <- result %>% ggplot(aes(y=factor(names, levels=rev(combined_variables$names)), x=mean, xmin=low, xmax=high, color=gender)) +     
    geom_vline(aes(xintercept=0), color="lightgray") +
    #geom_pointrange() +
    ggstance::geom_pointrangeh(position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
    labs(title="", #title="Effects sizes of variables on Hb prediction",
         x="Regression coefficient", y=NULL, colour="Sex") +
    scale_colour_discrete(drop=FALSE, breaks = c("male", "female")) + 
    #guides(colour = guide_legend(reverse = TRUE)) + # Reverse the order of keys in the legend
    theme_gray(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")

  return(list(plot=plot, cis=cis))
}

# create_variable_importance_plot <- function(rrfFit_roc, descript, base_size = 11) {
#   rrfFit_rocImp <- varImp(rrfFit_roc, scale = FALSE)
#   rrfFit.varimp <- as_tibble(cbind(rownames(rrfFit_rocImp$importance),rrfFit_rocImp$importance))
#   colnames(rrfFit.varimp) <- c("Variable","Importance")
#   rrfFit.varimp <- left_join(rrfFit.varimp, descript, by=c("Variable"="Variable")) %>% select(Variable,Pretty,Importance) %>% arrange(Importance)
#   
#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "previous_Hb_defTRUE"] <- "Previous donation deferred"
#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "warm_seasonTRUE"] <- "Warm season"
#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "genderWomen"] <- "Sex"
#   
#   varimp.plot <- ggplot(rrfFit.varimp) + 
#     geom_col(aes(y=Importance, x=reorder(Pretty, Importance)), alpha=0.7) + 
#     coord_flip() +
#     xlab(NULL) + 
#     theme_gray(base_size = base_size)
#     #xlab("Variable")
#   
#   #filename="../results/rrfFit_roc_importance.pdf"
#   #ggsave(filename=filename, varimp.plot, width = 180,  height = 80,units="mm", dpi=600, scale=1.0)
#   return(varimp.plot)
# }

create_variable_importance_plot2 <- function(rrfFit_roc, descript, combined_variables, base_size = 11) {
  rrfFit_rocImp <- varImp(rrfFit_roc, scale = FALSE)
  rrfFit.varimp <- as_tibble(cbind(rownames(rrfFit_rocImp$importance),rrfFit_rocImp$importance))
  colnames(rrfFit.varimp) <- c("Variable","Importance")
  rrfFit.varimp <- left_join(rrfFit.varimp, descript, by=c("Variable"="Variable")) %>% select(Variable,Pretty,Importance) %>% arrange(Importance)

  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "previous_Hb_defTRUE"] <- "Previous donation deferred"
  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "warm_seasonTRUE"] <- "Warm season"
  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "genderWomen"] <- "Sex"

  #pretty <- descript["Pretty"]
  pretty <- combined_variables
  rrfFit.varimp <- left_join(pretty, rrfFit.varimp, by="Pretty")
  varimp.plot <- rrfFit.varimp %>% 
    ggplot(aes(y=factor(Pretty, levels=rev(Pretty)), x=Importance, xmin=0, xmax=Importance, colour = factor("both", levels=c("female", "both", "male")))) + 
    ggstance::geom_pointrangeh(position=position_dodge2(width=0.8, padding=0.1)) +  # ggstance is required to make legend keys horizontal
                               #colour = hue_pal()(3)[2], size=0.25) +
    labs(y=NULL, colour="Sex") + 
    guides(colour="none", y="none") +
    scale_colour_discrete(drop=FALSE) +
    theme_gray(base_size = base_size)

  return(varimp.plot)
}

create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
                                           filename=NULL,
                                           width = 180,  # width of the combined figure in millimetres
                                          base_size = 11
) {
  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
  combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb"))
  forest <- create_double_forest_plot(male_posterior, female_posterior, variables, combined_variables, base_size = base_size)$plot
  importance <- create_variable_importance_plot2(rrfFit_roc, descript, combined_variables, base_size = base_size)

  use_cowplot <- TRUE

  if (use_cowplot) {
    forest_importance <- cowplot::plot_grid(forest, importance, labels = c('A', 'B'), label_size = 12, nrow=1, scale=1.0, axis="tb", align="h",
                                            rel_widths = c(2, 1))
    if (!is.null(filename)) {
      cowplot::save_plot(filename, forest_importance, title="Effect sizes and importances",
                         ncol = 2, base_asp = 1.0, base_width = width / 25.4 / 2, base_height = NULL)
    }
  } else {
    forest_importance <- gridExtra::grid.arrange(forest, importance, nrow = 1, respect=TRUE)   # Combine the plots
    if (!is.null(filename)) {
      ggsave(filename=filename, forest_importance, width = width, units="mm", dpi=600, scale=1.0)
    }
  }
  forest_importance
}
```

```{r}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
male_posterior <- finngenn_male_raw$samples
female_posterior <- finngenn_female_raw$samples
variables <- finngenn_male_raw$pretty_variable_names
```

```{r}
if (save_figs) {
    filename <- paste(fig_path, "effect_size_importance.pdf", sep="/")
} else {
    filename <- NULL
}
create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
```

## Performance forest plot

```{r}
# Show AUROC, AUPR, F1, E6, and E12 as forest plot
create_performance_forest_plot <- function() {
  df <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
  df <- df %>%
    mutate(Id=factor(Id, levels=rev(ids)),
           type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
           sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
           model=factor(str_split_fixed(Id, "-", 3)[,3]))

  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
  xrange <- as.numeric(df %>% filter(type %in% c("E6", "E12")) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
  xrange <- rep(xrange, 2)
  mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, type=mytype)  # Dummy table to use the same x-axis limits for E6 and E12
  #print(dummy)

  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
    #ggstance::geom_pointrangeh(aes(colour=sex, shape=model)) + 
    geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
    labs(x="Value", y="Data and method", colour="Sex", shape="Model") + 
    scale_colour_discrete(breaks = c("male", "female", "both")) +
    scale_shape_discrete(breaks=c("lmm", "dlmm", "dt", "rf")) +
    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
    geom_blank(data=dummy) +
    facet_wrap("type", scales="free_x")
  #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
  return(g)
}

```

```{r}
g <- create_performance_forest_plot()
if (save_figs)
  ggsave(filename="performance_forest_plot.pdf", title="Performance forest plot", path=fig_path, plot=g, dpi=600, units="mm", width=180)
g
```
## Calibration plots

```{r cars}
g <- calibration_plots(ids)
if (save_figs)
  ggsave(filename="calibration_plots.pdf", title="Calibration plots", path=fig_path, plot=g, dpi=600, units="mm", width=180)
g
```

## Classification scatter plot for male Finngen DLMM

```{r}

generate_my_breaks <- function(step) {
  # Break between limits at every position that is multiple of 'step' 
  my_breaks <- function(limits) {
    #step <- 0.2
    m <- limits %/% step
    m <- ifelse(m < 0, m+1, m)
    m <- m*step
    return(seq(m[1], m[2], step))
  }
  return(my_breaks)
}

create_classification_scatter_plot <- function(df, hb_threshold, probability_of_deferral_threshold) {
  xymin <- min(min(df$predicted), min(df$observed))
  xymax <- max(max(df$predicted), max(df$observed))
  df <- df %>% 
    mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0)) %>%
    mutate(confusion_class = factor(ifelse(deferral == 1, 
                                    ifelse(new_predicted_label == 1, "True positive", "False negative"),
                                    ifelse(new_predicted_label == 1, "False positive", "True negative")),
           levels=c("True positive", "False negative", "False positive", "True negative")))
  scatter_plot <- ggplot(df, aes(x = observed, y=predicted, color = confusion_class)) +
    geom_point() +
    #xlim(xymin,xymax) + ylim(xymin,xymax) +
    scale_x_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    scale_y_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
    geom_abline(intercept = 0, slope = 1) +
    labs(x = "Observed", y = "Predicted", colour = "Deferral status") +
    #scale_colour_discrete(labels=c("Accepted", "Deferred")) +
    geom_smooth(mapping=aes(x = observed, y=predicted), colour="black", show.legend = FALSE) +
    geom_vline(xintercept = hb_threshold, linetype = "dashed") +
    geom_hline(yintercept = hb_threshold, linetype = "dashed") +
    theme(legend.position = "bottom") +
    ggtitle("Observed vs predicted Hb-values")
  return(scatter_plot)
}
```

```{r}
finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")

```

```{r}
g <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.1)
if (save_figs)
  ggsave(filename="classification-scatter-plot-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
         plot=g, dpi=600, units="mm", width=180)
g
```

### 2020-10-30 src/economic_effect_functions.R

@@ -4,6 +4,35 @@ library(tidyverse)
 library(caret)
 library(boot)
 
+#ids <- c("progesa-female-lmm", "progesa-female-dlmm", "progesa-male-lmm", "progesa-male-dlmm", "finngen-male-dlmm", "finngen-female-dlmm", 
+#         "findonor-female-dlmm", "findonor-male-dlmm", "progesa-both-dt", "progesa-both-rf")
+ids <- c("progesa-male-lmm", "progesa-female-lmm", "progesa-male-dlmm", "progesa-female-dlmm", "finngen-male-dlmm", "finngen-female-dlmm", 
+         "findonor-male-dlmm", "findonor-female-dlmm", "progesa-both-dt", "progesa-both-rf")
+
+dummy_ids <- c("finngen-male-stratified", "finngen-male-most-frequent", "finngen-male-prior", "finngen-male-uniform", "finngen-male-deferred")
+
+# These are from Progesa data
+get_mikkos_cost_constants <- function() {
+  return(list(
+    mr=0.4732352,
+    fr=0.5267648,
+    mdr=0.2335795,
+    fdr=0.7664205,
+    d=0.03269718
+  ))
+}
+
+data_parameters <- get_mikkos_cost_constants()
+fixed_parameters <- list(Pm=2, Pd=60, Fn=0.1066, rl=0)
+
+parameters <- c(data_parameters, fixed_parameters)
+
+load_single <- function(filename) {
+  names <- load(filename, verbose=FALSE)
+  stopifnot(length(names) == 1)
+  return(get(names))
+}
+
 filter_donations <- function(data, 
                              donor=NULL, donation13=NULL, site=NULL,
                              date=NULL, phleb_start=NULL, status=NULL,
@@ -107,21 +136,6 @@ get_cost_constants <- function(data) {
   ))
 }
 
-# These are from Progesa data
-get_mikkos_cost_constants <- function() {
-  return(list(
-    mr=0.4732352,
-    fr=0.5267648,
-    mdr=0.2335795,
-    fdr=0.7664205,
-    d=0.03269718
-  ))
-}
-
-data_parameters <- get_mikkos_cost_constants()
-fixed_parameters <- list(Pm=2, Pd=60, Fn=0.1066, rl=0)
-
-parameters <- c(data_parameters, fixed_parameters)
 
[…]

### 2020-11-02 src/article_figures.Rmd
Improved article figures

@@
 date: "10/30/2020"
-output: html_document
+output: 
+  html_document: default
+  md_document: default
 ---
@@
 
+```{r}
+
+descript <- tibble(Variable = c("donor", "Hb", "days_to_previous_fb", "age", "previous_Hb_def", 
+                                "year", "warm_season", "consecutive_deferrals", "recent_donations",
+                                "recent_deferrals", "hour", 
+                                "previous_Hb", "Hb_first", "Hb_deferral","nb_donat","gender"), 
+                   Pretty = c("Donor ID", "Hemoglobin", "Days to previous full blood donation", "Age", 
+                              "Previous donation deferred", #"Previous Hb deferral",   # This is a more descriptive name
+                              "Year", "Warm season", "Consecutive deferrals", "Recent donations", 
+                              "Recent deferrals", "Hour", 
+                              "Previous Hb", "First Hb", "Hb deferral","Life time donations","Sex"),
+                   Type = c("Factor", "numeric", "numeric (int)", "numeric", "boolean",
+                            "numeric (int)", "boolean", "numeric (int)", "numeric (int)", "numeric (int)", "numeric",
+                            "numeric", "numeric", "boolean","numeric (int)","Factor"),
+                   Explanation = c("Donor identifier",
+                                   "Amount of Hemoglobin",
+                                   "Time (in days) between Hb measurement and previous full blood donation event",
+                                   "Age of donor",
+                                   "Indicates whether the donor was deferred from blood donation due to low hemoglobin at previous donation event",
+                                   "Year of donation",
+                                   "True if donation was given in April-September",
+                                   "Amount of times the donor has been deferred due to low hemoglobin since last succesful whole blood donation",
+                                   "Amount of donations in the last two years",
+                                   "Amount of deferrals due to low hemoglobin in the last two years",
+                                   "Time of day when donation was given as hours (e.g. 13:45 = 13.75)",
+                                   "Hb value at previous measurement (ICP-model)",
+                                   "Hb value at first donation of this donor (Non ICP-model)",
+                                   "Deferred based on low hemoglogin",
+                                  "Life time total count of donations",
+                                  "Sex"
+                   )
+                   )
+
+```
+
 ## Effect sizes and importances of variables
@@
 ```{r}
+
+# create_variable_importance_plot <- function(rrfFit_roc, descript, base_size = 11) {
+#   rrfFit_rocImp <- varImp(rrfFit_roc, scale = FALSE)
+#   rrfFit.varimp <- as_tibble(cbind(rownames(rrfFit_rocImp$importance),rrfFit_rocImp$importance))
+#   colnames(rrfFit.varimp) <- c("Variable","Importance")
+#   rrfFit.varimp <- left_join(rrfFit.varimp, descript, by=c("Variable"="Variable")) %>% select(Variable,Pretty,Importance) %>% arrange(Importance)
+#   
+#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "previous_Hb_defTRUE"] <- "Previous donation deferred"
+#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "warm_seasonTRUE"] <- "Warm season"
+#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "genderWomen"] <- "Sex"
+#   
+#   varimp.plot <- ggplot(rrfFit.varimp) + 
+#     geom_col(aes(y=Importance, x=reorder(Pretty, Importance)), alpha=0.7) + 
+#     coord_flip() +
+#     xlab(NULL) + 
+#     theme_gray(base_size = base_size)
+#     #xlab("Variable")
+#   
+#   #filename="../results/rrfFit_roc_importance.pdf"
+#   #ggsave(filename=filename, varimp.plot, width = 180,  height = 80,units="mm", dpi=600, scale=1.0)
+#   return(varimp.plot)
+# }
+
 # Creates a forest plot of means and confidence intervals from posterior distributions.
@@
   result <- result %>%
-    mutate(gender = factor(gender, levels=c("female", "both", "male")))
+    mutate(gender = factor(gender, levels=c("female", "both", "male"))) %>%
+    rename(Pretty=names) %>%
+    mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)))
   cis <- result
@@
   print(head(result))
-  combined_variables <- combined_variables %>% rename(names=Pretty)
-  result <- left_join(combined_variables, result, by="names")
-  plot <- result %>% ggplot(aes(y=factor(names, levels=rev(combined_variables$names)), x=mean, xmin=low, xmax=high, color=gender)) +     
+  #combined_variables <- combined_variables %>% 
+    #rename(names=Pretty) %>%
+    #mutate(stripe=factor(row_number()%%2))
+  result <- left_join(combined_variables, result, by="Pretty") %>%
+    mutate(Pretty=fct_rev(Pretty))
+
+    #mutate(Pretty=factor(Pretty, levels=rev(combined_variables$Pretty)))
+    # group_by(gender) %>%
+    # mutate(stripe=factor(row_number()%%2)) %>%
+    # ungroup()
+  df1 <<- result
+  plot <- result %>% 
+    # mutate(Pretty=factor(names, levels=rev(combined_variables$names)), 
+    #        stripe=factor(row_number()%%2)) %>%
+    ggplot() +     
     geom_vline(aes(xintercept=0), color="lightgray") +
     #geom_pointrange() +
-    ggstance::geom_pointrangeh(position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
+    ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
+                               position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
+    geom_rect(data=combined_variables %>% filter(stripe==1), 
+              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
+                  ymin = as.numeric(Pretty) - 0.5),
+                  #xmin = -Inf,
+                  #xmax = Inf,
+                  fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA) +
     labs(title="", #title="Effects sizes of variables on Hb prediction",
@@
     #guides(colour = guide_legend(reverse = TRUE)) + # Reverse the order of keys in the legend
-    theme_gray(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
-
+    #theme_gray(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
+    theme_classic(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
   return(list(plot=plot, cis=cis))
 }
-
-# create_variable_importance_plot <- function(rrfFit_roc, descript, base_size = 11) {
-#   rrfFit_rocImp <- varImp(rrfFit_roc, scale = FALSE)
-#   rrfFit.varimp <- as_tibble(cbind(rownames(rrfFit_rocImp$importance),rrfFit_rocImp$importance))
-#   colnames(rrfFit.varimp) <- c("Variable","Importance")
-#   rrfFit.varimp <- left_join(rrfFit.varimp, descript, by=c("Variable"="Variable")) %>% select(Variable,Pretty,Importance) %>% arrange(Importance)
-#   
-#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "previous_Hb_defTRUE"] <- "Previous donation deferred"
-#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "warm_seasonTRUE"] <- "Warm season"
-#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "genderWomen"] <- "Sex"
-#   
-#   varimp.plot <- ggplot(rrfFit.varimp) + 
-#     geom_col(aes(y=Importance, x=reorder(Pretty, Importance)), alpha=0.7) + 
-#     coord_flip() +
-#     xlab(NULL) + 
-#     theme_gray(base_size = base_size)
-#     #xlab("Variable")
-#   
-#   #filename="../results/rrfFit_roc_importance.pdf"
-#   #ggsave(filename=filename, varimp.plot, width = 180,  height = 80,units="mm", dpi=600, scale=1.0)
-#   return(varimp.plot)
-# }
 
@@
   #pretty <- descript["Pretty"]
-  pretty <- combined_variables
-  rrfFit.varimp <- left_join(pretty, rrfFit.varimp, by="Pretty")
-  varimp.plot <- rrfFit.varimp %>% 
-    ggplot(aes(y=factor(Pretty, levels=rev(Pretty)), x=Importance, xmin=0, xmax=Importance, colour = factor("both", levels=c("female", "both", "male")))) + 
-    ggstance::geom_pointrangeh(position=position_dodge2(width=0.8, padding=0.1)) +  # ggstance is required to make legend keys horizontal
+  #pretty <- combined_variables
+  rrfFit.varimp <- rrfFit.varimp %>% mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)))
+  rrfFit.varimp <- left_join(combined_variables, rrfFit.varimp, by="Pretty")
+
+  var_imp <- rrfFit.varimp %>%
+      mutate(Pretty=fct_rev(Pretty),
+             gender=factor("both", levels=c("female", "both", "male"))) 
+             #stripe=factor(row_number()%%2))
+  print(var_imp) 
+  df2 <<- var_imp
+  varimp.plot <- var_imp %>% 
+      # rrfFit.varimp %>%
+      # mutate(Pretty=factor(Pretty, levels=rev(Pretty)), 
+      #        stripe=factor(row_number()%%2)) %>%
+    ggplot() + 
+    ggstance::geom_pointrangeh(aes(y=Pretty, x=Importance, xmin=0, xmax=Importance, colour = gender), # ggstance is required to make legend keys horizontal
+                               position=position_dodge2(width=0.8, padding=0.1), size=0.25, inherit.aes = FALSE) +   
                                #colour = hue_pal()(3)[2], size=0.25) +
+      geom_rect(data=combined_variables %>% filter(stripe==1),
+              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
+                    ymin = as.numeric(Pretty) - 0.5),
+                    #xmin = -Inf,
+                    #xmax = Inf,
+                    #fill = stripe), 
+              xmin=-Inf, xmax=Inf, alpha = 0.1, colour=NA, fill="gray", show.legend = FALSE, inherit.aes = FALSE) +
     labs(y=NULL, colour="Sex") + 
-    guides(colour="none", y="none") +
-    scale_colour_discrete(drop=FALSE) +
-    theme_gray(base_size = base_size)
+    guides(y="none") +
+    scale_colour_discrete(drop=FALSE, breaks=c("both")) +
+    #theme_gray(base_size = base_size)
+    theme_classic(base_size = base_size) + theme(legend.position = "bottom", legend.direction = "horizontal")
 
@@
   rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
-  combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb"))
+  combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
+    mutate(stripe=factor(row_number()%%2),
+           Pretty=factor(Pretty, levels=Pretty))
+  combined_variables <<- combined_variables
   forest <- create_double_forest_plot(male_posterior, female_posterior, variables, combined_variables, base_size = base_size)$plot
@@
 variables <- finngenn_male_raw$pretty_variable_names
+# Give a more descriptive name to this variable
+variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred") #"Previous donation deferred", #"Previous Hb deferral"
 ```
@@
 
+  #key_size <- as.numeric(grid::convertX(theme_get()$legend.key.size, "mm"))
+  key_size <- 1.8
+  #key_size <- unit(1.2, "lines")
   g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
-    #ggstance::geom_pointrangeh(aes(colour=sex, shape=model)) + 
-    geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
+    ggstance::geom_pointrangeh(aes(colour=sex, shape=model)) + 
+    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
     labs(x="Value", y="Data and method", colour="Sex", shape="Model") + 
+    #scale_colour_manual(breaks = c("male", "female", "both"), values=c("yellow", "brown", "black"), aesthetics = c("color", "fill", "shape")) +
     scale_colour_discrete(breaks = c("male", "female", "both")) +
@@
     #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
+    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size))) +
     geom_blank(data=dummy) +
@@
 ## Calibration plots
+
+```{r calibration plots}
+calibration_plot <- function(df, use_my_method=FALSE) {
+  if (use_my_method) {
+    df$cat <- cut(df$Deferred, seq(0.0, 1.0, length.out = 11))
+    df2 <- df %>% group_by(cat, .drop=FALSE) %>% summarise(true_bin_prob = mean(obs=="Deferred"))
+    df2 %>% ggplot(aes(cat, true_bin_prob)) + geom_point() + ylim(0, 1)
+  } else {
+    cal <- calibration(obs ~ Deferred, data=df, class="Deferred", cuts=11)
+    ggplot(cal) + ylim(0, 100)
+  }
+}
+
+calibration_plots <- function(ids) {
+  data_frames <- map(ids, get_data_frame)
+  names(data_frames) <- ids
+  df <- bind_rows(data_frames, .id="Id")
+  df <- df %>% 
+    group_by(Id) %>%
+    mutate(cat=cut(Deferred, seq(0.0, 1.0, length.out = 11))) %>%
+    group_by(Id, cat, .drop=FALSE) %>% summarise(true_bin_prob = mean(obs=="Deferred"), count=n()) %>%
+    ungroup() %>%
+    mutate(Id = factor(Id, levels=ids))
+
+  g <- df %>% ggplot(aes(cat, true_bin_prob, colour=count)) + 
+    geom_point() + 
+    geom_text(aes(label=count), colour="black", nudge_y=0.05) +
+    ylim(0, 1.05) +
+    labs(x="Predicted probability of deferral", y="True probability of deferral in each bin") +
+    scale_colour_gradient(name = "Count", trans = "log10") +
+    geom_abline(aes(intercept=-0.1, slope=0.1)) +
+    facet_wrap("Id") +
+    theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
+
+  #return(df)
+  return(g)
+}
+```
 

### 2020-11-02 src/economic_effect_functions.R

@@ -504,42 +504,7 @@ test_get_all_costs <- function() {
 }
 
 
-calibration_plot <- function(df, use_my_method=FALSE) {
-  if (use_my_method) {
-    df$cat <- cut(df$Deferred, seq(0.0, 1.0, length.out = 11))
-    df2 <- df %>% group_by(cat, .drop=FALSE) %>% summarise(true_bin_prob = mean(obs=="Deferred"))
-    df2 %>% ggplot(aes(cat, true_bin_prob)) + geom_point() + ylim(0, 1)
-  } else {
-    cal <- calibration(obs ~ Deferred, data=df, class="Deferred", cuts=11)
-    ggplot(cal) + ylim(0, 100)
-  }
-}
 
-
-calibration_plots <- function(ids) {
-  data_frames <- map(ids, get_data_frame)
-  names(data_frames) <- ids
-  df <- bind_rows(data_frames, .id="Id")
-  df <- df %>% 
-    group_by(Id) %>%
-    mutate(cat=cut(Deferred, seq(0.0, 1.0, length.out = 11))) %>%
-    group_by(Id, cat, .drop=FALSE) %>% summarise(true_bin_prob = mean(obs=="Deferred"), count=n()) %>%
-    ungroup() %>%
-    mutate(Id = factor(Id, levels=ids))
-  
-  g <- df %>% ggplot(aes(cat, true_bin_prob, colour=count)) + 
-    geom_point() + 
-    geom_text(aes(label=count), colour="black", nudge_y=0.05) +
-    ylim(0, 1.05) +
-    labs(x="Predicted probability of deferral", y="True probability of deferral in each bin") +
-    scale_colour_gradient(name = "Count", trans = "log10") +
-    geom_abline(aes(intercept=-0.1, slope=0.1)) +
-    facet_wrap("Id") +
-    theme(axis.text.x = element_text(angle = 45, vjust = 0.5))
-  
-  #return(df)
-  return(g)
-}
 
 roc_wrapper <- function(df) {
   df <- df %>% rename(scores = Deferred, labels=obs) %>% mutate(labels = ifelse(labels=="Accepted", 0, 1))

### 2020-11-03 src/article_figures.Rmd
Updated article figures

@@
 output: 
+  md_document: default
   html_document: default
-  md_document: default
 ---
@@
 
-# create_variable_importance_plot <- function(rrfFit_roc, descript, base_size = 11) {
-#   rrfFit_rocImp <- varImp(rrfFit_roc, scale = FALSE)
-#   rrfFit.varimp <- as_tibble(cbind(rownames(rrfFit_rocImp$importance),rrfFit_rocImp$importance))
-#   colnames(rrfFit.varimp) <- c("Variable","Importance")
-#   rrfFit.varimp <- left_join(rrfFit.varimp, descript, by=c("Variable"="Variable")) %>% select(Variable,Pretty,Importance) %>% arrange(Importance)
-#   
-#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "previous_Hb_defTRUE"] <- "Previous donation deferred"
-#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "warm_seasonTRUE"] <- "Warm season"
-#   rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "genderWomen"] <- "Sex"
-#   
-#   varimp.plot <- ggplot(rrfFit.varimp) + 
-#     geom_col(aes(y=Importance, x=reorder(Pretty, Importance)), alpha=0.7) + 
-#     coord_flip() +
-#     xlab(NULL) + 
-#     theme_gray(base_size = base_size)
-#     #xlab("Variable")
-#   
-#   #filename="../results/rrfFit_roc_importance.pdf"
-#   #ggsave(filename=filename, varimp.plot, width = 180,  height = 80,units="mm", dpi=600, scale=1.0)
-#   return(varimp.plot)
-# }
-
-# Creates a forest plot of means and confidence intervals from posterior distributions.
-# Puts both male and female results in the same plot.
-create_double_forest_plot <- function(male_posterior, female_posterior, variables, combined_variables, base_size = 11) {
+get_dlmm_df <- function(male_posterior, female_posterior, variables, combined_variables) {
     for (gender in c("male", "female")) {
@@
     mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)))
+  result <- left_join(combined_variables, result, by="Pretty")
+  return(result)
+
+}
+
+get_rf_df <- function(rrfFit_roc, descript, combined_variables) {
+  rrfFit_rocImp <- varImp(rrfFit_roc, scale = FALSE)
+  #rrfFit.varimp <- as_tibble(cbind(rownames(rrfFit_rocImp$importance),rrfFit_rocImp$importance))
+  rrfFit.varimp <- rownames_to_column(rrfFit_rocImp$importance)
+  colnames(rrfFit.varimp) <- c("Variable","Importance")
+  rrfFit.varimp <- left_join(rrfFit.varimp, descript, by=c("Variable"="Variable")) %>% select(Variable, Pretty, Importance) %>% arrange(Importance)
+
+  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "previous_Hb_defTRUE"] <- "Previous donation deferred"
+  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "warm_seasonTRUE"] <- "Warm season"
+  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "genderWomen"] <- "Sex"
+
+  rrfFit.varimp <- rrfFit.varimp %>% 
+    mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)),
+           gender=factor("both", levels=c("female", "both", "male")))
+  rrfFit.varimp <- left_join(combined_variables, rrfFit.varimp, by="Pretty")
+  return(rrfFit.varimp)  
+}
+
+# Creates a forest plot of means and confidence intervals from posterior distributions.
+# Puts both male and female results in the same plot.
+create_double_forest_plot <- function(male_posterior, female_posterior, variables, combined_variables, base_size = 11) {
+  result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
   cis <- result
-  #result <- DescTools::Rev(result, margin=1)   # Reverse the order of rows so they appear in correct order in the forest plot
-  print(head(result))
-  #combined_variables <- combined_variables %>% 
-    #rename(names=Pretty) %>%
-    #mutate(stripe=factor(row_number()%%2))
-  result <- left_join(combined_variables, result, by="Pretty") %>%
+
+  result <- result %>%
     mutate(Pretty=fct_rev(Pretty))
 
-    #mutate(Pretty=factor(Pretty, levels=rev(combined_variables$Pretty)))
-    # group_by(gender) %>%
-    # mutate(stripe=factor(row_number()%%2)) %>%
-    # ungroup()
-  df1 <<- result
   plot <- result %>% 
-    # mutate(Pretty=factor(names, levels=rev(combined_variables$names)), 
-    #        stripe=factor(row_number()%%2)) %>%
     ggplot() +     
     geom_vline(aes(xintercept=0), color="lightgray") +
-    #geom_pointrange() +
     ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
@@
 create_variable_importance_plot2 <- function(rrfFit_roc, descript, combined_variables, base_size = 11) {
-  rrfFit_rocImp <- varImp(rrfFit_roc, scale = FALSE)
-  rrfFit.varimp <- as_tibble(cbind(rownames(rrfFit_rocImp$importance),rrfFit_rocImp$importance))
-  colnames(rrfFit.varimp) <- c("Variable","Importance")
-  rrfFit.varimp <- left_join(rrfFit.varimp, descript, by=c("Variable"="Variable")) %>% select(Variable,Pretty,Importance) %>% arrange(Importance)
-
-  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "previous_Hb_defTRUE"] <- "Previous donation deferred"
-  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "warm_seasonTRUE"] <- "Warm season"
-  rrfFit.varimp$Pretty[rrfFit.varimp$Variable == "genderWomen"] <- "Sex"
-
-  #pretty <- descript["Pretty"]
-  #pretty <- combined_variables
-  rrfFit.varimp <- rrfFit.varimp %>% mutate(Pretty=factor(Pretty, levels=levels(combined_variables$Pretty)))
-  rrfFit.varimp <- left_join(combined_variables, rrfFit.varimp, by="Pretty")
+  rrfFit.varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
 
   var_imp <- rrfFit.varimp %>%
-      mutate(Pretty=fct_rev(Pretty),
-             gender=factor("both", levels=c("female", "both", "male"))) 
-             #stripe=factor(row_number()%%2))
-  print(var_imp) 
-  df2 <<- var_imp
+      mutate(Pretty=fct_rev(Pretty))
+
   varimp.plot <- var_imp %>% 
-      # rrfFit.varimp %>%
-      # mutate(Pretty=factor(Pretty, levels=rev(Pretty)), 
-      #        stripe=factor(row_number()%%2)) %>%
     ggplot() + 
@@
   rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
+
   combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
-    mutate(stripe=factor(row_number()%%2),
+    mutate(stripe=factor(row_number()%%2),    # this is used to make the gray/white background
            Pretty=factor(Pretty, levels=Pretty))
-  combined_variables <<- combined_variables
+
   forest <- create_double_forest_plot(male_posterior, female_posterior, variables, combined_variables, base_size = base_size)$plot
@@
 # Give a more descriptive name to this variable
-variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred") #"Previous donation deferred", #"Previous Hb deferral"
+variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
 ```
@@
 ```
+## Alternative effect sizes and importances of variables
+
+```{r}
+alternative_create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
+                                           filename=NULL,
+                                           width = 180,  # width of the combined figure in millimetres
+                                          base_size = 11
+) {
+  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
+
+  combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
+    mutate(stripe=factor(row_number()%%2),    # this is used to make the gray/white background
+           Pretty=factor(Pretty, levels=Pretty))
+
+  result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
+  varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
+
+  result <- result %>%
+    mutate(Pretty=fct_rev(Pretty))
+
+  var_imp <- varimp %>%
+    mutate(Pretty=fct_rev(Pretty),
+           mean=Importance,
+           low=0,
+           high=Importance)
+
+  final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
+    mutate(method=factor(method, levels=c("Regression coefficient", "Importance")))
+
+  plot <- final %>% 
+    ggplot() +     
+    geom_vline(aes(xintercept=0), color="lightgray") +
+    ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
+                               position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
+    geom_rect(data=combined_variables %>% filter(stripe==1), 
+              mapping=aes(ymax = as.numeric(Pretty) + 0.5,
+                          ymin = as.numeric(Pretty) - 0.5),
+              fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA) +
+    labs(title=NULL,
+         x=NULL, y=NULL, colour="Sex") +
+    scale_colour_discrete(drop=FALSE, breaks = c("male", "female", "both")) + 
+    facet_wrap("method", scales="free_x") +
+    theme_classic(base_size = base_size) + 
+    theme(legend.position = "bottom", legend.direction = "horizontal", strip.background = element_rect(fill = NULL, color = "white", size = 1))
+
+#    ggstance::geom_pointrangeh(aes(y=Pretty, x=Importance, xmin=0, xmax=Importance, colour = gender), # ggstance is required to make legend keys horizontal
+#                               position=position_dodge2(width=0.8, padding=0.1), size=0.25, inherit.aes = FALSE) +
+
+    #inherit.aes = FALSE) +
+#    guides(y="none") +
+
+  return(list(plot=plot, final=final))
+}
+```
+
+```{r}
+g <- alternative_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
+
+if (save_figs)
+  ggsave(filename="alternative_effect_size_importance.pdf", title="Performance forest plot", path=fig_path, plot=g, dpi=600, units="mm", width=180)
+g
+```
 
@@
 ```{r}
+
+get_threshold_df <- function() {
+  t6 <- tibble(
+      Id = factor(ids),
+      type = factor("Threshold"),
+      value = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1, 0.4, 0.6))
+  t12 <- tibble(
+      Id = factor(ids),
+      type = factor("Threshold"),
+      value = c(0.2, 0.2, 0.1, 0.1, 0.1, 0.2, 0.3, 0.3, 0.8, 0.8)
+  )
+  t <- bind_rows(t6, t12)
+  t <- t %>%
+    mutate(low=value, 
+           high=value,
+           sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
+           model=factor(str_split_fixed(Id, "-", 3)[,3]))
+  return(t)
+}
+
 # Show AUROC, AUPR, F1, E6, and E12 as forest plot
@@
 
+  thresholds <- get_threshold_df()
+  df <- bind_rows(df, thresholds)
+
   # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
@@
 
-  #key_size <- as.numeric(grid::convertX(theme_get()$legend.key.size, "mm"))
-  key_size <- 1.8
+  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
+  #print(key_size)
+  key_size <- 1.8   # This results in box of size 6 mm
+  #key_size <- unit(6.096, "mm")
   #key_size <- unit(1.2, "lines")
+  #grid::convertX(grid::unit(1.2, "lines"), "mm")
   g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
@@
     guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size))) +
+    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
+    #theme(legend.key.size = theme_get()$legend.key.size) +
     geom_blank(data=dummy) +

### 2020-11-04 src/article_figures.Rmd
Polished article plots

@@
     geom_vline(aes(xintercept=0), color="lightgray") +
-    ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
-                               position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
     geom_rect(data=combined_variables %>% filter(stripe==1), 
@@
                   fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA) +
+    ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
+                               position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
     labs(title="", #title="Effects sizes of variables on Hb prediction",
@@
     scale_colour_discrete(drop=FALSE, breaks = c("male", "female")) + 
+    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
+                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
     #guides(colour = guide_legend(reverse = TRUE)) + # Reverse the order of keys in the legend
@@
     ggplot() + 
-    ggstance::geom_pointrangeh(aes(y=Pretty, x=Importance, xmin=0, xmax=Importance, colour = gender), # ggstance is required to make legend keys horizontal
-                               position=position_dodge2(width=0.8, padding=0.1), size=0.25, inherit.aes = FALSE) +   
-                               #colour = hue_pal()(3)[2], size=0.25) +
     geom_rect(data=combined_variables %>% filter(stripe==1),
@@
               xmin=-Inf, xmax=Inf, alpha = 0.1, colour=NA, fill="gray", show.legend = FALSE, inherit.aes = FALSE) +
+    ggstance::geom_pointrangeh(aes(y=Pretty, x=Importance, xmin=0, xmax=Importance, colour = gender), # ggstance is required to make legend keys horizontal
+                               position=position_dodge2(width=0.8, padding=0.1), size=0.25, inherit.aes = FALSE) +   
+                               #colour = hue_pal()(3)[2], size=0.25) +
     labs(y=NULL, colour="Sex") + 
@@
     scale_colour_discrete(drop=FALSE, breaks=c("both")) +
+    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
+                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
     #theme_gray(base_size = base_size)
@@
 ```
+
 ## Alternative effect sizes and importances of variables
@@
            mean=Importance,
-           low=0,
+           low=ifelse(is.na(Importance), NA, 0),
            high=Importance)
@@
   final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
-    mutate(method=factor(method, levels=c("Regression coefficient", "Importance")))
+    mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
+    select(-Importance, -Variable, stripe)
+
+  print(combined_variables)
 
@@
     geom_vline(aes(xintercept=0), color="lightgray") +
-    ggstance::geom_pointrangeh(aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
-                               position=position_dodge2(width=0.8, padding=0.1), size=0.25) + # ggstance is required to make legend keys horizontal
+    # ggstance is required to make legend keys horizontal
     geom_rect(data=combined_variables %>% filter(stripe==1),
@@
                           ymin = as.numeric(Pretty) - 0.5),
-              fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA) +
-    labs(title=NULL,
-         x=NULL, y=NULL, colour="Sex") +
+              fill = "gray", xmin=-Inf, xmax=Inf, alpha = 0.1, show.legend = FALSE, colour=NA, inherit.aes = FALSE) +
+    ggstance::geom_pointrangeh(mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high, color=gender),
+                               position=position_dodge2(width=0.8, padding=0.1), 
+                               size=0.25,  
+                               inherit.aes = FALSE) +
+    labs(title=NULL, x=NULL, y=NULL, colour="Sex") +
     scale_colour_discrete(drop=FALSE, breaks = c("male", "female", "both")) + 
+    scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
+                         # Otherwise following error results: Error: Discrete value supplied to continuous scale
     facet_wrap("method", scales="free_x") +
@@
     theme(legend.position = "bottom", legend.direction = "horizontal", strip.background = element_rect(fill = NULL, color = "white", size = 1))
-
-#    ggstance::geom_pointrangeh(aes(y=Pretty, x=Importance, xmin=0, xmax=Importance, colour = gender), # ggstance is required to make legend keys horizontal
-#                               position=position_dodge2(width=0.8, padding=0.1), size=0.25, inherit.aes = FALSE) +
-
-    #inherit.aes = FALSE) +
-#    guides(y="none") +
 
@@
 ```{r}
-g <- alternative_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
-
+res <- alternative_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
+g <- res$plot
+final <- res$final
 if (save_figs)
@@
            sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
-           model=factor(str_split_fixed(Id, "-", 3)[,3]))
+           #model=factor(str_split_fixed(Id, "-", 3)[,3])
+           model=rep(c("t6", "t12"), each=length(ids))
+           )
   return(t)
@@
   df <- bind_rows(df, thresholds)
-
+  df <- df %>% 
+    mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "t6", "t12")))
   # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
@@
     scale_colour_discrete(breaks = c("male", "female", "both")) +
-    scale_shape_discrete(breaks=c("lmm", "dlmm", "dt", "rf")) +
+    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf"), values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
     #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
@@
   #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
-  return(g)
+  return(list(g=g, df=df))
 }
@@
 ```{r}
-g <- create_performance_forest_plot()
+res <- create_performance_forest_plot()
+g <- res$g
+df <- res$df
 if (save_figs)
@@
 ```
+
 ## Calibration plots
@@
 ```
+
+```{r, eval=FALSE}
+jarkko_test <- function(male_posterior, female_posterior, variables,
+                                           filename=NULL,
+                                           width = 180,  # width of the combined figure in millimetres
+                                          base_size = 11
+) {
+  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
+
+  combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
+    mutate(stripe=factor(row_number()%%2),    # this is used to make the gray/white background
+           Pretty=factor(Pretty, levels=Pretty))
+
+  result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
+  varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
+
+  result <- result %>%
+    mutate(Pretty=fct_rev(Pretty))
+
+  var_imp <- varimp %>%
+    mutate(Pretty=fct_rev(Pretty),
+           mean=Importance,
+           low=ifelse(is.na(Importance), NA, 0),
+           high=Importance)
+
+  final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
+    mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
+    select(-Importance, -Variable, stripe)
+
+#   final <- tibble(
+#             method = c("Regression coefficient",
+#                        "Regression coefficient","Importance"),
+#             Pretty = factor(c("Days to previous full blood donation","Days to previous full blood donation",
+#                        "Days to previous full blood donation")),
+#             #gender = c("male", "female", "both"),
+#               mean = c(0.0152091681221149,
+#                        0.0445741103338152,0.0255604042298423),
+#                low = c(0.00691102953212081, 0.0353192308283623, 0),
+#               high = c(0.0240855557569598,
+#                        0.0537929726515129,0.0255604042298423)
+# )
+
+  #combined_variables <- tibble(Pretty=factor("Days to previous full blood donation"), stripe=factor(1))
+
+  print(final)         
+  print(combined_variables)
+  vars <- c("Days to previous full blood donation", "Age", "Previous donation deferred", "Year")
+  plot <- final %>% #filter(Pretty %in% vars) %>%
+    ggplot(aes(color=gender)) +
+    geom_rect(data=combined_variables %>% filter(stripe==1),
+              mapping=aes(#ymax = as.numeric(Pretty) + 0.5,
+                          #ymin = as.numeric(Pretty) - 0.5
+                          ymax = as.numeric(Pretty) + 0.5,
+                          ymin = as.numeric(Pretty) - 0.5),
+              fill = "blue", 
+              #position="identity",
+              xmin=-Inf, xmax=Inf, alpha = 0.1, inherit.aes = FALSE) +
+    ggstance::geom_pointrangeh(#data=final, 
+                               mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high),
+                               #position=position_dodge2(0.8)
+                               position=position_dodge2(width=0.8, padding=0.1)
+                               ) +
+    scale_y_discrete() +
+    facet_wrap("method", scales = "free_x") +
+    #coord_flip() +
+    theme_classic(base_size = base_size)
+
+  return(list(plot=plot, final=final))
+}
+
+res <- jarkko_test(male_posterior, female_posterior, variables, filename=filename)
+final <- res$final
+g <- res$plot
+g
+```

### 2020-11-05 src/economic_effect_functions.R
Added dummy classifiers for both genders

@@ -9,6 +9,7 @@ library(boot)
 ids <- c("progesa-male-lmm", "progesa-female-lmm", "progesa-male-dlmm", "progesa-female-dlmm", "finngen-male-dlmm", "finngen-female-dlmm", 
          "findonor-male-dlmm", "findonor-female-dlmm", "progesa-both-dt", "progesa-both-rf")
 
+dummy2_ids <- c("progesa-female-dummy", "progesa-male-dummy")
 dummy_ids <- c("finngen-male-stratified", "finngen-male-most-frequent", "finngen-male-prior", "finngen-male-uniform", "finngen-male-deferred")
 
 # These are from Progesa data
@@ -192,8 +193,8 @@ get_cost <- function(TPR6, FPR6, TPR12, FPR12, p = parameters) {
   }
 }
 
-get_optimal_thresholds <- function(df, p=parameters) {
-  thresholds <- seq(0.1, .9, .1)
+get_optimal_thresholds <- function(df, p=parameters, thresholds = seq(0.1, .9, .1)) {
+  
   pred.classes <- data.frame(matrix(nrow=nrow(df),ncol=length(thresholds)))
   colnames(pred.classes) <- paste0("p_",thresholds)
   for (t in thresholds ) {
@@ -295,6 +296,12 @@ get_data_frame <- function(id) {
     df <- as_tibble(load_single("~/FRCBS/interval-sims/rrfFit__dtree_roc_validate_probs.rdata"))
     df <- df %>% mutate(obs = factor(ifelse(obs == 2, "Deferred", "Accepted"), levels=c("Accepted", "Deferred")))
     return(df)
+  } else if (id == "progesa-female-dummy") {
+    df <- load_single("~/FRCBS/interval-sims/progesa-validate-female-dummy.rdata")
+    return(df)
+  } else if (id == "progesa-male-dummy") {
+    df <- load_single("~/FRCBS/interval-sims/progesa-validate-male-dummy.rdata")
+    return(df)
   } else if (id == "finngen-male-stratified") {
     df <- get_data_frame("finngen-male-dlmm")
     n <- nrow(df)
@@ -321,7 +328,7 @@ get_data_frame <- function(id) {
     df <- get_data_frame("finngen-male-dlmm")
     n <- nrow(df)
     df <- df %>% mutate(Deferred = 1.0)
-  } else {
+  } else {         # Linear models
     df <- get_raw_result_list(id)$comp_df
     return(process(df))
   }
[…]

### 2020-11-06 src/article_figures.Rmd
Updated article figures

@@
 source("economic_effect_functions.R")
-save_figs = FALSE
+save_figs = TRUE
 fig_path="~/FRCBS/results-for-eba-paper/pdf"
@@
 
-```{r}
+```{r, include=FALSE}
 
@@
 
-```{r}
+```{r, include=FALSE}
 finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
@@
 
-```{r}
+```{r, include=FALSE}
 alternative_create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
@@
     labs(title=NULL, x=NULL, y=NULL, colour="Sex") +
-    scale_colour_discrete(drop=FALSE, breaks = c("male", "female", "both")) + 
+    scale_colour_discrete(drop=FALSE, breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) + 
     scale_y_discrete() + # !!!!! This is important. Solves the problem with position_dodge2 and the order of rect and pointrange geoms !!!!!!
@@
     theme_classic(base_size = base_size) + 
-    theme(legend.position = "bottom", legend.direction = "horizontal", strip.background = element_rect(fill = NULL, color = "white", size = 1))
+    theme(legend.position = "bottom", legend.direction = "horizontal", strip.background = element_rect(fill = NULL, color = "white", size = 1))#,
+          #plot.background = element_rect(fill = "lightblue"))
 
@@
 
+get_name_of_data <- function(s) {
+  s <- str_split_fixed(s, "-", 3)[,1]
+  s <- map(s, function(x) when(x, 
+                               x == "progesa" ~ "eProgesa",
+                               x == "finngen" ~ "Biobank",
+                               x == "findonor" ~ "FinDonor",
+                               ~ "koe"))
+  return(s)
+}
+
 # Show AUROC, AUPR, F1, E6, and E12 as forest plot
@@
   df <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
+  new_names <- c(`Economic effect of 6 month deferral` = "E6", `Economic effect of 12 month deferral` = "E12")
   df <- df %>%
@@
            sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
-           model=factor(str_split_fixed(Id, "-", 3)[,3]))
+           model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
+    mutate(type = fct_recode(type, !!!new_names))
 
@@
   dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, type=mytype) # Dummy table to use the same x-axis limits for E6 and E12
-  #print(dummy)
+
+  dummy <- dummy %>%
+    mutate(type = fct_recode(type, !!!new_names))
+
+  print(df)
 
@@
   g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
-    ggstance::geom_pointrangeh(aes(colour=sex, shape=model)) + 
+    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
     #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
-    labs(x="Value", y="Data and method", colour="Sex", shape="Model") + 
-    #scale_colour_manual(breaks = c("male", "female", "both"), values=c("yellow", "brown", "black"), aesthetics = c("color", "fill", "shape")) +
-    scale_colour_discrete(breaks = c("male", "female", "both")) +
-    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf"), values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
+    labs(x="Value", y="Data and method", colour="Sex", shape="Model", fill="Deferral length") + 
+    scale_y_discrete(labels=get_name_of_data) +
+    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
+    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:6) + #breaks = new_names) +
+    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
+                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
+                       values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
+    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
+    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
+    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
     #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
-    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size))) +
+    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
+           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
     #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
@@
     geom_blank(data=dummy) +
-    facet_wrap("type", scales="free_x")
+    facet_wrap("type", scales="free_x") +
+    #theme(legend.position = "right", legend.direction = "vertical") 
+    theme(legend.position = "bottom", legend.direction = "vertical")
   #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
@@
 
-```{r calibration plots}
+```{r calibration plots, include=FALSE}
 calibration_plot <- function(df, use_my_method=FALSE) {
@@
   xymax <- max(max(df$predicted), max(df$observed))
+  if (is.null(probability_of_deferral_threshold)) { # predict deferral based on Hb
     df <- df %>% 
-    mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0)) %>%
+      mutate(new_predicted_label= ifelse(predicted < hb_threshold, 1, 0))
+  } else { # predict deferral based on probability of deferral
+    df <- df %>% 
+      mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0))
+  }
+  df <- df %>%
     mutate(confusion_class = factor(ifelse(deferral == 1, 
@@
     scale_y_continuous(breaks = generate_my_breaks(20), limits=c(xymin,xymax)) +
+    scale_colour_discrete(drop=FALSE) +
     geom_abline(intercept = 0, slope = 1) +
@@
     geom_hline(yintercept = hb_threshold, linetype = "dashed") +
-    theme(legend.position = "bottom") +
-    ggtitle("Observed vs predicted Hb-values")
+    theme(legend.position = "bottom")
+    #ggtitle("Observed vs predicted Hb-values")
   return(scatter_plot)
@@
 ```{r}
-g <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.1)
-if (save_figs)
-  ggsave(filename="classification-scatter-plot-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
-         plot=g, dpi=600, units="mm", width=180)
-g
+g1 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, NULL)
+g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.1)
+if (save_figs) {
+  ggsave(filename="classification-scatter-plot-hb-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
+         plot=g1, dpi=600, units="mm", width=180)
+  ggsave(filename="classification-scatter-plot-probability-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
+         plot=g2, dpi=600, units="mm", width=180)
+}
+g1
+g2
 ```
 
-```{r, eval=FALSE}
+```{r, eval=FALSE, include=FALSE}
 jarkko_test <- function(male_posterior, female_posterior, variables,

### 2020-11-06 src/dummy_classifier.Rmd
Added dummy predictor

---
title: "Dummy classifier"
author: "Jarkko Toivonen"
date: "11/4/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("economic_effect_functions.R")
```

## Dummy model for females and males on Progesa data

Use the preprocessed validate dataset from random forest runs. Dummy predictor uses -previous_Hb as score for deferral.

```{r}
load("~/FRCBS/interval-sims/rf-train-and-validate-datas.rdata", verbose=TRUE)
```

```{r cars}

logistic <- function(x) 1 / (1 + exp(-x))

dummy_predictor <- function(df, Hb_threshold=NULL) {
  if (is.null(Hb_threshold)) {
    result <- tibble(Deferred = -df$previous_Hb,
                     obs = df$Hb_deferral)
  } else {
    result <- tibble(Deferred = logistic(-(df$previous_Hb - Hb_threshold)), 
                     obs = df$Hb_deferral)

  }
  return(result)
}

# get_metrics <- function(df, previous_Hb_threshold) {
#   df <- df %>% mutate(predicted_Hb_deferral=factor(ifelse(previous_Hb < previous_Hb_threshold, "Deferred", "Accepted"), levels=c("Accepted", "Deferred")))
#   
#   cm <- confusionMatrix(reference = df$Hb_deferral, data = df$predicted_Hb_deferral, positive = "Deferred", mode = "everything")
#   f1 <- cm$byClass["F1"]
#          #Sensitivity  = TPR
#   TPR <- cm$byClass['Sensitivity']
#   FPR <- 1 - cm$byClass['Specificity']
#   temp <- get_cost(TPR, FPR, TPR, FPR, p = parameters) # The parameters are from file economic_effect_functions.R
#   E6  <- temp$E6
#   E12 <- temp$E12
#   return(tibble(previous_Hb_threshold=previous_Hb_threshold, f1=f1, E6=E6, E12=E12))
# }
# 
# optimize_threshold <- function(df) {
#   thresholds <- seq(120, 150, by=2)
#   result <- map_dfr(thresholds, function(t) get_metrics(df, t))
#   
#   threshold <- result %>%
#     slice_max(f1) %>%
#     pull(previous_Hb_threshold)
#   return(list(threshold=threshold, result=result))
# }

```

```{r}
train_female <- train %>% filter(gender=="Women")
train_male <- train %>% filter(gender=="Men")
validate_female <- validate %>% filter(gender=="Women")
validate_male <- validate %>% filter(gender=="Men")
```

```{r}
# res_female <- optimize_threshold(train_female)
# res_male <- optimize_threshold(train_male)
```

Optimized thresholds on previous Hb for females and males
```{r}
# female_previous_Hb_threshold <- res_female$threshold
# male_previous_Hb_threshold <- res_male$threshold
# female_previous_Hb_threshold
# male_previous_Hb_threshold
```

```{r}
# res_female$result
# res_male$result
```

Final result on validate data
```{r}

# f1_female <- get_f1(validate_female, female_previous_Hb_threshold)
# f1_female
```

```{r}

# f1_male <- get_f1(validate_male, male_previous_Hb_threshold)
# f1_male
```

```{r}
prediction_female <- dummy_predictor(validate_female, 125)  # or would mean Hb be better centering point than deferral threshold
prediction_male <- dummy_predictor(validate_male, 135)

head(prediction_female)
head(prediction_male)
```

```{r}
save(prediction_female, file="~/FRCBS/interval-sims/progesa-validate-female-dummy.rdata")
save(prediction_male, file="~/FRCBS/interval-sims/progesa-validate-male-dummy.rdata")
```

```{r}
dummy_result <- process_all_data(dummy2_ids)
```

```{r}
dummy_result
```

### 2020-11-06 src/economic_effect_functions.R

@@ -477,9 +477,9 @@ get_intervals <- function(df, f1_threshold, threshold6=0.6, threshold12=0.8) {
 }
 
 process_data <- function(df, id) {
-  print("Here")
+  #print("Here")
   if (id %in% c("progesa-female-dummy", "progesa-male-dummy")) {
-    thresholds <-  seq(-150, -120, by=2)   # thresholds for inverse of hemoglobin
+    thresholds <-  seq(-160, -110, by=2)   # thresholds for inverse of hemoglobin
     f1_threshold <- ifelse(id=="progesa-female-dummy", -125, -135)
   } else {
     thresholds <- seq(0.1, .9, .1)  # thresholds for probability of deferral
@@ -501,7 +501,8 @@ process_all_data <- function(ids) {
     mutate(type=map_chr(type, function(x) when(x, . == "actual" ~ "", . == "CI_inf" ~ "low", . == "CI_sup" ~ "high"))) %>%
     pivot_wider(names_from=c("variable", "type"), values_from=value, names_glue="{variable} {type}")
   names(df) <- str_trim(names(df))
-  df <- df %>% select("Id", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high", "F1", "F1 low", "F1 high", "threshold6", "threshold12")
+  #df <- df %>% select("Id", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high", "F1", "F1 low", "F1 high", "threshold6", "threshold12")
+  df <- df %>% select("Id", "F1", "F1 low", "F1 high", "threshold6", "threshold12", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high")
   return(df)
 }
 

### 2020-11-09 src/article_figures.Rmd
Added dummy models to the performance plot

@@
 source("economic_effect_functions.R")
-save_figs = TRUE
+save_figs = FALSE
 fig_path="~/FRCBS/results-for-eba-paper/pdf"
@@
 
+all_ids <- c(ids, dummy2_ids)
+
 get_threshold_df <- function() {
+
   t6 <- tibble(
-      Id = factor(ids),
+      Id = factor(all_ids),
       type = factor("Threshold"),
-      value = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1, 0.4, 0.6))
+      value = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1, 0.4, 0.6,   0.2, 0.3))
   t12 <- tibble(
-      Id = factor(ids),
+      Id = factor(all_ids),
       type = factor("Threshold"),
-      value = c(0.2, 0.2, 0.1, 0.1, 0.1, 0.2, 0.3, 0.3, 0.8, 0.8)
+      value = c(0.2, 0.2, 0.1, 0.1, 0.1, 0.2, 0.3, 0.3, 0.8, 0.8,   0.4, 0.5)
   )
@@
            #model=factor(str_split_fixed(Id, "-", 3)[,3])
-           model=rep(c("t6", "t12"), each=length(ids))
+           model=rep(c("t6", "t12"), each=length(all_ids))
            )
@@
 create_performance_forest_plot <- function() {
-  df <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
-  new_names <- c(`Economic effect of 6 month deferral` = "E6", `Economic effect of 12 month deferral` = "E12")
+  df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
+  df2 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data_dummy.csv")
+  df <- bind_rows(df1, df2)
+  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")
   df <- df %>%
-    mutate(Id=factor(Id, levels=rev(ids)),
+    mutate(Id=factor(Id, levels=rev(all_ids)),
            type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
@@
   df <- df %>% 
-    mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "t6", "t12")))
+    mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
+  write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")
+
   # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
-  xrange <- as.numeric(df %>% filter(type %in% c("E6", "E12")) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
+  xrange <- as.numeric(df %>% filter(type %in% names(new_names)) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
   xrange <- rep(xrange, 2)
@@
     mutate(type = fct_recode(type, !!!new_names))
-
+  print(dummy)
   print(df)
@@
     scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
-    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:6) + #breaks = new_names) +
-    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
-                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
-                       values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
+    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
+    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "dummy"),
+                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Dummy"),
+                       values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
     # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
@@
 
+Testing the effect of base_size parameter.
+
+```{r}
+g1 + theme_gray(base_size=20)
+g2 + theme_gray(base_size=20)
+```
+
 ```{r, eval=FALSE, include=FALSE}

### 2020-11-09 src/dummy_classifier.Rmd
Use logistic function in the dummy classifier

@@
 date: "11/4/2020"
-output: html_document
+output: md_document
 ---
@@
 
+library(tidyverse)
 source("economic_effect_functions.R")
+source('../../Hb_predictor_container/src/validate_stan_fit.R', chdir=T)
 ```
@@
 
+We use [Logistic function](https://en.wikipedia.org/wiki/Logistic_function) to map previous_Hb values to
+the probability range [0, 1].
+
 ```{r cars}
 
-logistic <- function(x) 1 / (1 + exp(-x))
+logistic <- function(x, k=1, x0=0) 1 / (1 + exp(-k*(x-x0)))
+
+score <- function(previous_Hb, Hb_threshold ) {
+  logistic(previous_Hb, k=-0.1, x0=Hb_threshold)
+}
+
+make_threshold_conversion_data_frame <- function(Hb_threshold, k) {
+  hb <- seq(110, 160, by=2)
+  df <- tibble(Hb=hb, 
+               Probability = logistic(hb, k=k, x0=Hb_threshold))
+  return(df)
+}
 
 dummy_predictor <- function(df, Hb_threshold=NULL) {
-  if (is.null(Hb_threshold)) {
+  if (is.null(Hb_threshold)) {    # use -previous_Hb as score
     result <- tibble(Deferred = -df$previous_Hb,
                      obs = df$Hb_deferral)
-  } else {
-    result <- tibble(Deferred = logistic(-(df$previous_Hb - Hb_threshold)), 
+  } else {                        # use logistic function of previous_Hb as score
+    result <- tibble(Deferred = score(df$previous_Hb, Hb_threshold), 
                      obs = df$Hb_deferral)
@@
 
+```
+
+```{r}
+df1 <- make_threshold_conversion_data_frame(125, -1) 
+df2 <- make_threshold_conversion_data_frame(125, -0.1)
+df <- bind_rows(`k=-1`=df1, `k=-0.1`= df2, .id="k")
+df %>% ggplot(aes(x=Hb, y=Probability, color=k)) + geom_point() + scale_y_continuous(breaks=seq(0, 1, by=0.1)) + labs(colour="Logistic function parameter") + ggtitle("Female")
+```
+
+```{r}
+df1 <- make_threshold_conversion_data_frame(135, -1) 
+df2 <- make_threshold_conversion_data_frame(135, -0.1)
+df <- bind_rows(`k=-1`=df1, `k=-0.1`= df2, .id="k")
+df %>% ggplot(aes(x=Hb, y=Probability, color=k)) + geom_point() + scale_y_continuous(breaks=seq(0, 1, by=0.1)) + labs(colour="Logistic function parameter") + ggtitle("Male")
 ```
@@
 ```
+
+```{r}
+roc_female <- roc_wrapper(prediction_female)
+pr_female <- pr_wrapper(prediction_female)
+```
+
+```{r}
+roc_male <- roc_wrapper(prediction_male)
+pr_male <- pr_wrapper(prediction_male)
+```
+
+```{r}
+v <- c(roc_female$roc$ci[c(2,1,3)], pr_female$ci[c(1,2,3)], 
+       roc_male$roc$ci[c(2,1,3)], pr_male$ci[c(1,2,3)])
+aucs <- as_tibble(matrix(v, ncol=6, byrow=TRUE))
+colnames(aucs) <- c("AUROC", "AUROC low", "AUROC high", "AUPR value", "AUPR low", "AUPR high")
+aucs <- aucs %>% 
+  mutate(Id=dummy2_ids) %>%
+  select(Id, everything())
+
+aucs
+```
+
+```{r}
+dummy_result <- inner_join(aucs, dummy_result, by="Id")
+dummy_result
+```

### 2020-11-09 src/economic_effect_functions.R

@@ -9,7 +9,7 @@ library(boot)
 ids <- c("progesa-male-lmm", "progesa-female-lmm", "progesa-male-dlmm", "progesa-female-dlmm", "finngen-male-dlmm", "finngen-female-dlmm", 
          "findonor-male-dlmm", "findonor-female-dlmm", "progesa-both-dt", "progesa-both-rf")
 
-dummy2_ids <- c("progesa-female-dummy", "progesa-male-dummy")
+dummy2_ids <- c("progesa-male-dummy", "progesa-female-dummy")
 dummy_ids <- c("finngen-male-stratified", "finngen-male-most-frequent", "finngen-male-prior", "finngen-male-uniform", "finngen-male-deferred")
 
 # These are from Progesa data
@@ -404,7 +404,7 @@ test_get_cost <- function(name, threshold=0.5) {
 
 get_f1 <- function(df, threshold = 0.5) {
   pred.class <- factor( ifelse(df$Deferred >= threshold, "Deferred", "Accepted"), levels=c("Accepted", "Deferred"))
-  cm <- confusionMatrix(reference = df$obs, data = pred.class, positive = "Deferred", mode = "prec_recall")
+  cm <- caret::confusionMatrix(reference = df$obs, data = pred.class, positive = "Deferred", mode = "prec_recall")
   f1 <- cm$byClass["F1"]
   return(f1)
 }
@@ -467,8 +467,8 @@ compute_CI <-function(fit_boot){
 }
 
 get_intervals <- function(df, f1_threshold, threshold6=0.6, threshold12=0.8) {
-  nb_boot <- 100
-  #nb_boot <- nrow(df)
+  #nb_boot <- 100
+  nb_boot <- nrow(df)
   fit_boot <- boot(df, f1_threshold=f1_threshold, threshold6=threshold6, threshold12=threshold12, p=parameters, 
                    statistic = boot_cost, R = nb_boot, parallel="multicore", ncpus=4)
   ci <- compute_CI(fit_boot)
@@ -478,7 +478,7 @@ get_intervals <- function(df, f1_threshold, threshold6=0.6, threshold12=0.8) {
 
 process_data <- function(df, id) {
   #print("Here")
-  if (id %in% c("progesa-female-dummy", "progesa-male-dummy")) {
+  if (FALSE && id %in% c("progesa-female-dummy", "progesa-male-dummy")) {
[…]

### 2020-12-01 src/dummy_classifier.Rmd
Dummy model is now logistic regression, and predicts test data

@@
 ```{r}
-load("~/FRCBS/interval-sims/rf-train-and-validate-datas.rdata", verbose=TRUE)
+load("~/FRCBS/interval-sims/rf-train-and-validate-datas.rdata", verbose=TRUE)  # returns train and validate dataframes
+load("~/FRCBS/interval-sims/rf-test-data.rdata", verbose=TRUE) # returns test dataframe
 ```
@@
 
-dummy_predictor <- function(df, Hb_threshold=NULL) {
+dummy_predictor <- function(df, Hb_threshold=NULL) { # fake logistic regression
   if (is.null(Hb_threshold)) {    # use -previous_Hb as score
@@
   }
+  return(result)
+}
+
+dummy_predictor2 <- function(train, validate) {  # real logistic regression
+  glm.fit <- glm(Hb_deferral ~ previous_Hb,
+               data = train,
+               family = binomial)
+
+  print(glm.fit)
+
+  glm.probs <- predict(glm.fit,
+                      newdata = validate,
+                      type = "response")
+
+  result <- tibble(Deferred = glm.probs, 
+                   obs = validate$Hb_deferral)
+
   return(result)  
@@
 validate_male <- validate %>% filter(gender=="Men")
+test_female <- test %>% filter(gender=="Women")
+test_male <- test %>% filter(gender=="Men")
 ```
@@
 ```{r}
+use_logistic_regression=TRUE
+if (use_logistic_regression) {
+  prediction_female <- dummy_predictor2(train_female, test_female)  # or would mean Hb be better centering point than deferral threshold
+  prediction_male <- dummy_predictor2(train_male, test_male)
+} else {
   prediction_female <- dummy_predictor(validate_female, 125)  # or would mean Hb be better centering point than deferral threshold
   prediction_male <- dummy_predictor(validate_male, 135)
-
+}
 head(prediction_female)
@@
 ```{r}
-save(prediction_female, file="~/FRCBS/interval-sims/progesa-validate-female-dummy.rdata")
-save(prediction_male, file="~/FRCBS/interval-sims/progesa-validate-male-dummy.rdata")
+save(prediction_female, file="~/FRCBS/interval-sims/progesa-validate-female-dummy2.rdata")
+save(prediction_male, file="~/FRCBS/interval-sims/progesa-validate-male-dummy2.rdata")
 ```
@@
 ```{r}
-dummy_result <- process_all_data(dummy2_ids)
+dummy_result <- process_all_data(dummy2_ids, replicates=100)
 ```
@@
 ```{r}
-dummy_result
+df <- dummy_result %>% 
+  pivot_wider(names_from=c("variable", "type"), values_from=value, names_glue="{variable} {type}")
+names(df) <- str_remove(names(df), " value$")
+df <- df %>%
+  select("Id", "F1", "F1 low", "F1 high", "threshold6", "threshold12", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high")
+
+df
 ```
@@
 ```
-
 ```{r}
-v <- c(roc_female$roc$ci[c(2,1,3)], pr_female$ci[c(1,2,3)], 
-       roc_male$roc$ci[c(2,1,3)], pr_male$ci[c(1,2,3)])
-aucs <- as_tibble(matrix(v, ncol=6, byrow=TRUE))
-colnames(aucs) <- c("AUROC", "AUROC low", "AUROC high", "AUPR value", "AUPR low", "AUPR high")
-aucs <- aucs %>% 
-  mutate(Id=dummy2_ids) %>%
-  select(Id, everything())
-
-aucs
+roc_female
+pr_female
+roc_male
+pr_male
 ```
@@
 ```{r}
-dummy_result <- inner_join(aucs, dummy_result, by="Id")
-dummy_result
+# v <- c(roc_female$roc$ci[c(2,1,3)], pr_female$ci[c(1,2,3)], 
+#        roc_male$roc$ci[c(2,1,3)], pr_male$ci[c(1,2,3)])
+# aucs <- as_tibble(matrix(v, ncol=6, byrow=TRUE))
+# colnames(aucs) <- c("AUROC", "AUROC low", "AUROC high", "AUPR value", "AUPR low", "AUPR high")
+# aucs <- aucs %>% 
+#   mutate(Id=dummy2_ids) %>%
+#   select(Id, everything())
+# 
+# aucs
 ```
+
+```{r}
+# dummy_result <- inner_join(aucs, dummy_result, by="Id")
+# dummy_result
+```

### 2020-12-01 src/economic_effect_functions.R
Many changes.
The decision tree and random forest know predict test data labels instead
of validate data's. The boot_cost now computes additional statistics
AUROC and AUPR. More automation and code cleanup. The resulting dataframe
from process_all_data contains all results in long form.

@@ -12,6 +12,8 @@ ids <- c("progesa-male-lmm", "progesa-female-lmm", "progesa-male-dlmm", "progesa
 dummy2_ids <- c("progesa-male-dummy", "progesa-female-dummy")
 dummy_ids <- c("finngen-male-stratified", "finngen-male-most-frequent", "finngen-male-prior", "finngen-male-uniform", "finngen-male-deferred")
 
+all_ids <- c(ids, dummy2_ids)
+
 # These are from Progesa data
 get_mikkos_cost_constants <- function() {
   return(list(
@@ -290,17 +292,19 @@ get_data_frame <- function(id) {
     df <- df %>% select(Deferred=fractions, obs=deferral) %>% mutate(obs=factor(ifelse(obs == 1, "Deferred", "Accepted"), levels=c("Accepted", "Deferred")))
     return(df)
   } else if (id == "progesa-both-rf") {
-    df <- load_single("~/FRCBS/interval-sims/rrfFit_roc_validate_probs.rdata")
+#    df <- load_single("~/FRCBS/interval-sims/rrfFit_roc_validate_probs.rdata")
+    df <- load_single("~/FRCBS/interval-sims/rrfFit_roc_test_probs.rdata")
     return(as_tibble(df))
   } else if (id == "progesa-both-dt") {
-    df <- as_tibble(load_single("~/FRCBS/interval-sims/rrfFit__dtree_roc_validate_probs.rdata"))
+    #df <- as_tibble(load_single("~/FRCBS/interval-sims/rrfFit__dtree_roc_validate_probs.rdata"))
+    df <- as_tibble(load_single("~/FRCBS/interval-sims/rrfFit__dtree_roc_test_probs.rdata"))
     df <- df %>% mutate(obs = factor(ifelse(obs == 2, "Deferred", "Accepted"), levels=c("Accepted", "Deferred")))
     return(df)
   } else if (id == "progesa-female-dummy") {
-    df <- load_single("~/FRCBS/interval-sims/progesa-validate-female-dummy.rdata")
+    df <- load_single("~/FRCBS/interval-sims/progesa-validate-female-dummy2.rdata")
     return(df)
   } else if (id == "progesa-male-dummy") {
-    df <- load_single("~/FRCBS/interval-sims/progesa-validate-male-dummy.rdata")
+    df <- load_single("~/FRCBS/interval-sims/progesa-validate-male-dummy2.rdata")
     return(df)
[…]

### 2020-12-01 src/article_figures.Rmd
Use the raw result dataframe as input for performance forest plot.

@@
 # Give a more descriptive name to this variable
-variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
+#variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
+variables <- recode(variables,    # Note that recode uses (stupidly) oldname = newname order, unlike many other functions in tidyverse
+                    `Previous Hb deferral` = "Previous donation deferred",
+                    `Polygenic risk score` = "Polygenic score",
+                    `RNF43 mutant` = "RNF43")
 ```
 
-```{r}
+```{r, eval=FALSE}
 if (save_figs) {
-    filename <- paste(fig_path, "effect_size_importance.pdf", sep="/")
+    filename <- paste(fig_path, "old-effect-size-importance.pdf", sep="/")
 } else {
@@
 
-  result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
-  varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
-
-  result <- result %>%
+  # Get effect sizes of the linear model  
+  dlmm <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
+  dlmm <- dlmm %>%
     mutate(Pretty=fct_rev(Pretty))
 
+  # Get variable importance from the RF object
+  varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
   var_imp <- varimp %>%
@@
 
-  final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
+  final <- bind_rows(`Regression coefficient`=dlmm, Importance=var_imp, .id="method") %>%
     mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
@@
 if (save_figs)
-  ggsave(filename="alternative_effect_size_importance.pdf", title="Performance forest plot", path=fig_path, plot=g, dpi=600, units="mm", width=180)
+  ggsave(filename="effect-size-importance.pdf", title="Effect sizes and importances", path=fig_path, plot=g, dpi=600, units="mm", width=180)
+  ggsave(filename="presentation-effect-size-importance.pdf", title="Effect sizes and importances", path=fig_path, plot=g, dpi=600, units="mm", width=180, height=90)
 g
@@
 get_name_of_data <- function(s) {
-  s <- str_split_fixed(s, "-", 3)[,1]
-  s <- map(s, function(x) when(x, 
-                               x == "progesa" ~ "eProgesa",
-                               x == "finngen" ~ "Biobank",
-                               x == "findonor" ~ "FinDonor",
-                               ~ "koe"))
+  s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
+  s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
+  # s <- map(s, function(x) when(x, 
+  #                              x == "progesa" ~ "eProgesa",
+  #                              x == "finngen" ~ "Biobank",
+  #                              x == "findonor" ~ "FinDonor",
+  #                              ~ "koe"))
   return(s)
@@
 create_performance_forest_plot <- function() {
+  new_names <- c(`€/donation, 6 month deferral` = "E6", `€/donation, 12 month deferral` = "E12")
+  if (FALSE) {
     df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
@@
     df <- bind_rows(df1, df2)
-  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")
+    #  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")
+
     df <- df %>%
@@
       mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
-  write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")
+    #write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")
+  } else {
+#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
+    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
+
+    df <- df %>% 
+      pivot_wider(names_from=c("type"), values_from=value) %>%
+      mutate(low = ifelse(is.na(low), value, low),     # To make the plot work, copy low=high=value for thresholds
+             high = ifelse(is.na(high), value, high))
+
+    df <- df %>%
+      mutate(Id=factor(Id, levels=rev(all_ids)),
+             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
+             model=case_when(variable == "threshold6" ~ "t6",
+                             variable == "threshold12" ~ "t12",
+                             TRUE ~ str_split_fixed(Id, "-", 3)[,3]),
+             model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")),
+             variable = recode(variable, "threshold6" = "Threshold", "threshold12" = "Threshold"),
+             variable = factor(variable, levels=c("AUROC", "AUPR", "F1", "E6", "E12", "Threshold")),
+             variable = fct_recode(variable, !!!new_names))
+
+  }
 
   # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
-  xrange <- as.numeric(df %>% filter(type %in% names(new_names)) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
+  xrange <- as.numeric(df %>% filter(variable %in% names(new_names)) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
   xrange <- rep(xrange, 2)
   mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
-  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, type=mytype) # Dummy table to use the same x-axis limits for E6 and E12
+  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, variable=mytype) # Dummy table to use the same x-axis limits for E6 and E12
 
   dummy <- dummy %>%
-    mutate(type = fct_recode(type, !!!new_names))
+    mutate(variable = fct_recode(variable, !!!new_names))
   print(dummy)
@@
     geom_blank(data=dummy) +
-    facet_wrap("type", scales="free_x") +
+    facet_wrap("variable", scales="free_x") +
     #theme(legend.position = "right", legend.direction = "vertical") 
     theme(legend.position = "bottom", legend.direction = "vertical")
-  #ggsave(filename="~/FRCBS/results-for-eba-paper/forest_plot.pdf", plot=g, dpi=600, units="mm", width=180)
   return(list(g=g, df=df))
@@
 if (save_figs)
-  ggsave(filename="performance_forest_plot.pdf", title="Performance forest plot", path=fig_path, plot=g, dpi=600, units="mm", width=180)
+  ggsave(filename="performance-forest-plot.pdf", #  title="Performance forest plot", 
+         path=fig_path, plot=g, dpi=600, units="mm", width=180, device=cairo_pdf)
 g
@@
 ```{r}
-g1 + theme_gray(base_size=20)
-g2 + theme_gray(base_size=20)
+base_size <- 20
+presentation1 <- g1 + coord_fixed() + theme_gray(base_size=base_size)
+presentation2 <- g2 + coord_fixed() + theme_gray(base_size=base_size)
+if (save_figs) {
+  ggsave(filename="presentation-classification-scatter-plot-hb-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
+         plot=presentation1, dpi=600, units="mm", width=180, height=120)
+  ggsave(filename="presentation-classification-scatter-plot-probability-finngen-male-dlmm.pdf", title="Classifiction scatter plot finngen-male-dlmm", path=fig_path, 
+         plot=presentation2, dpi=600, units="mm", width=180, height=120)
+}
+
+```
+
+Pretty print results as a table.
+
+```{r}
+df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
+df <- df %>% 
+  pivot_wider(names_from=c("variable", "type"), values_from=value, names_glue="{variable} {type}")
+names(df) <- str_remove(names(df), " value$")
+df %>%
+  select("Id", "F1", "F1 low", "F1 high", "threshold6", "threshold12", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high")
 ```
@@
 ```{r, eval=FALSE, include=FALSE}
-jarkko_test <- function(male_posterior, female_posterior, variables,
-                                           filename=NULL,
-                                           width = 180,  # width of the combined figure in millimetres
-                                          base_size = 11
-) {
-  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
-
-  combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
-    mutate(stripe=factor(row_number()%%2),    # this is used to make the gray/white background
-           Pretty=factor(Pretty, levels=Pretty))
-
-  result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
-  varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
-
-  result <- result %>%
-    mutate(Pretty=fct_rev(Pretty))
-
-  var_imp <- varimp %>%
-    mutate(Pretty=fct_rev(Pretty),
-           mean=Importance,
-           low=ifelse(is.na(Importance), NA, 0),
-           high=Importance)
-
-  final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
-    mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
-    select(-Importance, -Variable, stripe)
-
-#   final <- tibble(
-#             method = c("Regression coefficient",
-#                        "Regression coefficient","Importance"),
-#             Pretty = factor(c("Days to previous full blood donation","Days to previous full blood donation",
-#                        "Days to previous full blood donation")),
-#             #gender = c("male", "female", "both"),
-#               mean = c(0.0152091681221149,
-#                        0.0445741103338152,0.0255604042298423),
-#                low = c(0.00691102953212081, 0.0353192308283623, 0),
-#               high = c(0.0240855557569598,
-#                        0.0537929726515129,0.0255604042298423)
-# )
-
-  #combined_variables <- tibble(Pretty=factor("Days to previous full blood donation"), stripe=factor(1))
-
-  print(final)         
-  print(combined_variables)
-  vars <- c("Days to previous full blood donation", "Age", "Previous donation deferred", "Year")
-  plot <- final %>% #filter(Pretty %in% vars) %>%
-    ggplot(aes(color=gender)) +
-    geom_rect(data=combined_variables %>% filter(stripe==1),
-              mapping=aes(#ymax = as.numeric(Pretty) + 0.5,
-                          #ymin = as.numeric(Pretty) - 0.5
-                          ymax = as.numeric(Pretty) + 0.5,
-                          ymin = as.numeric(Pretty) - 0.5),
-              fill = "blue", 
-              #position="identity",
-              xmin=-Inf, xmax=Inf, alpha = 0.1, inherit.aes = FALSE) +
-    ggstance::geom_pointrangeh(#data=final, 
-                               mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high),
-                               #position=position_dodge2(0.8)
-                               position=position_dodge2(width=0.8, padding=0.1)
-                               ) +
-    scale_y_discrete() +
-    facet_wrap("method", scales = "free_x") +
-    #coord_flip() +
-    theme_classic(base_size = base_size)
-
-  return(list(plot=plot, final=final))
-}
-
-res <- jarkko_test(male_posterior, female_posterior, variables, filename=filename)
-final <- res$final
-g <- res$plot
-g
+# jarkko_test <- function(male_posterior, female_posterior, variables,
+#                                            filename=NULL,
+#                                            width = 180,  # width of the combined figure in millimetres
+#                                           base_size = 11
+# ) {
+#   rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
+# 
+#   combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
+#     mutate(stripe=factor(row_number()%%2),    # this is used to make the gray/white background
+#            Pretty=factor(Pretty, levels=Pretty))
+# 
+#   result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
+#   varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
+# 
+#   result <- result %>%
+#     mutate(Pretty=fct_rev(Pretty))
+# 
+#   var_imp <- varimp %>%
+#     mutate(Pretty=fct_rev(Pretty),
+#            mean=Importance,
+#            low=ifelse(is.na(Importance), NA, 0),
+#            high=Importance)
+# 
+#   final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
+#     mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
+#     select(-Importance, -Variable, stripe)
+# 
+# #   final <- tibble(
+# #             method = c("Regression coefficient",
+# #                        "Regression coefficient","Importance"),
+# #             Pretty = factor(c("Days to previous full blood donation","Days to previous full blood donation",
+# #                        "Days to previous full blood donation")),
+# #             #gender = c("male", "female", "both"),
+# #               mean = c(0.0152091681221149,
+# #                        0.0445741103338152,0.0255604042298423),
+# #                low = c(0.00691102953212081, 0.0353192308283623, 0),
+# #               high = c(0.0240855557569598,
+# #                        0.0537929726515129,0.0255604042298423)
+# # )
+#   
+#   #combined_variables <- tibble(Pretty=factor("Days to previous full blood donation"), stripe=factor(1))
+#   
+#   print(final)         
+#   print(combined_variables)
+#   vars <- c("Days to previous full blood donation", "Age", "Previous donation deferred", "Year")
+#   plot <- final %>% #filter(Pretty %in% vars) %>%
+#     ggplot(aes(color=gender)) +
+#     geom_rect(data=combined_variables %>% filter(stripe==1),
+#               mapping=aes(#ymax = as.numeric(Pretty) + 0.5,
+#                           #ymin = as.numeric(Pretty) - 0.5
+#                           ymax = as.numeric(Pretty) + 0.5,
+#                           ymin = as.numeric(Pretty) - 0.5),
+#               fill = "blue", 
+#               #position="identity",
+#               xmin=-Inf, xmax=Inf, alpha = 0.1, inherit.aes = FALSE) +
+#     ggstance::geom_pointrangeh(#data=final, 
+#                                mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high),
+#                                #position=position_dodge2(0.8)
+#                                position=position_dodge2(width=0.8, padding=0.1)
+#                                ) +
+#     scale_y_discrete() +
+#     facet_wrap("method", scales = "free_x") +
+#     #coord_flip() +
+#     theme_classic(base_size = base_size) 
+# 
+#   return(list(plot=plot, final=final))
+# }
+# 
+# res <- jarkko_test(male_posterior, female_posterior, variables, filename=filename)
+# final <- res$final
+# g <- res$plot
+# g
 ```

### 2020-12-02 src/dummy_classifier.Rmd
Added plot of logistic regression

@@
 date: "11/4/2020"
-output: md_document
+output: 
+  html_document: default
+  md_document: default
 ---
@@
 library(tidyverse)
+library(knitr)
 source("economic_effect_functions.R")
@@
 
-dummy_predictor2 <- function(train, validate) {  # real logistic regression
+logistic_regression <- function(train) {
   glm.fit <- glm(Hb_deferral ~ previous_Hb,
@@
                family = binomial)
+  return(glm.fit)  
+}
 
+dummy_predictor2 <- function(train, validate) {  # real logistic regression
+  glm.fit <- logistic_regression(train)
   print(glm.fit)
@@
 
-  return(result)  
+  return(list(result=result, fit=glm.fit))  
 }
-
-# get_metrics <- function(df, previous_Hb_threshold) {
-#   df <- df %>% mutate(predicted_Hb_deferral=factor(ifelse(previous_Hb < previous_Hb_threshold, "Deferred", "Accepted"), levels=c("Accepted", "Deferred")))
-#   
-#   cm <- confusionMatrix(reference = df$Hb_deferral, data = df$predicted_Hb_deferral, positive = "Deferred", mode = "everything")
-#   f1 <- cm$byClass["F1"]
-#          #Sensitivity  = TPR
-#   TPR <- cm$byClass['Sensitivity']
-#   FPR <- 1 - cm$byClass['Specificity']
-#   temp <- get_cost(TPR, FPR, TPR, FPR, p = parameters) # The parameters are from file economic_effect_functions.R
-#   E6  <- temp$E6
-#   E12 <- temp$E12
-#   return(tibble(previous_Hb_threshold=previous_Hb_threshold, f1=f1, E6=E6, E12=E12))
-# }
-# 
-# optimize_threshold <- function(df) {
-#   thresholds <- seq(120, 150, by=2)
-#   result <- map_dfr(thresholds, function(t) get_metrics(df, t))
-#   
-#   threshold <- result %>%
-#     slice_max(f1) %>%
-#     pull(previous_Hb_threshold)
-#   return(list(threshold=threshold, result=result))
-# }
 
@@
 ```{r}
-# res_female <- optimize_threshold(train_female)
-# res_male <- optimize_threshold(train_male)
-```
-
-Optimized thresholds on previous Hb for females and males
-```{r}
-# female_previous_Hb_threshold <- res_female$threshold
-# male_previous_Hb_threshold <- res_male$threshold
-# female_previous_Hb_threshold
-# male_previous_Hb_threshold
-```
-
-```{r}
-# res_female$result
-# res_male$result
-```
-
-Final result on validate data
-```{r}
-
-# f1_female <- get_f1(validate_female, female_previous_Hb_threshold)
-# f1_female
-```
-
-```{r}
-
-# f1_male <- get_f1(validate_male, male_previous_Hb_threshold)
-# f1_male
-```
-
-```{r}
 use_logistic_regression=TRUE
 if (use_logistic_regression) {
-  prediction_female <- dummy_predictor2(train_female, test_female)  # or would mean Hb be better centering point than deferral threshold
-  prediction_male <- dummy_predictor2(train_male, test_male)
+  res <- dummy_predictor2(train_female, test_female)
+  prediction_female <- res$result
+  fit_female <- res$fit
+
+  res <- dummy_predictor2(train_male, test_male)
+  prediction_male <- res$result
+  fit_male <- res$fit
 } else {
@@
 head(prediction_male)
+fit_male
+fit_female
 ```
@@
 ```{r}
+if (FALSE) {
   save(prediction_female, file="~/FRCBS/interval-sims/progesa-validate-female-dummy2.rdata")
   save(prediction_male, file="~/FRCBS/interval-sims/progesa-validate-male-dummy2.rdata")
+}
 ```
@@
 ```{r}
-dummy_result <- process_all_data(dummy2_ids, replicates=100)
+dummy_result <- suppressMessages(process_all_data(dummy2_ids, replicates=100))
 ```
@@
 
-df
+kable(df)
+```
+
+```{r}
+data <- tibble(previous_Hb = seq(100, 200))
+data[["male"]] <- predict(fit_male, newdata=data, type="response")
+data[["female"]] <- predict(fit_female, newdata=data, type="response")
+data2 <- data %>% pivot_longer(cols=c(male, female), names_to="Sex", values_to="Probability")
+#data2
+
+mt <- df %>% filter(Id=="progesa-male-dummy") %>% select(threshold6, threshold12) %>% as.numeric()   # male thresholds
+ft <- df %>% filter(Id=="progesa-female-dummy") %>% select(threshold6, threshold12) %>% as.numeric() # female thresholds
+mhb <- approx(x = data$male,   y = data$previous_Hb, xout=mt)$y                                          # corresponding male hemoglobins
+fhb <- approx(x = data$female, y = data$previous_Hb, xout=ft)$y                                          # corresponding female hemoglobins
+threshold_data <- tibble(previous_Hb = c(mhb, fhb), Probability=c(mt, ft), text=c("t6", "t12", "t6", "t12"), Sex=c("male", "male", "female", "female"))
+kable(threshold_data)
+
+g <- data2 %>% ggplot(aes(x=previous_Hb, y=Probability, color=Sex)) + 
+  geom_line() +
+  geom_vline(data=threshold_data, mapping=aes(xintercept=previous_Hb, colour=Sex)) +
+  geom_text(data=threshold_data, mapping=aes(x=previous_Hb, label=text), y=0.1, show.legend=FALSE, nudge_x=2.5)
+g
 ```
@@
 roc_female
-pr_female
+pr_female$pr_plot
 roc_male
-pr_male
+pr_male$pr_plot
 ```
-
-```{r}
-# v <- c(roc_female$roc$ci[c(2,1,3)], pr_female$ci[c(1,2,3)], 
-#        roc_male$roc$ci[c(2,1,3)], pr_male$ci[c(1,2,3)])
-# aucs <- as_tibble(matrix(v, ncol=6, byrow=TRUE))
-# colnames(aucs) <- c("AUROC", "AUROC low", "AUROC high", "AUPR value", "AUPR low", "AUPR high")
-# aucs <- aucs %>% 
-#   mutate(Id=dummy2_ids) %>%
-#   select(Id, everything())
-# 
-# aucs
-```
-
-```{r}
-# dummy_result <- inner_join(aucs, dummy_result, by="Id")
-# dummy_result
-```

### 2020-12-02 src/economic_effect_functions.R
Cleaned up code and parameterized the bootstrap method

@@ -420,16 +420,16 @@ get_f1 <- function(df, threshold = 0.5) {
 # Computes the area under the precision-recall curve
 # See get_f1 for parameter description.
 get_aupr <- function(df) {
-  aupr <- PRROC::pr.curve(scores.class0=df$Deferred, 
-                          weights.class0=df$obs=="Deferred")$auc.davis.goadrich
+  aupr <- suppressMessages(PRROC::pr.curve(scores.class0=df$Deferred, 
+                          weights.class0=df$obs=="Deferred")$auc.davis.goadrich)
   return(aupr)   # Returns a single value
 }
 
 # Computest the area under the receiver operating characteristic curve
 # See get_f1 for parameter description.
 get_auroc <- function(df) {
-  auroc <- pROC::auc(response = df$obs=="Deferred",
-                     predictor = df$Deferred)
+  auroc <- suppressMessages(pROC::auc(response = df$obs=="Deferred",
+                     predictor = df$Deferred))
   return(auroc)   # Returns a single value
 }
           
@@ -463,7 +463,8 @@ boot_cost <- function(boot_data, boot_ind, f1_threshold, threshold6 = 0.6, thres
   return(costs)   # returns a vector of statistics with element names E6, E12, F1, AUPR, and AUROC
 }
 
-compute_cis <-function(fit_boot, conf = 0.95){
+# method is either "norm", "basic", "perc", or "bca"
+compute_cis <-function(fit_boot, conf = 0.95, method="norm"){
   statistics <- fit_boot$t0  # for example E6, E12, F1, AUPR, and AUROC
   n_statistics <- length(statistics)
   CI_low = rep(0, n_statistics)  # Create arrays for end points of confidence intervals
@@ -482,9 +483,15 @@ compute_cis <-function(fit_boot, conf = 0.95){
         #Bca_low[i_regressor] <- CI$bca[4] # N.B type = 'normal' has 3 columns, while the other types 5 columns
         #Bca_high[i_regressor] <- CI$bca[5] #
         #there is just too much data for bca to be used
-        CI <- boot.ci(fit_boot, conf = conf, type = "norm", index=i_regressor)  # using normal approximation
-        CI_low[i_regressor] <- CI$normal[2]
-        CI_high[i_regressor] <- CI$normal[3]
[…]

### 2020-12-03 src/article_figures.Rmd
Updated the table in md

@@
 source("economic_effect_functions.R")
+library(knitr)
+library(tidyverse)
 save_figs = TRUE
@@
 
-  print(combined_variables)
+  #print(combined_variables)
 
@@
 
-get_threshold_df <- function() {
-
-  t6 <- tibble(
-      Id = factor(all_ids),
-      type = factor("Threshold"),
-      value = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1, 0.4, 0.6,   0.2, 0.3))
-  t12 <- tibble(
-      Id = factor(all_ids),
-      type = factor("Threshold"),
-      value = c(0.2, 0.2, 0.1, 0.1, 0.1, 0.2, 0.3, 0.3, 0.8, 0.8,   0.4, 0.5)
-  )
-  t <- bind_rows(t6, t12)
-  t <- t %>%
-    mutate(low=value, 
-           high=value,
-           sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
-           #model=factor(str_split_fixed(Id, "-", 3)[,3])
-           model=rep(c("t6", "t12"), each=length(all_ids))
-           )
-  return(t)
-}
+# get_threshold_df <- function() {
+#   
+#   t6 <- tibble(
+#       Id = factor(all_ids),
+#       type = factor("Threshold"),
+#       value = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1, 0.4, 0.6,   0.2, 0.3))
+#   t12 <- tibble(
+#       Id = factor(all_ids),
+#       type = factor("Threshold"),
+#       value = c(0.2, 0.2, 0.1, 0.1, 0.1, 0.2, 0.3, 0.3, 0.8, 0.8,   0.4, 0.5)
+#   )
+#   t <- bind_rows(t6, t12)
+#   t <- t %>%
+#     mutate(low=value, 
+#            high=value,
+#            sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
+#            #model=factor(str_split_fixed(Id, "-", 3)[,3])
+#            model=rep(c("t6", "t12"), each=length(all_ids))
+#            )
+#   return(t)
+# }
 
@@
     mutate(variable = fct_recode(variable, !!!new_names))
-  print(dummy)
-  print(df)
+  #print(dummy)
+  #print(df)
 
@@
 ```{r}
-df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
+df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
 df <- df %>% 
@@
 names(df) <- str_remove(names(df), " value$")
-df %>%
+df2 <- df %>%
   select("Id", "F1", "F1 low", "F1 high", "threshold6", "threshold12", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high")
+kable(df2, digits=2)
 ```
-
-```{r, eval=FALSE, include=FALSE}
-# jarkko_test <- function(male_posterior, female_posterior, variables,
-#                                            filename=NULL,
-#                                            width = 180,  # width of the combined figure in millimetres
-#                                           base_size = 11
-# ) {
-#   rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
-# 
-#   combined_variables <- tibble(Pretty=c(variables, "Life time donations", "Sex", "First Hb")) %>%
-#     mutate(stripe=factor(row_number()%%2),    # this is used to make the gray/white background
-#            Pretty=factor(Pretty, levels=Pretty))
-# 
-#   result <- get_dlmm_df(male_posterior, female_posterior, variables, combined_variables)
-#   varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
-# 
-#   result <- result %>%
-#     mutate(Pretty=fct_rev(Pretty))
-# 
-#   var_imp <- varimp %>%
-#     mutate(Pretty=fct_rev(Pretty),
-#            mean=Importance,
-#            low=ifelse(is.na(Importance), NA, 0),
-#            high=Importance)
-# 
-#   final <- bind_rows(`Regression coefficient`=result, Importance=var_imp, .id="method") %>%
-#     mutate(method=factor(method, levels=c("Regression coefficient", "Importance"))) %>%
-#     select(-Importance, -Variable, stripe)
-# 
-# #   final <- tibble(
-# #             method = c("Regression coefficient",
-# #                        "Regression coefficient","Importance"),
-# #             Pretty = factor(c("Days to previous full blood donation","Days to previous full blood donation",
-# #                        "Days to previous full blood donation")),
-# #             #gender = c("male", "female", "both"),
-# #               mean = c(0.0152091681221149,
-# #                        0.0445741103338152,0.0255604042298423),
-# #                low = c(0.00691102953212081, 0.0353192308283623, 0),
-# #               high = c(0.0240855557569598,
-# #                        0.0537929726515129,0.0255604042298423)
-# # )
-#   
-#   #combined_variables <- tibble(Pretty=factor("Days to previous full blood donation"), stripe=factor(1))
-#   
-#   print(final)         
-#   print(combined_variables)
-#   vars <- c("Days to previous full blood donation", "Age", "Previous donation deferred", "Year")
-#   plot <- final %>% #filter(Pretty %in% vars) %>%
-#     ggplot(aes(color=gender)) +
-#     geom_rect(data=combined_variables %>% filter(stripe==1),
-#               mapping=aes(#ymax = as.numeric(Pretty) + 0.5,
-#                           #ymin = as.numeric(Pretty) - 0.5
-#                           ymax = as.numeric(Pretty) + 0.5,
-#                           ymin = as.numeric(Pretty) - 0.5),
-#               fill = "blue", 
-#               #position="identity",
-#               xmin=-Inf, xmax=Inf, alpha = 0.1, inherit.aes = FALSE) +
-#     ggstance::geom_pointrangeh(#data=final, 
-#                                mapping=aes(y=Pretty, x=mean, xmin=low, xmax=high),
-#                                #position=position_dodge2(0.8)
-#                                position=position_dodge2(width=0.8, padding=0.1)
-#                                ) +
-#     scale_y_discrete() +
-#     facet_wrap("method", scales = "free_x") +
-#     #coord_flip() +
-#     theme_classic(base_size = base_size) 
-# 
-#   return(list(plot=plot, final=final))
-# }
-# 
-# res <- jarkko_test(male_posterior, female_posterior, variables, filename=filename)
-# final <- res$final
-# g <- res$plot
-# g
-```

### 2020-12-18 src/article_figures.Rmd
Small change

@@
 ) {
-  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
 
@@
 
-  # Get variable importance from the RF object
+  # Get variable importances from the RF object
+  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
   varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)
@@
 df2 <- df %>%
-  select("Id", "F1", "F1 low", "F1 high", "threshold6", "threshold12", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high")
+  select("Id", "AUROC", "AUROC low", "AUROC high", "AUPR", "AUPR low", "AUPR high", "F1", "F1 low", "F1 high", "threshold6", "threshold12", "E6", "E6 low", "E6 high", "E12", "E12 low", "E12 high")
 kable(df2, digits=2)

### 2021-06-07 src/dummy_classifier.Rmd
Small change
Cleaned up code.

@@
 
-```{r}
+```{r, eval=FALSE}
 df1 <- make_threshold_conversion_data_frame(125, -1) 
@@
 
-```{r}
+```{r, eval=FALSE}
 df1 <- make_threshold_conversion_data_frame(135, -1) 
@@
 if (use_logistic_regression) {
-  res <- dummy_predictor2(train_female, test_female)
-  prediction_female <- res$result
-  fit_female <- res$fit
+  res_female <- dummy_predictor2(train_female, test_female)
+  prediction_female <- res_female$result
+  fit_female <- res_female$fit
 
-  res <- dummy_predictor2(train_male, test_male)
-  prediction_male <- res$result
-  fit_male <- res$fit
-} else {
+  res_male <- dummy_predictor2(train_male, test_male)
+  prediction_male <- res_male$result
+  fit_male <- res_male$fit
+} else {  # Old way was to do it by hand using logistic function
   prediction_female <- dummy_predictor(validate_female, 125)  # or would mean Hb be better centering point than deferral threshold
@@
 ```{r}
-dummy_result <- suppressMessages(process_all_data(dummy2_ids, replicates=100))
+#dummy_result <- suppressMessages(process_all_data(dummy2_ids, replicates=100))
+dummy_result <- suppressMessages(process_all_data(dummy2_ids))
 ```
@@
 ```
+
+Compute the effects of applying baseline model. This needs still some cleanup.
+
+```{r}
+result_male <- res_male$result %>% mutate(prediction=as.factor(ifelse(Deferred>0.36, "Deferred", "Accepted")))
+result_female <- res_female$result %>% mutate(prediction=as.factor(ifelse(Deferred>0.64, "Deferred", "Accepted")))
+```
+
+```{r}
+summary(result_male)
+```
+
+```{r}
+summary(result_female)
+```
+
+```{r}
+nm <- nrow(result_male)
+nf <- nrow(result_female)
+md <- sum(result_male$obs=="Deferred")
+fd <- sum(result_female$obs=="Deferred")
+cat(sprintf("Men deferred %i (%.2f%%)\n", md, md/nm*100))
+cat(sprintf("Women deferred %i (%.2f%%)\n", fd, fd/nf*100))
+
+md2 <- sum(result_male$obs=="Deferred" & result_male$prediction=="Accepted")
+fd2 <- sum(result_female$obs=="Deferred" & result_female$prediction=="Accepted")
+cat(sprintf("Men deferred %i (%.2f%%)\n", md2, md2/nm*100))
+cat(sprintf("Women deferred %i (%.2f%%)\n", fd2, fd2/nf*100))
+
+```
+```{r}
+mextension <- sum(result_male$prediction=="Deferred")
+fextension <- sum(result_female$prediction=="Deferred")
+cat(sprintf("Men extended for 6 months %i (%.2f%%)\n", mextension, mextension/nm*100))
+cat(sprintf("Women extended for 6 months %i (%.2f%%)\n", fextension, fextension/nf*100))
+```

### 2021-06-07 src/economic_effect_functions.R

@@ -152,72 +152,89 @@ get_cost <- function(TPR6, FPR6, TPR12, FPR12, p = parameters) {
   Fn.=p$Fn
   rl.=p$rl
   
-  if (FALSE) {
-    get_cost2 <- function(TPR, FPR) {
+#  if (TRUE) {  # This version has less repetition
+    get_cost2 <- function(TPR, FPR, new_interval_length) {
+      male_interval_factor   <- new_interval_length / 2
+      female_interval_factor <- new_interval_length / 3
       q <- d. * TPR
-      a <- 1 * ((1 - FPR) * (1 - d.)) + 
-        3 * mr. * (FPR * (1 - d.)) + 2 * fr. * (FPR * (1 - d.)) + 
-        3 * mdr. * d. + 2 * fdr. * d.
+      # Average donation interval extension over all donors.
+      # The right parts of the terms sum to one.
+      a <- 1 * ((1 - FPR) * (1 - d.)) + # no extension, donors correctly predicted as accepted
+        male_interval_factor * mr. * (FPR * (1 - d.)) + female_interval_factor * fr. * (FPR * (1 - d.)) + # donors falsely predicted as deferred 
+        male_interval_factor * mdr. * d.              + female_interval_factor * fdr. * d.                # donors that are deferred or predicted as such 
       Fratio <- a / (1 + Fn. * (a - 1))
       Em <- Pm. * (Fratio - 1 - q * rl.)
       Ed <- Pd. * q
       E <- Em - Ed
       tibble(q=q, a=a, E=E, TPR=TPR, FPR=FPR)
     }
-    df6 <- get_cost2(TPR6, FPR6)
-    df12 <- get_cost2(TPR12, FPR12)
+    df6 <- get_cost2(TPR6, FPR6, 6)
+    df12 <- get_cost2(TPR12, FPR12, 12)
     names(df6) <- paste0(names(df6), "6")
     names(df12) <- paste0(names(df12), "12")
-    return(bind_cols(df6, df12))
-  } else {
-    q6 <- d. * TPR6
-    q12 <- d. * TPR12
-    a6 <- 1 * ((1 - FPR6) * (1 - d.)) +
-      3 * mr. * (FPR6 * (1 - d.)) + 2 * fr. * (FPR6 * (1 - d.)) +
-      3 * mdr. * d. + 2 * fdr. * d.
+    df <- bind_cols(df6, df12) %>% 
[…]

### 2021-06-07 src/article_figures.Rmd
Show also atot, q, TPR, and FPR in a table.

@@
 library(tidyverse)
-save_figs = TRUE
+save_figs = FALSE
 fig_path="~/FRCBS/results-for-eba-paper/pdf"
@@
 }
+```
 
+```{r, eval=FALSE}
 # Creates a forest plot of means and confidence intervals from posterior distributions.
@@
   s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
+  # Note that recode stupidly uses order "old"="new"
   s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
@@
     #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
-    labs(x="Value", y="Data and method", colour="Sex", shape="Model", fill="Deferral length") + 
+    labs(x="Value", y="Data and model", colour="Sex", shape="Model", fill="Deferral length") + 
     scale_y_discrete(labels=get_name_of_data) +
@@
     scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "dummy"),
-                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Dummy"),
+                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Baseline"),
                        values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
@@
       mutate(new_predicted_label= ifelse(predicted < hb_threshold, 1, 0))
+    prediction_threshold <- hb_threshold
   } else { # predict deferral based on probability of deferral
@@
       mutate(new_predicted_label= ifelse(scores >= probability_of_deferral_threshold, 1, 0))
+    prediction_threshold <- min(df %>% filter(scores <= probability_of_deferral_threshold) %>% pull(predicted))
   }
+
+  cat(sprintf("Prediction threshold is %.1f\n", prediction_threshold))
+
+  # Get the R2 correlation
+  r2 <- summary(lm(predicted ~ observed, data=df))$r.squared
+  r2_text <- sprintf("%.2f", r2)
+
   df <- df %>%
@@
            levels=c("True positive", "False negative", "False positive", "True negative")))
+
+  n <- nrow(df)
+  fp <- sum(df$confusion_class=="False positive")
+  tp <- sum(df$confusion_class=="True positive")
+  deferrals <- nrow(df %>% filter(confusion_class %in% c("True positive", "False negative")))
+  accepted <- n - deferrals
+
+  cat(sprintf("Incorrectly predicting %.1f%% of accepted donations as deferrals\n", 100*fp/accepted))
+  cat(sprintf("Correctly predicting %.1f%% of deferred donations as deferrals\n", 100*tp/deferrals))
+
   scatter_plot <- ggplot(df, aes(x = observed, y=predicted, color = confusion_class)) +
@@
     geom_vline(xintercept = hb_threshold, linetype = "dashed") +
-    geom_hline(yintercept = hb_threshold, linetype = "dashed") +
+    geom_hline(yintercept = prediction_threshold, linetype = "dashed") +
+    annotate("text", x=120, y=190, color="black", label=bquote(R^2==.(r2_text)), fontface="bold") +
     theme(legend.position = "bottom")
@@
 g1 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, NULL)
-g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.1)
+g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.04)
 if (save_figs) {
@@
 ```
+
+Create a table of atot and TPR and FPR values for each method/data. Q is the fraction of avoided deferrals.
+
+```{r}
+#data_frames <- map(dummy2_ids, get_data_frame)
+df2 <- df %>% select(Id, threshold6, threshold12)  # Get the optimal thresholds
+```
+
+```{r}
+helper <- function(Id, threshold6, threshold12) {
+  df2 <- get_data_frame(Id)
+  r6 <- get_rates(df2, threshold6)
+  r12 <- get_rates(df2, threshold12)
+  df <- get_cost(r6$TPR, r6$FPR, r12$TPR, r12$FPR, p = parameters)
+  #df <- tibble(a=1, b=2)
+  df$Id <- Id
+  return(df %>% select(Id, everything()))  
+}
+
+d <- get_mikkos_cost_constants()$d  # Fraction of deferrals of all donors
+result <- pmap_dfr(df2, helper) 
+result <- result %>%
+  mutate(across(c(q6, q12), function(x) x / d)) %>%
+  rename(Q6=q6, Q12=q12)
+kable(result, digits=3)
+```

### 2021-06-10 src/article_figures.Rmd
Fixed cost computation.
There was a problem with separation into males and females.

@@
 fig_path="~/FRCBS/results-for-eba-paper/pdf"
+table_path="~/FRCBS/results-for-eba-paper/table"
 ```
@@
 
-```{r}
+```{r, Get optimal thresholds}
 #data_frames <- map(dummy2_ids, get_data_frame)
-df2 <- df %>% select(Id, threshold6, threshold12)  # Get the optimal thresholds
+compute=TRUE
+if (compute) {
+  threshold_range <- seq(0.02, 0.98, 0.02)
+  data_frames <- map(all_ids, get_data_frame)
+  thresholds <- map2_dfr(data_frames, all_ids, function(df, id) get_optimal_thresholds(df, parameters, threshold_range, id)) %>%
+    mutate(Id=all_ids) %>%
+    select(Id, threshold6, threshold12)
+} else {
+  thresholds <- df %>% select(Id, threshold6, threshold12)  # Get the optimal thresholds
+}
 ```
@@
 helper <- function(Id, threshold6, threshold12) {
-  df2 <- get_data_frame(Id)
-  r6 <- get_rates(df2, threshold6)
-  r12 <- get_rates(df2, threshold12)
-  df <- get_cost(r6$TPR, r6$FPR, r12$TPR, r12$FPR, p = parameters)
+  df3 <- get_data_frame(Id)
+  r6 <- get_rates(df3, threshold6)
+  r12 <- get_rates(df3, threshold12)
+  sex <- get_sex(Id)
+  df <- get_cost(r6$TPR, r6$FPR, r12$TPR, r12$FPR, sex=sex, p = parameters)
   #df <- tibble(a=1, b=2)
@@
 d <- get_mikkos_cost_constants()$d  # Fraction of deferrals of all donors
-result <- pmap_dfr(df2, helper) 
+result <- pmap_dfr(thresholds, helper) 
 result <- result %>%
   mutate(across(c(q6, q12), function(x) x / d)) %>%
-  rename(Q6=q6, Q12=q12)
+  inner_join(thresholds, by="Id")
+  #rename(Q6=q6, Q12=q12)
 kable(result, digits=3)
 ```
+
+Get all variables and their confidence intervals
+
+```{r}
+  all_results <- process_all_data(all_ids)
+  filename <- sprintf("%s/raw-data-2021-06-10.tsv", table_path)
+  write_tsv(all_results, file=filename)
+```
+Same results in wide format xlsx
+```{r}
+library(openxlsx)
+temp <- all_results %>% pivot_wider(names_from=c(variable, type))
+filename <- sprintf("%s/pretty-data-2021-06-10.xlxs", table_path)
+openxlsx::write.xlsx(temp, file = filename, overwrite=TRUE)
+```
+
+```{r}
+g <- draw_surface(results=all_results %>% filter(Id == "progesa-both-rf"))
+if (save_figs) {
+  filename <- sprintf("%s/cme_cost_surface_werr.pdf", fig_path)
+  ggsave(filename=filename,  width = 180,  height = 120, units="mm", dpi=600, scale=1.0, device=cairo_pdf)
+}
+g
+```

### 2021-06-10 src/economic_effect_functions.R

@@ -9,10 +9,10 @@ library(boot)
 ids <- c("progesa-male-lmm", "progesa-female-lmm", "progesa-male-dlmm", "progesa-female-dlmm", "finngen-male-dlmm", "finngen-female-dlmm", 
          "findonor-male-dlmm", "findonor-female-dlmm", "progesa-both-dt", "progesa-both-rf")
 
-dummy2_ids <- c("progesa-male-dummy", "progesa-female-dummy")
-dummy_ids <- c("finngen-male-stratified", "finngen-male-most-frequent", "finngen-male-prior", "finngen-male-uniform", "finngen-male-deferred")
+dummy_ids <- c("progesa-male-dummy", "progesa-female-dummy")
+dummy2_ids <- c("finngen-male-stratified", "finngen-male-most-frequent", "finngen-male-prior", "finngen-male-uniform", "finngen-male-deferred")
 
-all_ids <- c(ids, dummy2_ids)
+all_ids <- c(ids, dummy_ids)
 
 # These are from Progesa data
 get_mikkos_cost_constants <- function() {
@@ -141,7 +141,7 @@ get_cost_constants <- function(data) {
 
 
 #get_cost <- function(TPR6, FPR6, TPR12, FPR12, d.=d, mr.=mr, fr.=fr, mdr.=mdr, fdr.=fdr, Pm.=Pm, Pd.=Pd, Fn.=Fn, rl.=rl) {
-get_cost <- function(TPR6, FPR6, TPR12, FPR12, p = parameters) {
+get_cost <- function(TPR6, FPR6, TPR12, FPR12, sex, p = parameters) {
   d.=p$d
   mr.=p$mr
   fr.=p$fr
@@ -152,53 +152,84 @@ get_cost <- function(TPR6, FPR6, TPR12, FPR12, p = parameters) {
   Fn.=p$Fn
   rl.=p$rl
   
-#  if (TRUE) {  # This version has less repetition
-    get_cost2 <- function(TPR, FPR, new_interval_length) {
-      male_interval_factor   <- new_interval_length / 2
-      female_interval_factor <- new_interval_length / 3
-      q <- d. * TPR
-      # Average donation interval extension over all donors.
-      # The right parts of the terms sum to one.
-      a <- 1 * ((1 - FPR) * (1 - d.)) + # no extension, donors correctly predicted as accepted
[…]

### 2021-06-11 src/article_figures.Rmd
all_results now includes TPR and FPR as well

--- 2021-06-07 src/article_figures.Rmd
@@
 save_figs = FALSE
+save_tables = FALSE
 fig_path="~/FRCBS/results-for-eba-paper/pdf"
+table_path="~/FRCBS/results-for-eba-paper/table"
+all_results_filename <- sprintf("%s/raw-data-2021-06-11.tsv", table_path)
+compute=FALSE
+```
+
+Get all variables and their confidence intervals
+
+```{r}
+if (compute) {
+  all_results <- process_all_data(all_ids)
+  write_tsv(all_results, file=all_results_filename)
+} else {
+  all_results <- read_tsv(file=all_results_filename)  
+}
+```
+
+Same results in wide format xlsx
+
+```{r}
+if (save_tables) {
+  library(openxlsx)
+  temp <- all_results %>% pivot_wider(names_from=c(variable, type))
+  filename <- sprintf("%s/pretty-data-2021-06-11.xlsx", table_path)
+  openxlsx::write.xlsx(temp, file = filename, overwrite=TRUE)
+}
+```
+
+```{r Cost surface}
+g <- draw_surface(results=all_results %>% filter(Id == "progesa-both-rf"))
+if (save_figs) {
+  filename <- sprintf("%s/cme_cost_surface_werr.pdf", fig_path)
+  ggsave(filename=filename,  width = 180,  height = 120, units="mm", dpi=600, scale=1.0, device=cairo_pdf)
+}
+g
+```
+
+```{r Probability thresholds}
+threshold_range <- seq(0.02, 0.98, 0.02)
+data_frames <- map(all_ids, get_data_frame)
+#get_thresholds <- function(df,  p=parameters, thresholds = seq(0.1, .9, .1), id=id)
+thresholds <- map2_dfr(data_frames, all_ids, function(df, id) get_thresholds(df, parameters, threshold_range, id))
+
+thresholds2 <- thresholds %>% 
+  filter(Id=="progesa-both-rf") %>%
+  select(probability, E6, E12) %>%
+  pivot_longer(cols = c(E6, E12), names_to = "variable", values_to = "E") %>%
+  extract(col="variable", regex = "([a-zA-Z]+)([0-9]+)", into=c("variable", "month")) %>%
+  select(-variable)
+
+min_thresholds <- thresholds2 %>%
+  group_by(month) %>%
+  slice_min(order_by=E, n=1, with_ties = FALSE) %>%
+  ungroup()
+```
+
+```{r Probability cutoff curve}
+g <- thresholds2 %>%
+  mutate(month=factor(month, levels=c(6, 12))) %>%
+  ggplot(aes(x=probability, y=E, color=month)) +
+  geom_line() +
+  geom_point(data=min_thresholds, mapping=aes(x=probability, y=E), 
+             color="black", size=2, inherit.aes = FALSE) +
+  geom_label(data=min_thresholds, mapping=aes(x=probability, y=E, label=sprintf("p >= %.2f : %.2f €", probability, E)), 
+             inherit.aes = FALSE, nudge_y=0.3, size=2) +
+  ylab("Cost effect (€ / donation)") + xlab("Probability of deferral cut-off") + labs(color="Deferral length (mo)") +
+  theme(legend.position="bottom")
+g
+if (save_figs) {
+  filename <- sprintf("%s/cme_prob_cut_curve.pdf", fig_path)
+  ggsave(filename=filename,  width = 90,  height = 80, units="mm", dpi=600, scale=1.0, device=cairo_pdf)
+}
 ```
@@
 
+This is the OLD visualisation.
+
 ```{r, eval=FALSE}
@@
 
-```{r, include=FALSE}
+```{r Read FinnGen DLMM posteriors, include=FALSE}
 finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
@@
 
-## Alternative effect sizes and importances of variables
+## New effect sizes and importances of variables
 
 ```{r, include=FALSE}
-alternative_create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
+new_create_forest_importance_plot <- function(male_posterior, female_posterior, variables,
                                            filename=NULL,
                                            width = 180,  # width of the combined figure in millimetres
-                                          base_size = 11
-) {
+                                          base_size = 11) 
+{
 
@@
 ```{r}
-res <- alternative_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
+res <- new_create_forest_importance_plot(male_posterior, female_posterior, variables, filename=filename)
 g <- res$plot
@@
 
-all_ids <- c(ids, dummy2_ids)
-
-# get_threshold_df <- function() {
-#   
-#   t6 <- tibble(
-#       Id = factor(all_ids),
-#       type = factor("Threshold"),
-#       value = c(0.1, 0.1, 0.1, 0.1, 0.1, 0.1, 0.2, 0.1, 0.4, 0.6,   0.2, 0.3))
-#   t12 <- tibble(
-#       Id = factor(all_ids),
-#       type = factor("Threshold"),
-#       value = c(0.2, 0.2, 0.1, 0.1, 0.1, 0.2, 0.3, 0.3, 0.8, 0.8,   0.4, 0.5)
-#   )
-#   t <- bind_rows(t6, t12)
-#   t <- t %>%
-#     mutate(low=value, 
-#            high=value,
-#            sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
-#            #model=factor(str_split_fixed(Id, "-", 3)[,3])
-#            model=rep(c("t6", "t12"), each=length(all_ids))
-#            )
-#   return(t)
-# }
+#all_ids <- c(ids, dummy_ids)
 
@@
   s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
-  # s <- map(s, function(x) when(x, 
-  #                              x == "progesa" ~ "eProgesa",
-  #                              x == "finngen" ~ "Biobank",
-  #                              x == "findonor" ~ "FinDonor",
-  #                              ~ "koe"))
   return(s)
@@
   if (FALSE) {
-    df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
-    df2 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data_dummy.csv")
-    df <- bind_rows(df1, df2)
-    #  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")
-
-    df <- df %>%
-      mutate(Id=factor(Id, levels=rev(all_ids)),
-             type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
-             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
-             model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
-      mutate(type = fct_recode(type, !!!new_names))
-
-    thresholds <- get_threshold_df()
-    df <- bind_rows(df, thresholds)
-    df <- df %>% 
-      mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
+    # df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
+    # df2 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data_dummy.csv")
+    # df <- bind_rows(df1, df2)
+    # #  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")
+    # 
+    # df <- df %>%
+    #   mutate(Id=factor(Id, levels=rev(all_ids)),
+    #          type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
+    #          sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
+    #          model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
+    #   mutate(type = fct_recode(type, !!!new_names))
+    # 
+    # thresholds <- get_threshold_df()
+    # df <- bind_rows(df, thresholds)
+    # df <- df %>% 
+    #   mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
     #write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")
@@
 #    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
-    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
+#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
+    df <- read_tsv(all_results_filename)
 
     df <- df %>% 
+      filter(variable %in% c("AUROC", "AUPR", "F1", "E6", "E12", "threshold6", "threshold12")) %>%
       pivot_wider(names_from=c("type"), values_from=value) %>%
@@
                              TRUE ~ str_split_fixed(Id, "-", 3)[,3]),
-             model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")),
+             model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "baseline", "t6", "t12")),
              variable = recode(variable, "threshold6" = "Threshold", "threshold12" = "Threshold"),
@@
     scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
-    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "dummy"),
+    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "baseline"),
                        labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Baseline"),
@@
 df <- res$df
-if (save_figs)
-  ggsave(filename="performance-forest-plot.pdf", #  title="Performance forest plot", 
-         path=fig_path, plot=g, dpi=600, units="mm", width=180, device=cairo_pdf)
+if (save_figs) {
+  filename <- sprintf("%s/performance-forest-plot.pdf", fig_path)
+  ggsave(filename=filename, #  title="Performance forest plot", 
+         plot=g, dpi=600, units="mm", width=180, device=cairo_pdf)
+}
 g
@@
 
-```{r}
-df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
+```{r Performances as LaTeX table}
+#df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
+df <- read_tsv(all_results_filename)
 df <- df %>% 
@@
 kable(df2, digits=2)
+if (save_tables) {
+  filename <- sprintf("%s/performance.tex", table_path)
+  latex_code <- kable(df2, format="latex", digits=2, 
+                      caption="The numeric values of the classification performance metrics and their 95\\% confidence intervals in Fig. 2.", 
+                      label="tab:performance",
+                      linesep="")
+  cat(latex_code, file=filename, sep="\n")
+}
 ```
 
-Create a table of atot and TPR and FPR values for each method/data. Q is the fraction of avoided deferrals.
+## Everything below this is unnecessary
 
-```{r}
+Create a table of atot and TPR and FPR values for each method/data. q is the fraction of avoided deferrals.
+
+```{r Get optimal thresholds, eval=FALSE}
 #data_frames <- map(dummy2_ids, get_data_frame)
-df2 <- df %>% select(Id, threshold6, threshold12)  # Get the optimal thresholds
+#compute=TRUE
+if (FALSE) {
+  threshold_range <- seq(0.02, 0.98, 0.02)
+  data_frames <- map(all_ids, get_data_frame)
+  thresholds <- map2_dfr(data_frames, all_ids, function(df, id) get_optimal_thresholds(df, parameters, threshold_range, id)) %>%
+    mutate(Id=all_ids) %>%
+    select(Id, threshold6, threshold12)
+} else {
+  optimal_thresholds <- all_results %>% 
+    filter(variable %in% c("threshold6", "threshold12")) %>%
+    pivot_wider(names_from=variable, values_from = value) %>%
+    select(-type)
+}
 ```
 
-```{r}
+```{r, eval=FALSE}
 helper <- function(Id, threshold6, threshold12) {
-  df2 <- get_data_frame(Id)
-  r6 <- get_rates(df2, threshold6)
-  r12 <- get_rates(df2, threshold12)
-  df <- get_cost(r6$TPR, r6$FPR, r12$TPR, r12$FPR, p = parameters)
+  df3 <- get_data_frame(Id)
+  r6 <- get_rates(df3, threshold6)
+  r12 <- get_rates(df3, threshold12)
+  sex <- get_sex(Id)
+  df <- get_cost(r6$TPR, r6$FPR, r12$TPR, r12$FPR, sex=sex, p = parameters)
   #df <- tibble(a=1, b=2)
@@
 
-d <- get_mikkos_cost_constants()$d  # Fraction of deferrals of all donors
-result <- pmap_dfr(df2, helper) 
-result <- result %>%
-  mutate(across(c(q6, q12), function(x) x / d)) %>%
-  rename(Q6=q6, Q12=q12)
+#d <- get_mikkos_cost_constants()$d  # Fraction of deferrals of all donors
+result <- pmap_dfr(optimal_thresholds, helper) 
+# result <- result %>%
+#   mutate(across(c(q6, q12), function(x) x ) %>%
+#   inner_join(thresholds, by="Id")
+#   #rename(Q6=q6, Q12=q12)
 kable(result, digits=3)

### 2021-06-11 src/economic_effect_functions.R

@@ -9,7 +9,7 @@ library(boot)
 ids <- c("progesa-male-lmm", "progesa-female-lmm", "progesa-male-dlmm", "progesa-female-dlmm", "finngen-male-dlmm", "finngen-female-dlmm", 
          "findonor-male-dlmm", "findonor-female-dlmm", "progesa-both-dt", "progesa-both-rf")
 
-dummy_ids <- c("progesa-male-dummy", "progesa-female-dummy")
+dummy_ids <- c("progesa-male-baseline", "progesa-female-baseline")
 dummy2_ids <- c("finngen-male-stratified", "finngen-male-most-frequent", "finngen-male-prior", "finngen-male-uniform", "finngen-male-deferred")
 
 all_ids <- c(ids, dummy_ids)
@@ -246,6 +246,15 @@ get_sex <- function(s) {
   str_split(s, "-", simplify = TRUE)[,2]   # The second part is the sex
 }
 
+get_thresholds <- function(df,  p=parameters, thresholds = seq(0.1, .9, .1), id=id) {
+  sex <- get_sex(id)
+  tprs <- map_dfr(thresholds, function(t) get_rates(df, t))
+  tprs$probability = thresholds
+  
+  tprs <- tprs %>% rowwise() %>% mutate(get_cost(TPR, FPR, TPR, FPR, sex=sex, p=p), Id=id) %>% ungroup()
+  return(tprs)
+}
+
 get_optimal_thresholds <- function(df, p=parameters, thresholds = seq(0.1, .9, .1), id=id) {
   
   # pred.classes <- data.frame(matrix(nrow=nrow(df),ncol=length(thresholds)))
@@ -266,11 +275,9 @@ get_optimal_thresholds <- function(df, p=parameters, thresholds = seq(0.1, .9, .
   #   sen <- c(sen,conf$byClass['Sensitivity']) 
   #   spe <- c(spe, 1 - conf$byClass['Specificity']) 
   # }
-  sex <- get_sex(id)
-  tprs <- map_dfr(thresholds, function(t) get_rates(df, t))
-  tprs$probability = thresholds
+
+  tprs <- get_thresholds(df, p, thresholds, id)
   
-  tprs <- tprs %>% rowwise() %>% mutate(get_cost(TPR, FPR, TPR, FPR, sex=sex, p=p)) %>% ungroup()
   #print(tprs)
   row <- tprs %>% slice_min(E6, with_ties=FALSE)
   E6 <- row %>% pull(E6)
[…]

### 2021-08-24 src/article_figures.Rmd
Tried new economic parameters, and other changes.
Order variables by regression coefficient.

Extracted some functions to an R script.

@@
 importance from random forest algorithm.
+
+```{r Read FinnGen DLMM posteriors, include=FALSE}
+finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
+finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
+male_posterior <- finngenn_male_raw$samples
+female_posterior <- finngenn_female_raw$samples
+variables <- finngenn_male_raw$pretty_variable_names
+# Give a more descriptive name to this variable
+#variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
+variables <- recode(variables,    # Note that recode uses (stupidly) oldname = newname order, unlike many other functions in tidyverse
+                    `Previous Hb deferral` = "Previous donation deferred",
+                    `Polygenic risk score` = "Polygenic score",
+                    `RNF43 mutant` = "RNF43")
+```
 
@@
 
-```{r Read FinnGen DLMM posteriors, include=FALSE}
-finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
-finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
-male_posterior <- finngenn_male_raw$samples
-female_posterior <- finngenn_female_raw$samples
-variables <- finngenn_male_raw$pretty_variable_names
-# Give a more descriptive name to this variable
-#variables <- replace(variables, variables=="Previous Hb deferral", "Previous donation deferred")
-variables <- recode(variables,    # Note that recode uses (stupidly) oldname = newname order, unlike many other functions in tidyverse
-                    `Previous Hb deferral` = "Previous donation deferred",
-                    `Polygenic risk score` = "Polygenic score",
-                    `RNF43 mutant` = "RNF43")
-```
-
 ```{r, eval=FALSE}
@@
 
+  # Get pretty variable names in the order of female regression coefficients and secondarily by the importances
+  get_ordered_levels <- function(df) {
+    df <- df %>%
+      filter(gender %in% c("female", "both")) %>%
+      pivot_wider(id_cols=Pretty, names_from = method, values_from = mean) %>%
+      arrange(desc(`Regression coefficient`), desc("Importance"))
+    print(df)
+    v <- unique(as.character(df$Pretty))
+    return(v)
+  }
+
+  # Get pretty variable names in the order of average regression coefficient over sex
+  get_ordered_levels2 <- function(df) {
+    df <- df %>%
+      #filter(method == "Regression coefficient") %>%
+      group_by(Pretty, method) %>%
+      summarise(mean=mean(mean), .groups="keep") %>%
+      ungroup() %>%
+      pivot_wider(id_cols=Pretty, names_from = method, values_from = mean) %>%
+      #mutate(`Regression coefficient`=replace_na(`Regression coefficient`, -1000)) %>%
+      arrange(desc(`Regression coefficient`), desc(Importance))
+    print(df)
+    v <- unique(as.character(df$Pretty))
+    return(v)
+  }
   #print(combined_variables)
+
+  # Order the variables
+  new_levels <- get_ordered_levels2(final)
+  new_levels <- rev(new_levels)
+  print(new_levels)
+  final <- final %>% mutate(Pretty = fct_relevel(Pretty, new_levels))
 
@@
 
-#all_ids <- c(ids, dummy_ids)
-
-get_name_of_data <- function(s) {
-  s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
-  # Note that recode stupidly uses order "old"="new"
-  s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
-  return(s)
-}
-
-# Show AUROC, AUPR, F1, E6, and E12 as forest plot
-create_performance_forest_plot <- function() {
-  new_names <- c(`€/donation, 6 month deferral` = "E6", `€/donation, 12 month deferral` = "E12")
-  if (FALSE) {
-    # df1 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data.csv")
-    # df2 <- read_csv("~/FRCBS/results-for-eba-paper/raw_data_dummy.csv")
-    # df <- bind_rows(df1, df2)
-    # #  new_names <- c(`Economic effect: 6 month deferral` = "E6", `Economic effect: 12 month deferral` = "E12")
-    # 
-    # df <- df %>%
-    #   mutate(Id=factor(Id, levels=rev(all_ids)),
-    #          type=factor(type, levels=c("AUROC", "AUPR", "F1", "E6", "E12")),
-    #          sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
-    #          model=factor(str_split_fixed(Id, "-", 3)[,3])) %>%
-    #   mutate(type = fct_recode(type, !!!new_names))
-    # 
-    # thresholds <- get_threshold_df()
-    # df <- bind_rows(df, thresholds)
-    # df <- df %>% 
-    #   mutate(model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "dummy", "t6", "t12")))
-    #write_csv(df, path = "~/FRCBS/results-for-eba-paper/raw_data_all.csv")
-  } else {
-#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
-#    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
-    df <- read_tsv(all_results_filename)
-
-    df <- df %>% 
-      filter(variable %in% c("AUROC", "AUPR", "F1", "E6", "E12", "threshold6", "threshold12")) %>%
-      pivot_wider(names_from=c("type"), values_from=value) %>%
-      mutate(low = ifelse(is.na(low), value, low),     # To make the plot work, copy low=high=value for thresholds
-             high = ifelse(is.na(high), value, high))
-
-    df <- df %>%
-      mutate(Id=factor(Id, levels=rev(all_ids)),
-             sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
-             model=case_when(variable == "threshold6" ~ "t6",
-                             variable == "threshold12" ~ "t12",
-                             TRUE ~ str_split_fixed(Id, "-", 3)[,3]),
-             model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "baseline", "t6", "t12")),
-             variable = recode(variable, "threshold6" = "Threshold", "threshold12" = "Threshold"),
-             variable = factor(variable, levels=c("AUROC", "AUPR", "F1", "E6", "E12", "Threshold")),
-             variable = fct_recode(variable, !!!new_names))
-
-  }
-
-  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
-  xrange <- as.numeric(df %>% filter(variable %in% names(new_names)) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
-  xrange <- rep(xrange, 2)
-  mytype <- factor(rep(c("E6", "E12"), each=2), levels=c("AUROC", "AUPR", "F1", "E6", "E12"))
-  dummy <- tibble(Id="progesa-female-lmm", value=xrange, low=xrange, high=xrange, variable=mytype) # Dummy table to use the same x-axis limits for E6 and E12
-
-  dummy <- dummy %>%
-    mutate(variable = fct_recode(variable, !!!new_names))
-  #print(dummy)
-  #print(df)
-
-  #key_size <- grid::convertX(theme_get()$legend.key.size, "mm", valueOnly=TRUE)
-  #print(key_size)
-  key_size <- 1.8   # This results in box of size 6 mm
-  #key_size <- unit(6.096, "mm")
-  #key_size <- unit(1.2, "lines")
-  #grid::convertX(grid::unit(1.2, "lines"), "mm")
-  g <- df %>% ggplot(aes(y=Id, x=value, xmin=low, xmax=high)) + 
-    ggstance::geom_pointrangeh(aes(colour=sex, shape=model, fill=model)) + 
-    #geom_pointrange(aes(colour=sex, fill=sex, shape=model), key_glyph = draw_key_rect) + 
-    labs(x="Value", y="Data and model", colour="Sex", shape="Model", fill="Deferral length") + 
-    scale_y_discrete(labels=get_name_of_data) +
-    scale_colour_discrete(breaks = c("male", "female", "both"), labels=c("Male", "Female", "Both")) +
-    scale_fill_manual(breaks=c("t6", "t12"), labels=c("Six months", "Twelve months"), values=1:7) + #breaks = new_names) +
-    scale_shape_manual(breaks=c("lmm", "dlmm", "dt", "rf", "baseline"),
-                       labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest", "Baseline"),
-                       values=c("square", "circle", "triangle", "diamond", "plus", "cross", "asterisk")) +
-    # scale_fill_manual(breaks=c("lmm", "dlmm", "dt", "rf"), 
-    #                    labels=c("Linear mixed model", "Dynamic linear mixed model", "Decision tree", "Random forest"),
-    #                    values=c("square", "circle", "triangle", "diamond", "cross", "asterisk")) +
-    #scale_colour_manual(values = c("female" = 1, "male" = 2, "both" = 3)) +
-    guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0, size=key_size)),
-           fill = guide_legend(override.aes = list(shape = c("cross", "asterisk"), linetype=0))) +
-    #guides(colour = guide_legend(override.aes = list(shape = 15, linetype=0))) +
-    #theme(legend.key.size = theme_get()$legend.key.size) +
-    geom_blank(data=dummy) +
-    facet_wrap("variable", scales="free_x") +
-    #theme(legend.position = "right", legend.direction = "vertical") 
-    theme(legend.position = "bottom", legend.direction = "vertical")
-  return(list(g=g, df=df))
-}
-
+source("article_figures_functions.R")
 ```
@@
 ```{r}
-res <- create_performance_forest_plot()
+df <- read_tsv(all_results_filename)
+res <- create_performance_forest_plot(df)
 g <- res$g
@@
   if (is.null(probability_of_deferral_threshold)) { # predict deferral based on Hb
+    hlinetype <- "dashed"
     df <- df %>% 
@@
   } else { # predict deferral based on probability of deferral
+    cat(sprintf("Probability threshold is %.2f\n", probability_of_deferral_threshold))
+    hlinetype <- "dotted"
     df <- df %>% 
@@
   r2 <- summary(lm(predicted ~ observed, data=df))$r.squared
+  #print(r2)
   r2_text <- sprintf("%.2f", r2)
+  cat(sprintf("R2 value is %s\n", r2_text))
 
   df <- df %>%
-    mutate(confusion_class = factor(ifelse(deferral == 1, 
-                                    ifelse(new_predicted_label == 1, "True positive", "False negative"),
-                                    ifelse(new_predicted_label == 1, "False positive", "True negative")),
+    mutate(confusion_class = factor(
+      case_when(
+        deferral == 1 & new_predicted_label == 1 ~ "True positive",
+        deferral == 1 & new_predicted_label == 0 ~ "False negative",
+        deferral == 0 & new_predicted_label == 1 ~ "False positive",
+        deferral == 0 & new_predicted_label == 0 ~ "True negative"
+      ),
+      # ifelse(deferral == 1, 
+      #        ifelse(new_predicted_label == 1, "True positive", "False negative"),
+      #        ifelse(new_predicted_label == 1, "False positive", "True negative")),
            levels=c("True positive", "False negative", "False positive", "True negative")))
@@
 
-  cat(sprintf("Incorrectly predicting %.1f%% of accepted donations as deferrals\n", 100*fp/accepted))
-  cat(sprintf("Correctly predicting %.1f%% of deferred donations as deferrals\n", 100*tp/deferrals))
+  cat(sprintf("Incorrectly predicting %.1f%% of accepted donations as deferrals (FPR)\n", 100*fp/accepted))
+  cat(sprintf("Correctly predicting %.1f%% of deferred donations as deferrals (TPR)\n", 100*tp/deferrals))
 
@@
     geom_vline(xintercept = hb_threshold, linetype = "dashed") +
-    geom_hline(yintercept = prediction_threshold, linetype = "dashed") +
-    annotate("text", x=120, y=190, color="black", label=bquote(R^2==.(r2_text)), fontface="bold") +
+    geom_hline(yintercept = prediction_threshold, linetype = hlinetype) +
+    annotate("text", x=xymin+10, y=xymax-10, color="black", label=bquote(R^2==.(r2_text)), fontface="bold") +
     theme(legend.position = "bottom")
@@
 finngenn_male_raw <- get_raw_result_list("finngen-male-dlmm")
-
+finngenn_female_raw <- get_raw_result_list("finngen-female-dlmm")
 ```
 
-```{r}
+```{r Male scatter plot}
 g1 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, NULL)
-g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.04)
+g2 <- create_classification_scatter_plot(finngenn_male_raw$comp_df, 135, 0.08)
 if (save_figs) {
@@
 g2
+```
+
+```{r Female scatter plot}
+f1 <- create_classification_scatter_plot(finngenn_female_raw$comp_df, 125, NULL)
+f2 <- create_classification_scatter_plot(finngenn_female_raw$comp_df, 125, 0.02)
+if (save_figs) {
+  ggsave(filename="classification-scatter-plot-hb-finngen-female-dlmm.pdf", title="Classifiction scatter plot finngen-female-dlmm", path=fig_path, 
+         plot=f1, dpi=600, units="mm", width=180)
+  ggsave(filename="classification-scatter-plot-probability-finngen-female-dlmm.pdf", title="Classifiction scatter plot finngen-female-dlmm", path=fig_path, 
+         plot=f2, dpi=600, units="mm", width=180)
+}
+f1
+f2
 ```

### 2021-08-24 src/new_economic_parameters.Rmd
Tried new economic parameters, and other changes.
Order variables by regression coefficient.

Extracted some functions to an R script.

---
title: "Untitled"
author: "Jarkko Toivonen"
date: "8/24/2021"
output: 
  md_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("economic_effect_functions.R")
source("article_figures_functions.R")
library(knitr)
library(tidyverse)
save_figs = TRUE
save_tables = FALSE
dir <- "~/FRCBS/interval-sims/src/new_cost_parameters"
fig_path <- dir
table_path <- dir
all_results_filename <- sprintf("%s/raw-data-2021-08-24.tsv", dir)
compute=FALSE
```

## R Markdown

Try the economic effect calculations with the new parameters values.

```{r}
new_fixed_parameters <- list(Pm=2.287, Pd=18.08, F=1, Fn=0.1066, rloss=0)

new_parameters <- c(data_parameters, new_fixed_parameters)
if (compute) {
  all_results <- process_all_data(all_ids, parameters=new_parameters)
  write_tsv(all_results, file=all_results_filename)
} else {
  all_results <- read_tsv(file=all_results_filename)  
}

```
The cost surface below is for the random forest on eProgesa data.

```{r Cost surface}
g <- draw_surface(results=all_results %>% filter(Id == "progesa-both-rf"), cost_function=do.call(cost_func_factory, new_parameters[c("Pm", "Pd", "F", "Fn", "rloss", "d")]))
if (save_figs) {
  filename <- sprintf("%s/cme_cost_surface_werr.pdf", fig_path)
  ggsave(filename=filename,  width = 180,  height = 120, units="mm", dpi=600, scale=1.0, device=cairo_pdf)
}
g
```

```{r Performance plot}
# finngen-male-dlmm  E12 a12 and f1 values are for some reason NA
#df <- all_results %>% 
#  replace_na(list(value=0))

res <- create_performance_forest_plot(all_results)
g <- res$g
df <- res$df
if (save_figs) {
  filename <- sprintf("%s/performance-forest-plot.pdf", fig_path)
  ggsave(filename=filename, #  title="Performance forest plot", 
         plot=g, dpi=600, units="mm", width=180, device=cairo_pdf)
}
g
```

### 2021-08-24 src/article_figures_functions.R

@@ -1,0 +1,80 @@
+#all_ids <- c(ids, dummy_ids)
+
+get_name_of_data <- function(s) {
+  s <- str_split_fixed(s, "-", 3)[,1]  # Extract the name of the data set
+  # Note that recode stupidly uses order "old"="new"
+  s <- recode(s, "progesa" = "eProgesa", "finngen" = "Biobank", "findonor" = "FinDonor")  # Prettify
+  return(s)
+} 
+
+# Show AUROC, AUPR, F1, E6, and E12 as forest plot
+create_performance_forest_plot <- function(df) {
+  new_names <- c(`€/donation, 6 month deferral` = "E6", `€/donation, 12 month deferral` = "E12")
+  #    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-11-30.csv")
+  #    df <- read_csv("~/FRCBS/results-for-eba-paper/raw-data-2020-12-01.csv")
+  
+  df <- df %>% 
+    filter(variable %in% c("AUROC", "AUPR", "F1", "E6", "E12", "threshold6", "threshold12")) %>%
+    pivot_wider(names_from=c("type"), values_from=value) %>%
+    mutate(low = ifelse(is.na(low), value, low),     # To make the plot work, copy low=high=value for thresholds
+           high = ifelse(is.na(high), value, high))
+  
+  df <- df %>%
+    mutate(Id=factor(Id, levels=rev(all_ids)),
+           sex=factor(str_split_fixed(Id, "-", 3)[,2], levels=c("female", "both", "male")),
+           model=case_when(variable == "threshold6" ~ "t6",
+                           variable == "threshold12" ~ "t12",
+                           TRUE ~ str_split_fixed(Id, "-", 3)[,3]),
+           model = factor(model, levels=c("lmm", "dlmm", "dt", "rf", "baseline", "t6", "t12")),
+           variable = recode(variable, "threshold6" = "Threshold", "threshold12" = "Threshold"),
+           variable = factor(variable, levels=c("AUROC", "AUPR", "F1", "E6", "E12", "Threshold")),
+           variable = fct_recode(variable, !!!new_names))
+  
+  
+  
+  
+  # Below we use a dummy table and geom_blank to have the same xlimits for E6 and E12 panels
+  xrange <- as.numeric(df %>% filter(variable %in% names(new_names)) %>% summarise(low=min(low), high=max(high))) # range of x-axis for E6 and E12
+  xrange <- rep(xrange, 2)
[…]

### 2021-08-24 src/economic_effect_functions.R

@@ -10,7 +10,7 @@ ids <- c("progesa-male-lmm", "progesa-female-lmm", "progesa-male-dlmm", "progesa
          "findonor-male-dlmm", "findonor-female-dlmm", "progesa-both-dt", "progesa-both-rf")
 
 dummy_ids <- c("progesa-male-baseline", "progesa-female-baseline")
-dummy2_ids <- c("finngen-male-stratified", "finngen-male-most-frequent", "finngen-male-prior", "finngen-male-uniform", "finngen-male-deferred")
+#dummy2_ids <- c("finngen-male-stratified", "finngen-male-most-frequent", "finngen-male-prior", "finngen-male-uniform", "finngen-male-deferred")
 
 all_ids <- c(ids, dummy_ids)
 
@@ -26,9 +26,9 @@ get_mikkos_cost_constants <- function() {
 }
 
 data_parameters <- get_mikkos_cost_constants()
-fixed_parameters <- list(Pm=2, Pd=60, Fn=0.1066, rl=0)
+fixed_parameters <- list(Pm=2, Pd=60, Fn=0.1066, rloss=0)
 
-parameters <- c(data_parameters, fixed_parameters)
+all_parameters <- c(data_parameters, fixed_parameters)
 
 load_single <- function(filename) {
   names <- load(filename, verbose=FALSE)
@@ -140,8 +140,18 @@ get_cost_constants <- function(data) {
 }
 
 
+cost_func <- function(q, atot, Pm, Pd, F, Fn, rloss, d) {Pm * ((F*atot) / (F + Fn*(atot-1)) - 1 - d*q*rloss) - Pd*d*q}
+
+cost_func_simplified <- function(q, atot) {2 / ((1-0.1066)/atot + 0.1066)  - 2 - 60*0.03269718*q}
+
+cost_func_factory <- function(Pm, Pd, F, Fn, rloss, d) {
+  function(q, atot) {
+    return(Pm * ((F*atot) / (F + Fn*(atot-1)) - 1 - d*q*rloss) - Pd*d*q)
+  }
+}
+
 #get_cost <- function(TPR6, FPR6, TPR12, FPR12, d.=d, mr.=mr, fr.=fr, mdr.=mdr, fdr.=fdr, Pm.=Pm, Pd.=Pd, Fn.=Fn, rl.=rl) {
[…]

### 2021-08-27 src/new_economic_parameters.Rmd
Yet again tweaked economic parameter values.

@@
 table_path <- dir
-all_results_filename <- sprintf("%s/raw-data-2021-08-24.tsv", dir)
+all_results_filename <- sprintf("%s/raw-data-2021-08-27.tsv", dir)
 compute=FALSE
@@
 ```{r}
-new_fixed_parameters <- list(Pm=2.287, Pd=18.08, F=1, Fn=0.1066, rloss=0)
+new_fixed_parameters <- list(Pm=2.287, Pd=19.96, F=1, Fn=0.1066, rloss=0)
+#new_fixed_parameters <- list(Pm=2.287, Pd=18.08, F=1, Fn=0.1066, rloss=0)
 

### 2021-08-31 src/new_economic_parameters.Rmd
RF done by oversampling.

@@
 library(tidyverse)
-save_figs = TRUE
-save_tables = FALSE
+save_figs = FALSE
+#save_tables = FALSE
 dir <- "~/FRCBS/interval-sims/src/new_cost_parameters"
@@
 table_path <- dir
-all_results_filename <- sprintf("%s/raw-data-2021-08-27.tsv", dir)
+all_results_filename <- sprintf("%s/raw-data-2021-08-31.tsv", dir)
 compute=FALSE

### 2021-08-31 src/economic_effect_functions.R

@@ -365,7 +365,8 @@ get_data_frame <- function(id) {
     return(df)
   } else if (id == "progesa-both-rf") {
 #    df <- load_single("~/FRCBS/interval-sims/rrfFit_roc_validate_probs.rdata")
-    df <- load_single("~/FRCBS/interval-sims/rrfFit_roc_test_probs.rdata")
+#    df <- load_single("~/FRCBS/interval-sims/rrfFit_roc_test_probs.rdata")
+    df <- load_single("/home/toivoja/FRCBS/interval_prediction/oversampling_results/rrfFit_roc_test_probs.rdata")
     return(as_tibble(df))
   } else if (id == "progesa-both-dt") {
     #df <- as_tibble(load_single("~/FRCBS/interval-sims/rrfFit__dtree_roc_validate_probs.rdata"))
@@ -712,7 +713,7 @@ draw_surface <- function(scale=FALSE, results=NULL, cost_function = cost_func_si
     results <- results %>%
       filter(#Id=="progesa-both-rf",
              variable %in% c("E6", "E12", "a6", "a12", "q6", "q12")) %>%
-      extract(col=variable, regex = "([a-zA-Z]+)([0-9]+)", into=c("variable", "month")) %>% 
+      tidyr::extract(col=variable, regex = "([a-zA-Z]+)([0-9]+)", into=c("variable", "month")) %>% 
       mutate(month=as.integer(month)) %>%
       pivot_wider(names_from=c(variable, type))
     print(results)

### 2021-09-16 src/article_figures.Rmd
Run with updated parameters.

@@
 table_path="~/FRCBS/results-for-eba-paper/table"
-all_results_filename <- sprintf("%s/raw-data-2021-06-11.tsv", table_path)
+all_results_filename <- sprintf("%s/raw-data-2021-09-03.tsv", table_path)
 compute=FALSE
+
+new_fixed_parameters <- list(Pm=2.287, Pd=20.342, F=1, Fn=0.1066, rloss=0.167)
+new_parameters <- c(data_parameters, new_fixed_parameters)
 ```
@@
 ```
+The results with RF model:
+
+```{r}
+all_results %>% filter(Id=="progesa-both-rf", type=="value", variable %in% c("E6", "a6", "q6"))
+```
+
+## Combine the sex specific models
+
+What is the average interval extension, economic effect, and fraction of deferrals avoided, if
+we use both male and female models?
+
+```{r}
+tmp <- all_results %>% 
+  filter(str_detect(Id, "lmm"), type=="value", variable %in% c("E6", "a6", "q6")) %>% 
+  separate(Id, sep="-", into=c("data", "sex", "model"), remove=FALSE) %>% 
+  pivot_wider(names_from = variable, values_from = value)
+
+tmp %>% group_by(data, model) %>%
+  summarise(a6 = sum(a6) - 1,
+            E6 = sum(E6),
+            q6 = sum(q6)
+            )
+```
 
@@
   temp <- all_results %>% pivot_wider(names_from=c(variable, type))
-  filename <- sprintf("%s/pretty-data-2021-06-11.xlsx", table_path)
+  filename <- sprintf("%s/pretty-data-2021-09-03.xlsx", table_path)
   openxlsx::write.xlsx(temp, file = filename, overwrite=TRUE)
@@
 ```{r Cost surface}
-g <- draw_surface(results=all_results %>% filter(Id == "progesa-both-rf"))
+g <- draw_surface(results=all_results %>% filter(Id == "progesa-both-rf"), cost_function=do.call(cost_func_factory, new_parameters[c("Pm", "Pd", "F", "Fn", "rloss", "d")]))
 if (save_figs) {
@@
 #get_thresholds <- function(df,  p=parameters, thresholds = seq(0.1, .9, .1), id=id)
-thresholds <- map2_dfr(data_frames, all_ids, function(df, id) get_thresholds(df, parameters, threshold_range, id))
+thresholds <- map2_dfr(data_frames, all_ids, function(df, id) get_thresholds(df, new_parameters, threshold_range, id))
 
@@
   # Get variable importances from the RF object
-  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
+  rrfFit_roc <- load_single("~/FRCBS/interval_prediction/oversampling_results/rrfFit_roc.rdata")
+  #rrfFit_roc <- load_single("~/FRCBS/interval_prediction/results/rrfFit_roc.rdata")
   varimp <- get_rf_df(rrfFit_roc, descript, combined_variables)

### 2021-09-16 src/dummy_classifier.Rmd
Run with updated parameters.

@@
 library(knitr)
+source('../../Hb_predictor_container/src/validate_stan_fit.R', chdir=T)
 source("economic_effect_functions.R")
-source('../../Hb_predictor_container/src/validate_stan_fit.R', chdir=T)
 ```
@@
 
-```{r cars}
+```{r Logistic regression functions}
 
@@
 
-  result <- tibble(Deferred = glm.probs, 
-                   obs = validate$Hb_deferral)
+  result <- tibble(score = glm.probs, 
+                   original_label = validate$Hb_deferral)
 
@@
 
-```{r, eval=FALSE}
+```{r Old logistic function plot to convert between Hb threshold and probability for females, eval=FALSE}
 df1 <- make_threshold_conversion_data_frame(125, -1) 
@@
 
-```{r, eval=FALSE}
+```{r  Old logistic function plot to convert between Hb threshold and probability for males, eval=FALSE}
 df1 <- make_threshold_conversion_data_frame(135, -1) 
@@
 
-```{r}
+```{r Fit logistic regression}
 use_logistic_regression=TRUE
@@
 
-```{r}
+```{r Compute performance measures}
 #dummy_result <- suppressMessages(process_all_data(dummy2_ids, replicates=100))
-dummy_result <- suppressMessages(process_all_data(dummy2_ids))
+dummy_result <- suppressMessages(process_all_data(dummy_ids))
 ```
 
-```{r}
+```{r Show the result table}
 df <- dummy_result %>% 

### 2021-09-16 src/new_economic_parameters.Rmd
Run with updated parameters.

--- 2021-08-24 src/new_economic_parameters.Rmd
@@
 table_path <- dir
-all_results_filename <- sprintf("%s/raw-data-2021-08-24.tsv", dir)
+all_results_filename <- sprintf("%s/raw-data-2021-09-03.tsv", dir)
 compute=FALSE
@@
 ```{r}
-new_fixed_parameters <- list(Pm=2.287, Pd=18.08, F=1, Fn=0.1066, rloss=0)
+new_fixed_parameters <- list(Pm=2.287, Pd=20.342, F=1, Fn=0.1066, rloss=0.167)
+#new_fixed_parameters <- list(Pm=2.287, Pd=19.96, F=1, Fn=0.1066, rloss=0.167)
+#new_fixed_parameters <- list(Pm=2.287, Pd=18.08, F=1, Fn=0.1066, rloss=0)
 
@@
 ```{r Cost surface}
-g <- draw_surface(results=all_results %>% filter(Id == "progesa-both-rf"), cost_function=do.call(cost_func_factory, new_parameters[c("Pm", "Pd", "F", "Fn", "rloss", "d")]))
+cost_function <- do.call(cost_func_factory, new_parameters[c("Pm", "Pd", "F", "Fn", "rloss", "d")])
+g <- draw_surface(results=all_results %>% filter(Id == "progesa-both-rf"), cost_function=cost_function)
 if (save_figs) {

### 2021-09-16 src/economic_effect_functions.R

@@ -258,6 +258,7 @@ get_sex <- function(s) {
   str_split(s, "-", simplify = TRUE)[,2]   # The second part is the sex
 }
 
+# Returns a dataframe with a row for each threshold
 get_thresholds <- function(df,  p, thresholds = seq(0.1, .9, .1), id=id) {
   sex <- get_sex(id)
   tprs <- map_dfr(thresholds, function(t) get_rates(df, t))
